---
title: "data_analysis_oddball_leap_231023"
author: "Nico Bast"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

# setup

- changes in July 2024
  - epoched data: every "trial" now consists of 8 trials as epoch to capture the longer pupillary response
  - linear time-invariant filter: data is filtered with a custom filter that considers pupillary responses within epochs as pulses from other trials that are corrected with a gamma distribution estimated in an independent data set of an oddball task

```{r setup, include=FALSE}

#chunk options
knitr::opts_chunk$set(echo = TRUE) #show code in markdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE) #warning to false
eval_bayes=FALSE #evaluate chunks with Bayesian modelling - FALSE to speed up markdown

## REQUIRED PACKAGES

#analysis
require(lme4) #linear mixed models
require(lmerTest) #p-value for fixed effects in LMM
require(emmeans) #estimated marginalized means
require(performance) #r²_nakagawa - marginalized and conditionalk R² 

#packages for MMN
require(pbapply) #progress bar apply
require(zoo) #rolling mean
require(data.table) #rbindlist

#visualization
require(kableExtra) #APA tables
library(greta) #Bayesian modelling
require(bayesplot) #plotting Bayesian
require(bayestestR) #credible interval ci, point estimates Bayesian
require(DiagrammeR) #plotting Bayesian
require(ggplot2) #plot posterior sampling estimates

require(R.matlab) # read matlab files
require(effectsize) # cohens D

require(RColorBrewer) #color palettes
require(gridExtra) #arrange plots
require(wesanderson) #custom color palettes

require(eegkit) #standard coordinates of EEG channels - data(eegcoord)
require(eegUtils)

require(car) #Anova function for type 3 SumSq 

#INFO: eegUtils is not on CRAN and thus loaded with remote packages from github
# install.packages("remotes")
# remotes::install_github("craddm/eegUtils")


#paths
project_path<-getwd()
#project_path<-"C:/Users/nico/PowerFolders/project_oddball_LEAP"
localation_of_EEG_allchannel_data<-"Z:/nico/backup_data/data_AIMS/mismatch_negativity/leap_oddball_eeg_list_allchannels.rdata"


##define parameters of posterior sampling in Bayesian modeling
number_of_warmups <- 4000
number_of_iterations <- 8000
credible_interval<-0.89 #credible interval Bayesian estimates


##define percentage of data sampling for visualization
fraction_to_sample<-0.2 ###relevant for scalp topography (0-1)
sampling_factor<-1/fraction_to_sample ### releavnt for MMN, pupil size change figures

###graphic settings

  #graphic parameters
  theme_set(theme_bw())
  
  #define color palette
  custom_condition_colors <- wes_palette('FantasticFox1',5,type='discrete')[2:5]
  custom_contrast_colors <- c(brewer.pal(6, "Blues")[4:6],brewer.pal(6, "Oranges")[4:6])
  custom_group_color <- brewer.pal(3, "Dark2")[1:2]
 
  #RColorBrewer::display.brewer.all()

```

# load data

Eye tracking and EEG data in separate file that are already preprocessed, see

<li> [Eye-Tracking preprocessing]( https://github.com/nicobast/oddball_LEAP/blob/master/code/data_preprocessing_oddball_leap.Rmd) </li>
<li> [EEG preprocessing](https://github.com/nicobast/oddball_LEAP/blob/master/code/read_eeg_data.R) </li>

```{r load_data_ET_EEG}

## pupil data
## - df_epoch: pupillary data per trial
## - df_dem: demographic data on participant level
## - d_timepoint: data for each timepoint
load(file=paste0(project_path,"/data/mmn_leap_pd_final_dfs_epoch_19072024"))

## EEG data - reduced to FZ channel - see read_eeg_data.R
#load(paste0(project_path,"/data/leap_oddball_eeg_list.rdata"))
list_eeg_redux<-readRDS(paste0(project_path,"/data/leap_oddball_eeg_list_redux.rds"))

```

# calculate mismatch negativity (MMN)

MMN amplitude as negativity in timeframe 50-350 ms (see function definition)

```{r calculate mismatch negativity}

###---> calculate MMN ####

fun_estimate_mmn<-function(one_set){

  #testing
  #one_set<-list_eeg[[100]]

  ###restrict to timeframe - 50-350ms
  one_set<-one_set[one_set$times>50 & one_set$times<350,]

  mmn_trial<-with(one_set,by(Fz,trial_counter,function(channel_data){

      channel_data_smooth<-rollmean(channel_data,k=20) #rolling mean of 20ms
      min(channel_data_smooth)

  }))

  trial_counter<-names(mmn_trial)
  mmn<-as.numeric(mmn_trial)
  mmn_per_trial<-data.frame(trial_counter,mmn)
  return(mmn_per_trial)

}

list_mmn<-pblapply(list_eeg_redux,fun_estimate_mmn)

##--> create data frame
ID<-substr(names(list_mmn),10,22)
condition<-substr(names(list_mmn),1,8)

list_mmn<-pbmapply(function(x,y,z){
  x$condition<-y
  x$ID<-z
  return(x)
  },x=list_mmn,y=condition,z=ID,SIMPLIFY = F)

df_mmn<-rbindlist(list_mmn)
df_mmn$trial_counter<-as.numeric(df_mmn$trial_counter)
df_mmn$condition<-as.factor(df_mmn$condition)

```

# merge eye-tracking and EEG data

```{r merge ET and EEG}

##prepare EEG data
df_mmn$ID<-sub('/','',df_mmn$ID)
df_mmn$ID<-sub('_','',df_mmn$ID)

    #with(df_timepoint,table(subjects,wave))

    #eye-tracking data with MMN data
    table(df_timepoint$subjects %in% df_mmn$ID)
    table(unique(df_mmn$ID) %in% df_timepoint$subjects)

#create merge variables    
df_mmn$merge_id<-interaction(df_mmn$ID,df_mmn$trial_counter)
df_epoch$merge_id<-interaction(df_epoch$subjects,df_epoch$EventCounter)

#merge - per-trial level (df_epoch)
df_epoch<-merge(df_epoch,df_mmn,by='merge_id',all.x=T)

    table(is.na(df_mmn$mmn))
    table(is.na(df_epoch$rpd_response))
    table(is.na(df_epoch$mmn)) #missing in MMN --> to be expected
    ##--> not all trials analyzed in eye-tracking were selected for EEG
    
##some trials are not correctly matched - correct for that
with(df_epoch,table(condition,EventData_epoch))
df_epoch_mmn<-df_epoch[(df_epoch$EventData_epoch=='201' & df_epoch$condition=='standard'|
                 df_epoch$EventData_epoch=='203' & df_epoch$condition=='duration'|
                 df_epoch$EventData_epoch=='202' & df_epoch$condition=='frequenc'|
                 df_epoch$EventData_epoch=='204' & df_epoch$condition=='combined'),]

###merge per-participant level (df_timepoint)
df_mmn_timepoint<-aggregate(mmn~id+EventData_epoch,data=df_epoch_mmn,FUN=mean,na.rm=T)
df_mmn_timepoint<-reshape(df_mmn_timepoint, idvar = "id", timevar = "EventData_epoch", direction = "wide")

df_timepoint<-merge(df_timepoint,df_mmn_timepoint,by='id',all.x = T)
rm(df_mmn_timepoint)

###which participants have MMN data
table(!is.na(df_timepoint$mmn.201),df_timepoint$t1_diagnosis) ##

```

# retained trials in EEG

calculate the number of retained trials in EEG data for the preprocessed set

```{r mmn retained trials}

unique_trials<-pbsapply(list_eeg_redux,function(x){length(unique(x$trial_counter))})
set_condition<-pbsapply(list_eeg_redux,function(x){unique(x$condition)})
set_ID<-pbsapply(list_eeg_redux,function(x){unique(x$ID)})

set_ID<-ifelse(set_condition=='frequenc',substr(set_ID,2,13),substr(set_ID,1,12))

#retained trials per condition per participant
xtabs(unique_trials ~ set_ID + set_condition,data=data.frame(unique_trials,set_ID,set_condition)[set_ID %in% df_dem$subjects,])

#retained trials per condition
by(unique_trials[set_ID %in% df_dem$subjects],set_condition[set_ID %in% df_dem$subjects],summary)
#--> selected per condition based on lowest number of retained trials
mean(unique_trials)
sd(unique_trials)

mean(unique_trials)/(1400*0.06)

```

# calculate medication 

```{r medication usage}

#define medication usage
df_dem$med_use<-ifelse(df_dem$t1_med_use %in% c('Medication use','Other medication use'),TRUE,FALSE)

df_timepoint$med_use<-ifelse(df_timepoint$med_use %in% c('Medication use','Other medication use'),TRUE,FALSE)

df_epoch$med_use<-ifelse(df_epoch$med_use %in% c('Medication use','Other medication use'),TRUE,FALSE)

```

# load and merge luminance data

- see script "luminance_data_extraction.R"

```{r}

#load
df_luminance<-readRDS(paste0(project_path,'/data/stimulus_luminance.rds'))

#merge
df_luminance$trials<-round(df_luminance$trials)
df_luminance<-df_luminance[!duplicated(df_luminance$trials),]
df_epoch<-merge(df_epoch,df_luminance,by.x='EventCounter_epoch',by.y='trials')

```


# descriptives table of sample

generates a table on characteristics of the sample between groups

```{r samples description}

#### sample descriptives - PARTICIPANT specific ####

fun_return_descriptive<-function(variable,group,rounding=2,scaling=1){
  mean_values<-by(variable,group,function(x){round(mean(x,na.rm=T),rounding)})
  sd_values<-by(variable,group,function(x){round(sd(x,na.rm=T),rounding)})
  min_values<-by(variable,group,function(x){round(min(x,na.rm=T),rounding)})
  max_values<-by(variable,group,function(x){round(max(x,na.rm=T),rounding)})
  mean_values<-mean_values*scaling
  sd_values<-sd_values*scaling
  min_values<-min_values*scaling
  max_values<-max_values*scaling
  paste0(mean_values,'/',sd_values,' ','[',min_values,'-',max_values,']')
}

#standard descriptives
n_group<-with(df_dem,by(subjects,t1_diagnosis,function(x){length(unique(x))}))
gender<-with(df_dem,by(t1_sex,t1_diagnosis,function(x){paste0(table(x)[2],'/',table(x)[1])}))
timepoints<-with(df_dem,by(timepoints,t1_diagnosis,function(x){paste0(table(x)[1],'/',table(x)[2],'/',table(x)[3])}))
age<-with(df_dem,fun_return_descriptive(variable=t1_ageyrs,group=t1_diagnosis))
iq<-with(df_dem,fun_return_descriptive(variable=t1_fsiq,group=t1_diagnosis))
piq<-with(df_dem,fun_return_descriptive(variable=t1_piq,group=t1_diagnosis))
viq<-with(df_dem,fun_return_descriptive(variable=t1_viq,group=t1_diagnosis))

#clinical variables
srs<-with(df_dem,fun_return_descriptive(variable=t1_srs_rawscore,group=t1_diagnosis))
rbs_total<-with(df_dem,fun_return_descriptive(variable=t1_rbs_total,group=t1_diagnosis))
sdq_total<-with(df_dem,fun_return_descriptive(variable=t1_sdq_total_difficulties_p,group=t1_diagnosis))
adhd_inatt<-with(df_dem,fun_return_descriptive(variable=adhd_inatt,group=t1_diagnosis))
adhd_hyper<-with(df_dem,fun_return_descriptive(variable=adhd_hyper,group=t1_diagnosis))
anx_beck<-with(df_dem,fun_return_descriptive(variable=anx_beck,group=t1_diagnosis))
dep_beck<-with(df_dem,fun_return_descriptive(variable=dep_beck,group=t1_diagnosis))

#medication
medi_use<-with(df_dem,by(med_use,t1_diagnosis,function(x){paste0(table(x)[2],'/',table(x)[1])}))

##data quality
#precision<-with(df_dem,fun_return_descriptive(variable=Precision,group=t1_diagnosis))
#accuracy<-with(df_dem,fun_return_descriptive(variable=Accuracy,group=t1_diagnosis))
#sampling_rate<-with(df_dem,by(sampling_rate,t1_diagnosis,function(x){paste0(table(x)[2],'/',table(x)[1])}))
missing_data_trial<-with(df_dem,fun_return_descriptive(variable=missing_data_trial,group=t1_diagnosis,rounding=4,scaling = 100))
sampling_rate<-with(df_timepoint,by(sampling_rate,t1_diagnosis,function(x){paste0(table(x)[2],'/',table(x)[1])}))

#et vars
center_dev_id<-with(df_dem,fun_return_descriptive(variable=center_dev,group=t1_diagnosis,rounding=4,scaling=100))
#screen_dist_id<-with(df_dem,fun_return_descriptive(variable=screen_dist,group=t1_diagnosis))
#pd_baseline_id<-with(df_dem,fun_return_descriptive(variable=pd_baseline,group=t1_diagnosis))
pd_baseline_id<-with(df_dem,fun_return_descriptive(variable=pd_res,group=t1_diagnosis))

rpd_auc_id<-with(df_epoch,fun_return_descriptive(variable=rpd_res_response,group=t1_diagnosis,rounding=4,scaling=100))

descriptives_table<-rbind(n_group,gender,timepoints,age,iq,piq,viq,
                          missing_data_trial,sampling_rate,
                          center_dev_id,
                          #screen_dist_id,
                          pd_baseline_id,medi_use,
                          srs,rbs_total,sdq_total,adhd_inatt,adhd_hyper,anx_beck,dep_beck
                          )

##group differences
groupdiff_p<-c(NA,
               round(as.numeric(with(df_dem,chisq.test(t1_sex,t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,chisq.test(timepoints,t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_ageyrs~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_fsiq~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_piq~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_viq~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(missing_data_trial~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_timepoint,chisq.test(sampling_rate,t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(center_dev~t1_diagnosis))['p.value']),3),
               #round(as.numeric(with(df_dem,t.test(screen_dist~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(pd_baseline~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,chisq.test(med_use,t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_srs_rawscore~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_rbs_total~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_sdq_total_difficulties_p~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(adhd_inatt~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(adhd_hyper~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(anx_beck~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(dep_beck~t1_diagnosis))['p.value']),3)
               )

groupdiff_p[which(groupdiff_p==0)]<-'<0.001'
groupdiff_p[which(is.na(groupdiff_p))]<-'-'

row_names<-c('n','gender (M/F)','timepoints (1/2/1+2)','age (in years)','IQ','perceptual IQ','verbal IQ',
             'missing data per trial (%)','sampling rate (300Hz / 120Hz)',
             'gaze center deviation (%)',
             #'screen distance (mm)',
             'mean pupil size (mm)',
             'medication (yes/no)',
             'SRS (total)','RBS (total)','SDQ (total)','ADHD inattention','ADHD hyperactivity','BAI (anxiety)','BDI (depression)'
             # 'pupillary response: amplitude - standard (%)','pupillary response: amplitude - oddball pitch (%)','pupillary response: amplitude - oddball length (%)','pupillary response: amplitude - oddball both (%)',
             # 'pupillary response: AUC - standard (%)','pupillary response: AUC - oddball pitch (%)','pupillary response: AUC - oddball length (%)','pupillary response: AUC - oddball both (%)'
             )

descriptives_table<-cbind(row_names,descriptives_table,groupdiff_p)

#add group differences
#descriptives_table<-ifelse(descriptives_table=='0','<0.001',descriptives_table)

table_sample<-descriptives_table %>%
  kbl(caption = "Sample description",
      col.names = c('','autistic individuals (ASD)','non-autistic individuals (NTC)','group difference (p)'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")


descriptives_table

table_sample

save_kable(table_sample, file=paste0(project_path,'/output/table_sampledescription.html'))




```

# group differences of dependent variables (per participant)

```{r group differences per participant}

#BPS
anova(lm<-lm(scale(pd_res)~t1_diagnosis,df_timepoint))
summary(lm)

cohens_d(x=scale(pd_res)~t1_diagnosis,data=df_timepoint)

#significant covariate model - PD
lm<-lm(scale(pd_res)~t1_diagnosis+
         scale(ageyrs)+scale(center_dev)
       ,df_timepoint)
Anova(lm,type='3')
summary(lm)

# anova(lm<-lm(scale(rpd_response.201)~t1_diagnosis,df_timepoint))
# anova(lm<-lm(scale(rpd_response.202)~t1_diagnosis,df_timepoint))
# anova(lm<-lm(scale(rpd_response.203)~t1_diagnosis,df_timepoint))
# anova(lm<-lm(scale(rpd_response.204)~t1_diagnosis,df_timepoint))
# anova(lm<-lm(scale(mmn.201)~t1_diagnosis,df_timepoint))
# anova(lm<-lm(scale(mmn.202)~t1_diagnosis,df_timepoint))
# anova(lm<-lm(scale(mmn.203)~t1_diagnosis,df_timepoint))
# anova(lm<-lm(scale(mmn.204)~t1_diagnosis,df_timepoint))


```

# covariate effects on dependent variables (per participant)

```{r covariate effects on dependent variables per participant}

#BPS
c(summary(lm(scale(pd_res)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(pd)~scale(ageyrs),df_timepoint))[2,])

#MMN - standards 
c(summary(lm(scale(mmn.201)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.201)~scale(ageyrs),df_timepoint))[2,])

#MMN - pitch oddballs 
c(summary(lm(scale(mmn.202)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.202)~scale(ageyrs),df_timepoint))[2,])

#MMN - length oddballs 
c(summary(lm(scale(mmn.203)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.203)~scale(ageyrs),df_timepoint))[2,])

#MMN - pitch & length oddballs
c(summary(lm(scale(mmn.204)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.204)~scale(ageyrs),df_timepoint))[2,])


```

# associations of dependent variables (per participant)

correlation coefficients on per-participant level of

<li> pupillary measure: baseline pupil size (BPS) </li>
<li> pupillary measure: stimulus-evoked pupilalry response (SEPR) </li>
<li> EEG measure: mismatch negativity (MMN) </li>
<li> computational model: neural gain (NG) </li>

Event Data coding

<li> 201 = standard </li>
<li> 202 = pitch </li>
<li> 203 = length </li>
<li> 204 = pitch & length </li>


```{r associations of dependent variables}

attach(df_timepoint)

cor_table<-data.frame(t(cbind(

#SEPR -PD
cbind(cor.test(rpd_res_response.201,pd_res)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_res_response.202,pd_res)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_res_response.203,pd_res)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_res_response.204,pd_res)[c('data.name','estimate','p.value')]),

#SEPR - MMN --> uncorrelated
cbind(cor.test(rpd_res_response.201,mmn.201)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_res_response.202,mmn.202)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_res_response.203,mmn.203)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_res_response.204,mmn.204)[c('data.name','estimate','p.value')]),

#BPS - MMN --> substantial negative correlations
cbind(cor.test(pd_res,mmn.201)[c('data.name','estimate','p.value')]),
cbind(cor.test(pd_res,mmn.202)[c('data.name','estimate','p.value')]),
cbind(cor.test(pd_res,mmn.203)[c('data.name','estimate','p.value')]),
cbind(cor.test(pd_res,mmn.204)[c('data.name','estimate','p.value')])

)))

#degree of freedom
summary(lm(mmn.201~pd_res,data=df_timepoint))
#confidence intervals
cor.test(rpd_res_response.201,pd_res)
cor.test(rpd_res_response.203,pd_res)
cor.test(rpd_res_response.201,mmn.201)
cor.test(pd_res,mmn.201)
cor.test(pd_res,mmn.202)
cor.test(pd_res,mmn.203)
cor.test(pd_res,mmn.204)


#format table
cor_table$data.name<-unlist(cor_table$data.name)
cor_table$estimate<-unlist(cor_table$estimate)
cor_table$p.value<-unlist(cor_table$p.value)

cor_table$estimate<-round(cor_table$estimate,3)
cor_table$p.value<-round(cor_table$p.value,3)

cor_table$data.name<-gsub('rpd_res_response','SEPR',cor_table$data.name)
cor_table$data.name<-gsub('pd_res','BPS',cor_table$data.name)
cor_table$data.name<-gsub('mmn','MMN',cor_table$data.name)
cor_table$data.name<-gsub('.201',' (standard)',cor_table$data.name)
cor_table$data.name<-gsub('.202',' (pitch oddball)',cor_table$data.name)
cor_table$data.name<-gsub('.203',' (length oddball)',cor_table$data.name)
cor_table$data.name<-gsub('.204',' (pitch+length oddball)',cor_table$data.name)


#Bonferroni correction
number_of_comparisons<-nrow(cor_table)
cor_table$p_adjust<-cor_table$p.value*number_of_comparisons
cor_table$p_adjust<-ifelse(cor_table$p_adjust>1,1,cor_table$p_adjust)

#produce table
cor_table_format<-cor_table %>%
    kbl(caption = "correlations of baseline pupil size (BPS), 
        stimulus-evoked pupillary response (SEPR), and
        mismatch negativity (MMN) for different stimuli.",
        col.names = c('correlation','coefficient (r)','p','p adjusted')) %>%
    kable_classic(full_width = F, html_font = "Cambria")

cor_table_format

save_kable(cor_table_format,file=paste0(project_path,'/output/supplements/associations_DV_perparticipant.html'))

detach(df_timepoint)
  
```

# significant associations between groups

```{r}

# BPS --> SEPR (standards)
df_timepoint$pd_res_z<-scale(df_timepoint$pd_res)
lm<-lm(scale(rpd_res_response.201)~pd_res_z*t1_diagnosis,df_timepoint)
summary(lm)
Anova(lm,type=3)
emtrends(lm, ~ t1_diagnosis,var='pd_res_z')
##--> specific to ASD

lm<-lm(scale(rpd_res_response.201)~pd_res_z*t1_diagnosis,df_timepoint)
Anova(lm,type=3)
emtrends(lm, ~ t1_diagnosis,var='pd_res_z')


# BPS --> SEPR (length)
df_timepoint$pd_res_z<-scale(df_timepoint$pd_res)
lm<-lm(scale(rpd_res_response.203)~pd_res_z*t1_diagnosis,df_timepoint)
Anova(lm,type=3)
##--> not influenced by group

# BPS --> MMN (standards)
df_timepoint$pd_res_z<-scale(df_timepoint$pd_res)
lm<-lm(scale(mmn.201)~pd_res_z*t1_diagnosis,df_timepoint)
Anova(lm,type=3)
##--> not influenced by group

# BPS --> MMN (standards)
df_timepoint$pd_res_z<-scale(df_timepoint$pd_res)
lm<-lm(scale(mmn.202)~pd_res_z*t1_diagnosis,df_timepoint)
Anova(lm,type=3)
##--> not influenced by group

# BPS --> MMN (standards)
df_timepoint$pd_res_z<-scale(df_timepoint$pd_res)
lm<-lm(scale(mmn.203)~pd_res_z*t1_diagnosis,df_timepoint)
Anova(lm,type=3)
##--> not influenced by group

# BPS --> MMN (standards)
df_timepoint$pd_res_z<-scale(df_timepoint$pd_res)
lm<-lm(scale(mmn.204)~pd_res_z*t1_diagnosis,df_timepoint)
Anova(lm,type=3)
##--> not influenced by group

# MMN --> SEPR (standards)
lm<-lm(scale(rpd_res_response.201)~scale(mmn.201)*t1_diagnosis,df_timepoint)
summary(lm)
Anova(lm,type=3)
contrast(emmeans(lm,  ~ t1_diagnosis),'pairwise')

lm<-lm(scale(mmn.201)~scale(rpd_res_response.201)*t1_diagnosis,df_timepoint)
Anova(lm,type=3)
confint(contrast(emmeans(lm,  ~ t1_diagnosis),'pairwise'))

```


# association of dependent variables (per trial)

```{r BPS_SEPR_association}

#interrelation of BPS and SEPR
lmm<-lmer(scale(rpd_res_response)~scale(pd_res)*EventData_epoch+
            (1|subjects),data=df_epoch)

anova(lmm)

cbind(round(fixef(lmm)['scale(pd_res)'],2),
              round(confint(lmm,parm = 'scale(pd_res)'),2))

#model fit
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(rpd_res_response)~scale(pd_baseline)+
              (1|subjects),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(rpd_res_response)~1+
                   (1|subjects),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  

```  

```{r BPS_SEPR_association_Bayesian, eval=eval_bayes}

#restrict to complete data
df_greta<-df_epoch[!is.na(df_epoch$pd_res),]
df_greta<-df_greta[!is.na(df_greta$rpd_res_response),]


 # data - create greta arrays
    x <- as_data(scale(df_greta$pd_res))
    y <- as_data(scale(df_greta$rpd_res_response))
    subject_id <- as.integer(factor(df_greta$subjects)) #random intercept (in LMM)

    # variables and priors
    int <- normal(0, 1)
    coef <- normal(0, 1)
    sd <- student(3, 0, 1, truncation = c(0, Inf))
    a_subject <- normal(0, 1, dim = max(subject_id))

    # operations
    mean <- int + coef * x + a_subject[subject_id]

    # likelihood - distribution over data
    distribution(y) <- normal(mean, sd)

    # defining the model
    m <- model(int, coef, sd)

    # plotting
    #require(DiagrammeR)
    plot(m)

    # sampling
    start <-Sys.time()
    draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups)
    Sys.time() - start
    
    summary(draws)
    
    mcmc_trace(draws) #see whether chains converge and are stationary
    mcmc_dens(draws) # + stable estimates
    
    #posterior predictive checking
    coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
    coda::effectiveSize(draws) #ESS

    #save model
    draws_SEPR_BPS<-draws
    save(draws_SEPR_BPS,file=paste0(project_path,'/data/Bayesian_model_SEPR_BPS.Rdata'))


```

- association of BPS and MMN

```{r BPS MMN association LMM}

###here use ID as random intercept (subject + wave) as MMN is only available for one timepoint

#interrelation of BPS and MMN
lmm<-lmer(scale(mmn)~scale(pd_res)*EventData_epoch+
            (1|id),data=df_epoch)

anova(lmm)

cbind(round(fixef(lmm)['scale(pd_res)'],2),
              round(confint(lmm,parm = 'scale(pd_res)'),2))

#model fit
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(mmn)~scale(pd_res)+EventData_epoch+
              (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(mmn)~EventData_epoch+
                   (1|id),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  



```

```{r SEPR MMN association LMM}

###here use ID as random intercept (subject + wave) as MMN is only available for one timepoint

#interrelation of BPS and SEPR
lmm<-lmer(scale(mmn)~scale(rpd_res_response)*EventData_epoch+
            (1|id),data=df_epoch)

anova(lmm)

cbind(round(fixef(lmm)['scale(rpd_res_response)'],2),
              round(confint(lmm,parm = 'scale(rpd_res_response)'),2))

#model fit
r2_nakagawa(lmm)  

#BIC Bayes Factor
lmm_ML<-lmer(scale(mmn)~scale(rpd_res_response)+
              (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(mmn)~1+
                   (1|id),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  


```

# figure - pupil size and Fz amplitude change within trials (figure 2)

visualization of pupile size change (BPS,SEPR) and Fz amplitude change (MMN) within trials

## MMN - scalp topography (part 1)

```{r scalp topography, warnings=F}

#load all EEG channel data - LARGE FILE ON DIFFERENT DRIVE
load(file=localation_of_EEG_allchannel_data)

#for standard coordinates of electrodes
data(eegcoord) #load sample set with standard coordinates
#rownames(eegcoord) #names of channels

###select only participants that are in final sample
subjects_in_list<-substr(names(list_mmn),10,22)
subjects_in_list<-sub('/','',subjects_in_list)
subjects_in_list<-sub('_','',subjects_in_list)
list_elements_in_final_sample<-which(subjects_in_list %in% unique(df_trial$subjects))
list_eeg_selected<-list_eeg[list_elements_in_final_sample]

#participants_to_sample<-round(length(list_eeg)/4*fraction_to_sample)
participants_to_sample<-round(length(list_eeg_selected)/4*fraction_to_sample)

#select 30 random from each condition - same participants
combined_sampled<-sample(1:(length(list_eeg_selected)/4),participants_to_sample)
duration_sampled<-combined_sampled+length(list_eeg_selected)/4
frequency_sampled<-combined_sampled+2*length(list_eeg_selected)/4
standard_sampled<-combined_sampled+3*length(list_eeg_selected)/4

select_samples<-c(combined_sampled,
                  duration_sampled,
                  frequency_sampled,
                  standard_sampled)

#select thirty random participants - from participants that have ET data
list_eeg_selected<-list_eeg_selected[select_samples]

fun_prepare_channeldata_plotting<-function(one_set,sample_coordinates){

  ##testing
  # one_set<-list_eeg_selected[[1]]
  # sample_coordinates<-eegcoord
  
  #match names of sample coordnisates and data set
  names(one_set)<-toupper(names(one_set)) #capitalize names - naming convention of eegkit
  sample_coordinates<-sample_coordinates[rownames(sample_coordinates) %in% names(one_set),]
  sample_coordinates$variable<-rownames(sample_coordinates)


  #limit site to interested timeframe - 150-250 ms after trial onset
  one_set<-one_set[one_set$TIMES >= 150 & one_set$TIMES <= 250,]
      ###--> less flexibel but reduces file size
  
  #melt to long format
  one_set$TIMES<-as.factor(one_set$TIMES) #preserve in melting
  one_set<-reshape2::melt(one_set,value.name='amplitude') #convert to long format

  #merge with sampel coordinates
  one_set<-merge(one_set,sample_coordinates,by='variable') #merge with sample coordniates
  names(one_set)[1]<-'electrode' #set name for later plotting function
  one_set$TIMES<-as.numeric(levels(one_set$TIMES))[one_set$TIMES]
  
  return(one_set)

      }

#takes around 20 minutes
list_eeg_selected<-pblapply(list_eeg_selected,fun_prepare_channeldata_plotting,sample_coordinates=eegcoord)
df_plot<-data.table::rbindlist(list_eeg_selected)

#scalp topography separated for condition
limits_of_plotted_range<-c(-3,3)

gf_4a<-topoplot(df_plot[df_plot$CONDITION=='standard',], 
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='standard') #scale coordniates to head

gf_4b<-topoplot(df_plot[df_plot$CONDITION=='frequenc',], 
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='pitch oddball') #scale coordniates to head

gf_4c<-topoplot(df_plot[df_plot$CONDITION=='duration',], 
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='length oddball') #scale coordniates to head

gf_4d<-topoplot(df_plot[df_plot$CONDITION=='combined',],
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='pitch & length oddball') #scale coordniates to head


#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend<-g_legend(gf_4a)

#plot in markdown
grid.arrange(
  arrangeGrob(
    gf_4a + theme(legend.position="none"),
    gf_4b + theme(legend.position="none"),
    gf_4c + theme(legend.position="none"),
    gf_4d + theme(legend.position="none"),nrow=2),
  mylegend, ncol=2, widths=c(5,1))


```

## MMN - Fz amplitude change within trials (part 2)

```{r Fz amplitude change}

df_eeg<-data.table::rbindlist(list_eeg_redux) ##very fast compared to do.call(rbind)

subsample<-sample(1:nrow(df_eeg),nrow(df_eeg)/sampling_factor) # select 10% of data for faster plotting


    ### use participants selected for scalp topography - much slower and same results
    # df_eeg$ID<-sub('/','',df_eeg$ID)
    # df_eeg$ID<-sub('_','',df_eeg$ID)
    # selected_participants<-substr(names(list_mmn)[combined_sampled],10,21)
    # subsample<-which(df_eeg$ID %in% selected_participants)


gf_1<-ggplot(df_eeg[subsample,],aes(x=times/1000,y=Fz,group=condition,color=condition,fill=condition))+
  geom_smooth()+
  labs(y='electrode Fz - amplitute (uV)',x='trial duration (s)')+
  scale_fill_manual(values = custom_condition_colors, labels=c("standard" = "standard", "frequenc" = "pitch oddball", "duration" = "length oddball ", "combined" = "pitch & length oddball"))+
  scale_color_manual(values = custom_condition_colors, labels=c("standard" = "standard", "frequenc" = "pitch oddball", "duration" = "length oddball ", "combined" = "pitch & length oddball"))+
  xlim(-0.1,0.6) #allign with eye-tracking epoch for plotting

gf_1

```

## pupil size change within trial - progression between groups (part 3+4)

- see preprocessing script
- composite figure is arranged in GIMP (image processing software)

# task effects

investigates influence of stimulus and task progression on pupillary measures (BPS, SEPR) and EEG measure (MMN)
in separate linear mixed models and supported by Bayesian posterior sampling

## task condition / stimulus effect

```{r stimulus_on_BPS}

##BPS
## include trial characteristics - TRIAL CONDITION
lmm<-lmer(scale(pd_res)~EventData_epoch+
            (1|id),data=df_epoch)

anova(lmm)
emmeans(lmm,~EventData_epoch)
confint(contrast(emmeans(lmm,~EventData_epoch),'pairwise'))
###--> 201 > 202

    
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(pd_res)~EventData_epoch+
              (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(pd_res)~1+
                  (1|id),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  
 
```

```{r stimulus_on_SEPR}

##SEPR
  
## include trial characteristics - TRIAL CONDITION
lmm<-lmer(scale(rpd_res_response)~EventData_epoch+
            (1|id),data=df_epoch)

anova(lmm)
emmeans(lmm,~EventData_epoch)
confint(contrast(emmeans(lmm,~EventData_epoch),'pairwise'))
###--> 201 > 202 + 203 + 204

r2_nakagawa(lmm)  
  
#BIC Bayes Factor  
lmm_ML<-lmer(scale(rpd_res_response)~EventData_epoch+
          (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(rpd_res_response)~1+
                 (1|id),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
BF_BIC
  
```

```{r stimulus_on_MMN}

## include trial characteristics - TRIAL CONDITION
lmm<-lmer(scale(mmn)~EventData_epoch+
            (1|id),data=df_epoch)

anova(lmm)
emmeans(lmm,~EventData_epoch)
confint(contrast(emmeans(lmm,~EventData_epoch),'pairwise'))

r2_nakagawa(lmm)  
  
#BIC Bayes Factor  
lmm_ML<-lmer(scale(mmn)~EventData_epoch+
          (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(mmn)~1+
                 (1|id),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
BF_BIC
  
```

## task progression effect

```{r task_progression_on_BPS}
  
## include trial characteristics - TASK PROGRESSION
  lmm<-lmer(scale(pd_res)~stats::poly(EventCounter_epoch,3)+scale(luminance_m)+
              (1|id),data=df_epoch)
  
  anova(lmm)
  
  confint(contrast(emmeans(lmm,~EventCounter_epoch,at=list(EventCounter_epoch =     seq(1,1400,699))),'revpairwise'))

r2_nakagawa(lmm)  
 
#effect of luminance
cbind(round(fixef(lmm)['scale(luminance_m)'],9),
                round(confint(lmm,parm = 'scale(luminance_m)'),9))

#BIC Bayes Factor
lmm_ML<-lmer(scale(pd_res)~stats::poly(EventCounter_epoch,3)+luminance_m+
              (1|subjects),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(pd_res)~luminance_m++
                   (1|subjects),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  


```

```{r task progression on SEPR}

## include trial characteristics - TASK PROGRESSION
  lmm<-lmer(scale(rpd_res_response)~EventCounter_epoch+luminance_m+
              (1|subjects),data=df_epoch)
  anova(lmm)
  
  cbind(round(fixef(lmm)['EventCounter'],9),
                round(confint(lmm,parm = 'EventCounter'),9))

  confint(contrast(emmeans(lmm,~EventCounter,at=list(EventCounter =   seq(1,1400,1399))),'pairwise'))
  
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(rpd_res_response)~EventCounter_epoch+
              (1|subjects),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(rpd_res_response)~1+
                   (1|subjects),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  


```

```{r task progression on MMN}

## include trial characteristics - TASK PROGRESSION
  lmm<-lmer(scale(mmn)~EventCounter_epoch+
              (1|id),data=df_epoch)
  anova(lmm)
  
  cbind(round(fixef(lmm)['EventCounter'],9),
                round(confint(lmm,parm = 'EventCounter'),9))

  confint(contrast(emmeans(lmm,~EventCounter,at=list(EventCounter =   seq(1,1400,1399))),'pairwise'))
  
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(mmn)~EventCounter_epoch+
              (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(mmn)~1+
                   (1|id),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  

```

## figure task effects (figure 3)

create a combined figure of stimulus and task progression effects on BPS, SEPR, MMN

```{r figure_task_effects}

#y-axis limitation for combined plot
y_axis_limitation<-c(-0.1, 0.18)

#BPS
lmm<-lmer(scale(pd_res)~EventData_epoch+
              (1|id),data=df_epoch)

#create plot --> of contrasts
plot_task_effect<-plot(contrast(emmeans(lmm,~EventData_epoch),'pairwise'))[['data']]

#modify plot
g1<-ggplot(plot_task_effect,aes(x=contrast,y=the.emmean))+
  #conventional error bar
  geom_errorbar(aes(min=asymp.LCL,max=asymp.UCL),width=0.2)+
  
  # #overplot predicted values
  # geom_jitter(data=df_plot_predicted_bps[complete.cases(df_plot_predicted_bps),],
  #             aes(x=predicted_bps_eventdata,
  #                 y=predicted_bps_mean,color=predicted_bps_eventdata),alpha=0.3,width=0.4,show.legend=F)+
  #boxplot
  geom_boxplot(aes(fill=contrast,
                   middle=the.emmean,
                   lower=the.emmean-1.5*SE,
                   upper=the.emmean+1.5*SE,
                   ymin=asymp.LCL,
                   ymax=asymp.UCL),stat = "identity",alpha=0.7)+
  scale_fill_manual(values = custom_contrast_colors,labels = c('standard - pitch','standard - length','standard - pitch & length','pitch - length','pitch - pitch & length','length - pitch & length'))+
  scale_x_discrete(labels = NULL, breaks = NULL)+ #remove x-axis tick labels
  coord_cartesian(ylim = y_axis_limitation)+
  labs(x='contrast of stimuli',y='BPS difference (z)')
  

#SEPR -pupillary response - technical model
lmm<-lmer(scale(rpd_res_response)~EventData_epoch+
              (1|id),data=df_epoch)

#create plot --> of contrasts
plot_task_effect<-plot(contrast(emmeans(lmm,~EventData_epoch),'pairwise'))[['data']]

#modify plot
g2<-ggplot(plot_task_effect,aes(x=contrast,y=the.emmean))+
  #conventional error bar
  geom_errorbar(aes(min=asymp.LCL,max=asymp.UCL),width=0.2)+
  
    #boxplot
  geom_boxplot(aes(fill=contrast,
                   middle=the.emmean,
                   lower=the.emmean-1.5*SE,
                   upper=the.emmean+1.5*SE,
                   ymin=asymp.LCL,
                   ymax=asymp.UCL),stat = "identity",alpha=0.7)+
  scale_fill_manual(values = custom_contrast_colors,labels = c('standard - pitch','standard - length','standard - pitch & length','pitch - length','pitch - pitch & length','length - pitch & length'))+
  scale_x_discrete(labels = NULL, breaks = NULL)+ #remove x-axis tick labels
  coord_cartesian(ylim = y_axis_limitation)+
  labs(x='contrast of stimuli',y='SEPR difference (z)')

##MMN
lmm<-lmer(scale(mmn)~EventData_epoch+(1|id),df_epoch)

      #create plot --> of contrasts
      plot_task_effect<-plot(contrast(emmeans(lmm,~EventData_epoch),'pairwise'))[['data']]

      #modify plot
      g3<-ggplot(plot_task_effect,aes(x=contrast,y=the.emmean))+
        #conventional error bar
        geom_errorbar(aes(min=asymp.LCL,max=asymp.UCL),width=0.2)+
        geom_boxplot(aes(fill=contrast,
                         middle=the.emmean,
                         lower=the.emmean-1.5*SE,
                         upper=the.emmean+1.5*SE,
                         ymin=asymp.LCL,
                         ymax=asymp.UCL),stat = "identity",alpha=0.7)+
        scale_fill_manual(values = custom_contrast_colors,labels = c('standard - pitch','standard - length','standard - pitch & length','pitch - length','pitch - pitch & length','length - pitch & length'))+
        scale_x_discrete(labels = NULL, breaks = NULL)+ #remove x-axis tick labels
        coord_cartesian(ylim = y_axis_limitation)+
        labs(x='contrast of stimuli',y='MMN difference (z)')


###task progression BPS
BPS_progression<-lmer(scale(pd_res)~stats::poly(EventCounter_epoch,3)+
              (1|id),data=df_epoch)
      
model_plot_data_BPS<-as.data.frame(emmeans(BPS_progression,~EventCounter_epoch,at=list(EventCounter_epoch = seq(1,1400,140))))
    
g4<-ggplot(model_plot_data_BPS,aes(x=EventCounter_epoch,y=emmean))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE),alpha=0.8,position='dodge',fill='grey35')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  coord_cartesian(ylim = c(-0.2,0.25))+
  labs(x='task progression (trial)',y='baseline pupil size (BPS, z)')

###task progression SEPR
SEPR_progression<-lmer(scale(rpd_res_response)~EventCounter_epoch+
              (1|id),data=df_epoch)
      
      
model_plot_data_SEPR<-as.data.frame(emmeans(SEPR_progression,~EventCounter_epoch,at=list(EventCounter_epoch = seq(1,1400,140))))
    
g5<-ggplot(model_plot_data_SEPR,aes(x=EventCounter_epoch,y=emmean))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE),alpha=0.8,position='dodge',fill='grey35')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  coord_cartesian(ylim = c(-0.2,0.25))+
  labs(x='task progression (trial)',y='pupillary response (SEPR, z)')

###task progression MMN
MMN_progression<-lmer(scale(mmn)~EventCounter_epoch+
              (1|id),data=df_epoch)
      
model_plot_data_MMN<-as.data.frame(emmeans(MMN_progression,~EventCounter_epoch,at=list(EventCounter_epoch = seq(1,1400,140))))
    
g6<-ggplot(model_plot_data_MMN,aes(x=EventCounter_epoch,y=emmean))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE),alpha=0.8,position='dodge',fill='grey35')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  coord_cartesian(ylim = c(-0.2,0.25))+
  labs(x='task progression (trial)',y='mismatch negativity (MMN, z)')

#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend_contrast<-g_legend(g3)

      
###composite figure layout
lay <- rbind(c(1,2,3,7),
             c(4,5,6,7))


#plot in markdown
grid.arrange(g1+theme(legend.position="none")+xlab(''),
             g2+theme(legend.position="none"),
             g3+theme(legend.position="none")+xlab(''),
             g4+xlab(''),
             g5,
             g6+xlab(''),
             mylegend_contrast,layout_matrix=lay)

#save to file
tiff(file=paste0(project_path,"/output/figures/task_effects_contrasts_progression.tiff"), # create a file in tiff format in 
  width=8, height=5, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(g1+theme(legend.position="none")+xlab(''),
             g2+theme(legend.position="none"),
             g3+theme(legend.position="none")+xlab(''),
             g4+xlab(''),
             g5,
             g6+xlab(''),
             mylegend_contrast,layout_matrix=lay)
      

dev.off() #close operation and save file


```

# group differences

## baseline pupil size (BPS)

```{r BPS task progression between groups - LMM}

#main model
lmm_REML<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+scale(luminance_m)+
              (1|id),data=df_epoch)

#fixed effect significances
anova(lmm_REML)

#model fit
r2_nakagawa(lmm_REML)

# #POST-HOC
# contrast(emmeans(lmm_REML,~EventData_epoch|t1_diagnosis),'pairwise')
# ##--> lower BPS in ASD for length versus standards

#Post-hoc - between group
emtrends(lmm_REML,~t1_diagnosis,var='EventCounter_epoch')

#emmeans(lmm_REML,~t1_diagnosis|EventCounter,at=list(EventCounter = seq(1,1400,1399)))
emmeans(lmm_REML,~t1_diagnosis|EventCounter_epoch,at=list(EventCounter_epoch = seq(1,1400,699)))


  confint(contrast(emmeans(lmm_REML,~EventCounter_epoch|t1_diagnosis,at=list(EventCounter_epoch =   seq(1,1400,699))),'revpairwise'))


#Bayes factor - group x task progression
lmm_ML<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+
              (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(pd_res)~EventData_epoch+t1_diagnosis+stats::poly(EventCounter_epoch,3)+
                   EventData_epoch:EventCounter_epoch+EventData_epoch:t1_diagnosis+
              (1|id),data=df_epoch,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
BF_BIC

### compare full model for influence of covariates

#full model
lmm_full<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|id),data=df_epoch)
              #--> group effect (interaction) is robust in reduced and full model

anova(lmm_full)

  
```

```{r BPS task progression between groups - Bayesian, eval=eval_bayes}

##define parameters of posterior sampling in Bayesian modeling
number_of_warmups <- 4000
number_of_iterations <- 8000
credible_interval<-0.89 #credible interval Bayesian estimates

df_greta<-df_epoch[!is.na(df_epoch$pd_res),]

# A. translate data - make greta arrays
y <- as_data(scale(df_greta$pd_res))
design <- as.matrix(cbind(ifelse(df_greta$t1_diagnosis=='ASD',T,F),df_greta$EventCounter_epoch))
subject_id <- as.integer(factor(df_greta$id)) #random intercept (in LMM)

#define priors
int <- normal(0, 1)
design_coef<- normal(0, 1, dim=ncol(design))
sd <- cauchy(0, 3, truncation = c(0, Inf))
#sd <- student(3, 0, 1, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- int + design %*% design_coef + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
m <- model(design_coef)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups,) ###--> test
Sys.time() - start

#save model
draws_BPS_twoway<-draws
save(draws_BPS_twoway,file=paste0(project_path,'/data/Bayesian_model_BPS_twoway.Rdata'))

#posterior estimates
point_estimate(draws_BPS_twoway)
ci(draws_BPS_twoway,ci=credible_interval)

```

## stimulus-evoked pupillary response

```{r SEPR task progression between groups LMM}

#MAIN MODEL - pupillary response by group - random intercept
lmm_REML<-lmer(scale(rpd_res_response)~EventData_epoch*t1_diagnosis*EventCounter_epoch+scale(luminance_m)+
            (1|id),data=df_epoch)
            
#fixed effect significances
anova(lmm_REML)

#coefficient of determination
r2_nakagawa(lmm_REML)

#Bayes factor - group x task progression
lmm_ML<-lmer(scale(rpd_res_response)~EventData_epoch*t1_diagnosis*EventCounter_epoch+
                (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(rpd_res_response)~EventData_epoch+EventCounter_epoch+
                 EventData_epoch:EventCounter_epoch+EventData_epoch:t1_diagnosis:EventCounter_epoch+
                (1|id),data=df_epoch,REML=F)

  BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
  BF_BIC

```

```{r SEPR_three_way_interaction Bayesian, eval=eval_bayes}

# A. translate data - make greta arrays
y <- as_data(scale(df_trial$rpd_auc))
trials <- as_data(scale(df_trial$EventCounter))
subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)
events <- as_data(model.matrix(~ EventData - 1, df_trial)) #alternative to dummy coding?
interaction_variable<-with(df_trial,interaction(t1_diagnosis,EventData))
group_by_events <- as_data(model.matrix(~interaction_variable-1, df_trial)) #alternative to dummy coding - remove intercept (-1)
#table(group_by_events[,1],group_by_events[,9])

#define priors
event_coefs<- normal(0, 1, dim=ncol(events))
group_event_coefs <- normal(0,1, dim=ncol(group_by_events))

sd <- cauchy(0, 3, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- trials * group_by_events %*% group_event_coefs + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
#m <- model(int, coef_201, coef_202, coef_203, coef_204, sd)
m <- model(group_event_coefs)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups) ###--> test
Sys.time() - start

#save model
draws_SEPR_threeway<-draws
save(draws_SEPR_threeway,file=paste0(project_path,'/data/Bayesian_model_SEPR_threewayinteraction.Rdata'))

#posterior estimates
table(interaction_variable) #names of coefficients
point_estimate(draws_SEPR_threeway)
ci(draws_SEPR_threeway,ci=credible_interval)


```

## mismatch negativity between groups

```{r MMN task progression - LLM}

#MAIN MODEL - MMN by group - random intercept
lmm_REML<-lmer(scale(mmn)~EventData_epoch*t1_diagnosis*EventCounter_epoch+
            (1|id),data=df_epoch)

#fixed effect significances
anova(lmm_REML)

#coefficient of determination
r2_nakagawa(lmm_REML)

#Bayes factor - group x task progression
lmm_ML<-lmer(scale(mmn)~EventData_epoch*t1_diagnosis*EventCounter_epoch+
                (1|id),data=df_epoch,REML=F)

null_lmm_ML <- lmer(scale(mmn)~EventData_epoch+EventCounter_epoch+
                 EventData_epoch:EventCounter_epoch+EventData_epoch:t1_diagnosis+
                (1|id),data=df_epoch,REML=F)

  BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
  BF_BIC


#full model
lmm_full<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|id),data=df_trial)
              #--> group effect (interaction) is robust in reduced and full model

anova(lmm_full)

#coefficient of determination
r2_nakagawa(lmm_full)


confint(contrast(emtrends(lmm_REML,~t1_diagnosis|EventData,var='EventCounter'),'pairwise'))

      plot(emtrends(lmm_REML,~t1_diagnosis|EventData,var='EventCounter'))

      
##translate to changes across the task
confint(contrast(emmeans(lmm_REML,~EventCounter|t1_diagnosis+EventData,at=list(EventCounter =   seq(1,1400,1399))),'revpairwise'))



```

```{r MMN task progression - Bayesian, eval=eval_bayes}

#remove NA
df_bayesian_mmn<-df_trial[!is.na(df_trial$mmn),]

# A. translate data - make greta arrays
y <- as_data(scale(df_bayesian_mmn$mmn))
trials <- as_data(scale(df_bayesian_mmn$EventCounter))
subject_id <- as.integer(factor(df_bayesian_mmn$subjects)) #random intercept (in LMM)
events <- as_data(model.matrix(~ EventData - 1, df_bayesian_mmn)) #alternative to dummy coding?
interaction_variable<-with(df_bayesian_mmn,interaction(t1_diagnosis,EventData))
group_by_events <- as_data(model.matrix(~interaction_variable-1, df_bayesian_mmn)) #alternative to dummy coding - remove intercept (-1)
#table(group_by_events[,1],group_by_events[,9])

#define priors
event_coefs<- normal(0, 1, dim=ncol(events))
group_event_coefs <- normal(0,1, dim=ncol(group_by_events))

sd <- cauchy(0, 3, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- trials * group_by_events %*% group_event_coefs + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
#m <- model(int, coef_201, coef_202, coef_203, coef_204, sd)
m <- model(group_event_coefs)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups) ###--> test
Sys.time() - start


#estimator convergence
mcmc_trace(draws) #see whether chains converge and are stationary
mcmc_dens(draws) # + stable estimates

##results
mcmc_areas(draws, prob= .66, prob_outer = .95)+scale_y_discrete(labels=c('ASD-standard','TD-standard ','ASD-pitch','TD-pitch','ASD-length ','TD-length','ASD-pitch+length','TD-pitch+length'))+coord_flip()+theme_bw()

#analyse
summary(draws)


#posterior predictive checking
coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
coda::effectiveSize(draws) #ESS - the higher the better

#save model
draws_MMN_threeway<-draws
save(draws_MMN_threeway,file=paste0(project_path,'/data/Bayesian_model_MMN_threewayinteraction.Rdata'))

#posterior estimate
table(interaction_variable) #names of coefficients
point_estimate(draws_MMN_threeway)
ci(draws_MMN_threeway,ci=credible_interval)


```

## figure group differences LMM + Bayesian

added to a compostie figure - not evaluated

```{r figure_BPS_SEP_MMN_between_groups, eval=FALSE}
#main models
BPS_progression<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+
              (1|id),data=df_epoch)

#plot interaction
model_plot_data_BPS1<-as.data.frame(emmeans(BPS_progression,~t1_diagnosis+EventCounter_epoch,
                                       at=list(EventCounter_epoch = seq(1,1400,140))))

g1<-ggplot(model_plot_data_BPS1,aes(x=EventCounter_epoch,y=emmean,group=t1_diagnosis,color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  labs(x='task progression (trial)',y='marginalized mean (z)')+
  theme(legend.position = 'none')


#main models - Bayesian --> load
load(file=paste0(project_path,'/data/Bayesian_model_BPS_twoway.Rdata')) #draws_BPS_twoway

#draws_BPS_twoway
plot_data<-mcmc_areas_data(draws_BPS_twoway,prob_outer = credible_interval)

g2<-ggplot(plot_data,aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_boxplot()+
  stat_boxplot(geom = "errorbar",width = 0.15) +
  scale_x_discrete(labels=c('autistic','non-autistic'))+
  labs(y='posterior sampling (z)',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = c(custom_group_color[1],custom_group_color[2]))

#plot in markdown
grid.arrange(g1,g2)


#save to file
tiff(file=paste0(project_path,"/output/figures/group_differences.tiff"), # create a file in tiff format in current working directory
  width=3, height=6, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure
      
grid.arrange(g1,g2)

dev.off() #close operation and save file


```

# supplementary analyses
## descriptives per participant 
### descriptive statistics of dependent variables - per participant

```{r descriptive statistics of dependent variables - per-participant, warnings=F}

descriptive_table_data<-
  
  rbind(
    psych::describe(df_timepoint$pd_res)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_res_response.201)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_res_response.202)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_res_response.203)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_res_response.204)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.201)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.202)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.203)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.204)[c('n','mean','sd','min','max','skew','kurtosis')]
    )


descriptive_table_data<-round(descriptive_table_data,3)

  variable<-c('BPS',
    'SEPR - standard','SEPR - pitch oddball',
    'SEPR - length oddball','SEPR - pitch+length oddball',
    'MMN - standard','MMN - pitch oddball',
    'MMN - length oddball','MMN - pitch+length oddball')

descriptive_table_data<-cbind(variable,descriptive_table_data)

#produce table
descriptive_table_format<-descriptive_table_data %>%
    kbl(caption = "Per-participant level: Descriptive statistics of baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), and mismatch negativity (MMN)",row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

descriptive_table_format

save_kable(descriptive_table_format,file=paste0(project_path,'/output/supplements/descriptivestatistics_DV_perparticipant.html'))


```

### histograms of dependent variables - per participant

```{r historgram of dependent variables - per participant}

grid.arrange(
  ggplot(df_timepoint,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot(df_timepoint,aes(x=gain_pcdm))+geom_histogram(bins=30)+labs(x='',title='neural gain (z)'),
  ggplot(df_timepoint,aes(x=rpd_auc.201))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.202))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.203))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.204))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_timepoint,aes(x=mmn.201))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_timepoint,aes(x=mmn.202))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.203))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.204))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

#save to file
tiff(file=paste0(project_path,"/output/supplements/histograms_dependentvariables_perparticipant.tiff"), # create a file in tiff format in current working directory
  width=8, height=12, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(
  ggplot(df_timepoint,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot(df_timepoint,aes(x=gain_pcdm))+geom_histogram(bins=30)+labs(x='',title='neural gain (z)'),
  ggplot(df_timepoint,aes(x=rpd_auc.201))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.202))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.203))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.204))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_timepoint,aes(x=mmn.201))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_timepoint,aes(x=mmn.202))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.203))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.204))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

dev.off() #close operation and save file


```






## descriptives per trial
### descriptive statistics of dependent variables - per trial

```{r descriptive statistics of dependent variables - per-trial, warnings=F}

descriptive_table_data<-
  
rbind(
psych::describe(df_epoch$pd_res)[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$rpd_res_response[df_epoch$EventData=='201'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$rpd_res_response[df_epoch$EventData=='202'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$rpd_res_response[df_epoch$EventData=='203'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$rpd_res_response[df_epoch$EventData=='204'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$mmn[df_epoch$EventData=='201'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$mmn[df_epoch$EventData=='202'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$mmn[df_epoch$EventData=='203'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_epoch$mmn[df_epoch$EventData=='204'])[c('n','mean','sd','min','max','skew','kurtosis')]

)


descriptive_table_data<-round(descriptive_table_data,3)

  variable<-c('BPS',
    'SEPR - standard','SEPR - pitch oddball',
    'SEPR - length oddball','SEPR - pitch+length oddball',
    'MMN - standard','MMN - pitch oddball',
    'MMN - length oddball','MMN - pitch+length oddball')

descriptive_table_data<-cbind(variable,descriptive_table_data)

#produce table
descriptive_table_format<-descriptive_table_data %>%
    kbl(caption = "Per-trial level: Descriptive statistics of baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), and mismatch negativity (MMN)",row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

descriptive_table_format

save_kable(descriptive_table_format,file=paste0(project_path,'/output/supplements/descriptivestatistics_DV_pertrial.html'))

```



### histograms of dependent variables - per trial

```{r historgram of dependent variables - per trial}

grid.arrange(
  ggplot(df_trial,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot()+theme_void(),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

#save to file
tiff(file=paste0(project_path,"/output/supplements/histograms_dependentvariables_pertrial.tiff"), # create a file in tiff format in current working directory
  width=8, height=12, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(
  ggplot(df_trial,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot()+theme_void(),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

dev.off() #close operation and save file


```

## group comparisons per-participant

```{r group comparisons per participant - table}
### table

BPS_group<-c(summary(lm<-lm(scale(pd_res)~t1_diagnosis,df_timepoint))[['fstatistic']],
  summary(lm<-lm(scale(pd_res)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])

SEPR_201_group<-c(summary(lm<-lm(scale(rpd_res_response.201)~t1_diagnosis,df_timepoint))[['fstatistic']],
             summary(lm<-lm(scale(rpd_res_response.201)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
SEPR_202_group<-c(summary(lm<-lm(scale(rpd_res_response.202)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(rpd_res_response.202)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
SEPR_203_group<-c(summary(lm<-lm(scale(rpd_res_response.203)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(rpd_res_response.203)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
SEPR_204_group<-c(summary(lm<-lm(scale(rpd_res_response.204)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(rpd_res_response.204)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])

MMN_201_group<-c(summary(lm<-lm(scale(mmn.201)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.201)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
MMN_202_group<-c(summary(lm<-lm(scale(mmn.202)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.202)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
MMN_203_group<-c(summary(lm<-lm(scale(mmn.203)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.203)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
MMN_204_group<-c(summary(lm<-lm(scale(mmn.204)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.204)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])

group_comparison_table<-round(rbind(BPS_group,
      SEPR_201_group,SEPR_202_group,SEPR_203_group,SEPR_204_group,
      MMN_201_group,MMN_202_group,MMN_203_group,MMN_204_group),3)


row_names<-c('BPS - across stimuli',
            'SEPR - standards','SEPR - pitch oddball','SEPR - length oddball','SEPR - pitch+length oddball',
            'MMN - standards','MMN - pitch oddball','MMN - length oddball','MMN - pitch+length oddball')

group_comparison_table<-cbind(row_names,group_comparison_table)

#produce table
group_comparison_table_format<-group_comparison_table %>%
  kbl(caption = "Per-participant level: Group comparisons of baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), and mismatch negativity (MMN)",
      row.names = F, col.names = c('','F','df1','df2','b','std. error','p')) %>%
  kable_classic(full_width = F, html_font = "Cambria")

group_comparison_table_format

save_kable(group_comparison_table_format,file=paste0(project_path,'/output/supplements/table_groupcomparison_perparticipant.html'))

```

## covariate effects on group comparisons per-partiicpant

```{r covariate effects on group comparisons per participant}

# significant covariate models ####
#full model - PD
lm<-lm(scale(pd_res)~t1_diagnosis+
         scale(ageyrs)+scale(t1_piq)+sex+
         as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)
       ,df_timepoint)

  ##supplements - full table
  row_names<-c('intercept','group',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'residuals')


  table_data_model<-round(Anova(lm,type=3),3)
  table_data_model<-cbind(row_names,table_data_model)

  table_data_model

  table_model<-table_data_model %>%
    kbl(caption = "Per participant: Full Linear model of baseline pupil size (BPS) with all potential covariates",
        col.names = c('','Sum Sq','df','F','p'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

  table_model

  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_linearmodel_BPS_allcovariates.html'))

#significant covariate model - PD
lm<-lm(scale(pd_res)~t1_diagnosis+
         scale(ageyrs)+scale(center_dev)
       ,df_timepoint)

  ##supplements - full table
  row_names<-c('intercept','group',
               'age','gaze center deviation','residuals')


  table_data_model<-round(Anova(lm,type=3),3)
  table_data_model<-cbind(row_names,table_data_model)

  table_data_model

  table_model<-table_data_model %>%
    kbl(caption = "Per participant: Reduced Linear model of baseline pupil size (BPS) with significant covariates",
        col.names = c('','Sum Sq','df','F','p'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

  table_model

  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_linearmodel_BPS_signcovariates.html'))

```


## medication effect on main group comparsions

```{r medication effect on group differences}

#describe medication usage
table(df_dem$med_use,df_dem$t1_diagnosis) #between groups
chisq.test(df_dem$med_use,df_dem$t1_diagnosis)

table(is.na(df_dem$t1_drugclass_2),df_dem$t1_diagnosis) # participants receiving more than one medication

round(table(c(df_dem$t1_drugclass_1,df_dem$t1_drugclass_2,df_dem$t1_drugclass_3))/table(df_dem$med_use)[2],2) #percentage of type of medication

#BPS
lm<-lm(scale(pd_res)~t1_diagnosis,df_timepoint) #model as in results
lm_med<-lm(scale(pd_res)~t1_diagnosis+med_use,df_timepoint) #model with medication as covariate
anova(lm,lm_med)
anova(lm_med)


#BPS - dynamic main model
lmm<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+scale(luminance_m)+
              (1|id),data=df_epoch,REML=F)
lmm_med<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+scale(luminance_m)+
                med_use+
              (1|id),data=df_epoch,REML=F)
anova(lmm,lmm_med)
anova(lmm_med)

fixef(lmm_med)

#SEPR - pupillary response by group - random intercept
lmm_med<-lmer(scale(rpd_res_response)~EventData_epoch*t1_diagnosis*EventCounter_epoch+
                med_use+
            (1|id),data=df_epoch,REML=F)
anova(lmm_med)

#MMN - MMN by group - random intercept
lmm_med<-lmer(scale(mmn)~EventData_epoch*t1_diagnosis*EventCounter_epoch+
                med_use+
            (1|id),data=df_epoch,REML=F)
anova(lmm_med)

```


## scaling of BPS SEPR and MMN for demographic variables between groups

### SEPR scaling with demographic variables

```{r SEPR scaling with demographics between groups}

  ###SCALING of pupillary response AUC WITH MAIN DEMOGRAPHIC VARIABLES ####

  hist(df_epoch$ageyrs) ## pick 10 - 25 years
  hist(df_epoch$t1_piq) ### IQ 80-135

    labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch+length oddball")

  #age
  g1<-ggplot(df_epoch[df_epoch$ageyrs>=10 & df_epoch$ageyrs<=25,],aes(x=ageyrs,y=scale(rpd_res_response),group=interaction(EventData_epoch,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='age (years)',y='')+
    coord_cartesian(ylim = c(-0.05, 0.15))+
    theme_bw()

  #IQ
  g2<-ggplot(df_epoch[df_epoch$t1_piq > 80 & df_epoch$t1_piq < 135,],aes(x=t1_piq,y=scale(rpd_res_response),group=interaction(EventData_epoch,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='perceptual IQ',y='')+
    coord_cartesian(ylim = c(-0.05, 0.15))+
    theme_bw()


    #summary functions
    stat_sum_df <- function(fun, geom="crossbar", ...) {
      stat_summary(fun.data=fun, geom=geom, width=0.2, ...)
    }

    stat_sum_single <- function(fun, geom="point", ...) {
      stat_summary(fun=fun, geom=geom, size = 3, ...)
    }

  group_labels<-c("female","female","male","male")
  g3<-ggplot(df_epoch,aes(x=interaction(t1_diagnosis,sex),y=scale(rpd_res_response),group=interaction(EventData_epoch,t1_diagnosis,sex),color=t1_diagnosis))+
    stat_sum_df("mean_cl_normal", geom = "errorbar")+
    stat_sum_single("mean")+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_x_discrete(labels=group_labels)+
    labs(x='biological sex',y='')+
    coord_cartesian(ylim = c(-0.05, 0.15))+
    theme_bw()

  
  grid.arrange(g1,g2,g3,nrow=3)
  
  
  #save to file
tiff(file=paste0(project_path,"/output/supplements/SEPR_scaling_withdemographics_between_groups.tiff"), # create a file in tiff format in current working directory
  width=10, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

  grid.arrange(g1,g2,g3,nrow=3,left='stimulus-evoked pupillary response (SEPR, z)')

        
dev.off() #close operation and save file

```

### BPS scaling with demographic variables

```{r BPS scaling with demographics between groups}
  ###SCALING of baseline pupil size WITH MAIN DEMOGRAPHIC VARIABLES ####

  labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch+length oddball")
  #age
  g1<-ggplot(df_epoch[df_epoch$ageyrs>=10 & df_epoch$ageyrs<=25,],aes(x=ageyrs,y=pd_res,group=interaction(EventData_epoch,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='age (years)',y='')+
    theme_bw()

  #IQ
  g2<-ggplot(df_epoch[df_epoch$t1_piq > 80 & df_epoch$t1_piq < 135,],aes(x=t1_piq,y=pd_res,group=interaction(EventData_epoch,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='perceptual IQ',y='')+
    theme_bw()


  #summary functions
  stat_sum_df <- function(fun, geom="crossbar", ...) {
    stat_summary(fun.data=fun, geom=geom, width=0.2, ...)
  }

  stat_sum_single <- function(fun, geom="point", ...) {
    stat_summary(fun=fun, geom=geom, size = 3, ...)
  }

  group_labels<-c("female","female","male","male")
  g3<-ggplot(df_epoch,aes(x=interaction(t1_diagnosis,sex),y=pd_res,group=interaction(EventData_epoch,t1_diagnosis,sex),color=t1_diagnosis))+
    stat_sum_df("mean_cl_normal", geom = "errorbar")+
    stat_sum_single("mean")+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_x_discrete(labels=group_labels)+
    labs(x='biological sex',y='')+
    theme_bw()

  grid.arrange(g1,g2,g3,nrow=3)
  ####--> higher IQ and lower age in autistic females?

      with(df_dem,by(t1_piq,interaction(t1_sex,t1_diagnosis),summary))
      with(df_dem,by(t1_ageyrs,interaction(t1_sex,t1_diagnosis),summary))

  
    #save to file
tiff(file=paste0(project_path,"/output/supplements/BPS_scaling_withdemographics_between_groups.tiff"), # create a file in tiff format in current working directory
  width=10, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

  grid.arrange(g1,g2,g3,nrow=3,left='baseline pupil size (mm)')

        
dev.off() #close operation and save file
    
      

```

### MMN scaling with demographic variables

```{r MMN scaling with demographics between groups}
  ###SCALING of MMN WITH MAIN DEMOGRAPHIC VARIABLES ####

    labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch+length oddball")

  #age
  g1<-ggplot(df_epoch[df_epoch$ageyrs>=10 & df_epoch$ageyrs<=25,],aes(x=ageyrs,y=mmn,group=interaction(EventData_epoch,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='age (years)',y='')+
    theme_bw()

  #IQ
  g2<-ggplot(df_epoch[df_epoch$t1_piq > 80 & df_epoch$t1_piq < 135,],aes(x=t1_piq,y=mmn,group=interaction(EventData_epoch,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='perceptual IQ',y='')+
    theme_bw()


  #summary functions
  stat_sum_df <- function(fun, geom="crossbar", ...) {
    stat_summary(fun.data=fun, geom=geom, width=0.2, ...)
  }

  stat_sum_single <- function(fun, geom="point", ...) {
    stat_summary(fun=fun, geom=geom, size = 3, ...)
  }

  group_labels<-c("female","female","male","male")
  g3<-ggplot(df_epoch,aes(x=interaction(t1_diagnosis,sex),y=mmn,group=interaction(EventData_epoch,t1_diagnosis,sex),color=t1_diagnosis))+
    stat_sum_df("mean_cl_normal", geom = "errorbar")+
    stat_sum_single("mean")+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    facet_grid(~EventData_epoch,labeller=labeller(EventData_epoch = labels))+
    scale_x_discrete(labels=group_labels)+
    labs(x='biological sex',y='')+
    theme_bw()

  grid.arrange(g1,g2,g3,nrow=3)
  ####--> higher IQ and lower age in autistic females?

  with(df_dem,by(t1_piq,interaction(t1_sex,t1_diagnosis),summary))
  with(df_dem,by(t1_ageyrs,interaction(t1_sex,t1_diagnosis),summary))


  #save to file
tiff(file=paste0(project_path,"/output/supplements/MMN_scaling_withdemographics_between_groups.tiff"), # create a file in tiff format in current working directory
  width=10, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

  grid.arrange(g1,g2,g3,nrow=3,left='mismatch negativity (uV)')

        
dev.off() #close operation and save file





```


## linearity of effect of trial number (polynomial fits)

```{r linearity BPS}

###polynomial fit of trial counter
  linear_fit<-lmer(scale(pd_res)~scale(EventCounter_epoch)+
          (1|id),data=df_epoch,REML=F)

  quadratic_fit<-lmer(scale(pd_res)~stats::poly(scale(EventCounter_epoch),2)+
              (1|id),data=df_epoch,REML=F)

  cubic_fit<-lmer(scale(pd_res)~stats::poly(scale(EventCounter_epoch),3)+
               (1|id),data=df_epoch,REML=F)
  
  anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-cbind(c('linear fit','quadratic fit','cubic fit'),table_model_compare)
  
  table_formatted<-table_model_compare %>%
  kbl(caption = "Model comparison on the linearity of trial number on baseline pupil size (BPS)",
      col.names = c('','number of parameters','AIC','BIC','log likelihood','deviance','Chi-squared','df','p-value'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_formatted

save_kable(table_formatted, file=paste0(project_path,'/output/supplements/table_modelcomparison_linearity_BPS.html'))

```

```{r linearity SEPR}

###polynomial fit of trial counter
  linear_fit<-lmer(scale(rpd_res_response)~scale(EventCounter_epoch)+
          (1|id),data=df_epoch,REML=F)

  quadratic_fit<-lmer(scale(rpd_res_response)~stats::poly(scale(EventCounter_epoch),2)+
              (1|id),data=df_epoch,REML=F)

  cubic_fit<-lmer(scale(rpd_res_response)~stats::poly(scale(EventCounter_epoch),3)+
               (1|id),data=df_epoch,REML=F)
  
  anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-cbind(c('linear fit','quadratic fit','cubic fit'),table_model_compare)
  
  table_formatted<-table_model_compare %>%
  kbl(caption = "Model comparison on the linearity of trial number on stimulus-evoked pupillary response (SEPR)",
      col.names = c('','number of parameters','AIC','BIC','log likelihood','deviance','Chi-squared','df','p-value'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_formatted

save_kable(table_formatted, file=paste0(project_path,'/output/supplements/table_modelcomparison_linearity_SEPR.html'))
  
  
```

```{r linearity MMN}

###polynomial fit of trial counter
  linear_fit<-lmer(scale(mmn)~scale(EventCounter_epoch)+
          (1|id),data=df_epoch,REML=F)

  quadratic_fit<-lmer(scale(mmn)~stats::poly(scale(EventCounter_epoch),2)+
              (1|id),data=df_epoch,REML=F)

  cubic_fit<-lmer(scale(mmn)~stats::poly(scale(EventCounter_epoch),3)+
               (1|id),data=df_epoch,REML=F)
  
  anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-cbind(c('linear fit','quadratic fit','cubic fit'),table_model_compare)
  
  table_formatted<-table_model_compare %>%
  kbl(caption = "Model comparison on the linearity of trial number on mismatch negativity (MMN)",
      col.names = c('','number of parameters','AIC','BIC','log likelihood','deviance','Chi-squared','df','p-value'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_formatted

save_kable(table_formatted, file=paste0(project_path,'/output/supplements/table_modelcomparison_linearity_MMN.html'))

```

## full model tables

```{r BPS full model LMM}

#full model  - save to supplements      
lmm_full<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+
      scale(luminance_m)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|id),data=df_epoch)

  plot(contrast(emmeans(lmm_full, ~ EventData_epoch| t1_diagnosis),'pairwise'))

##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'video luminance',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'S x group','S x TP','group x TP','S x group x TP')
  
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: baseline pupil size (BPS) - task progression",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_BPS.html'))

```

## Bayesian chains - convergence and stationarity of estimates

```{r figure Bayesian models chain convergence stationarity}

###load saved Bayeisan posterior estimates
load(file=paste0(project_path,'/data/Bayesian_model_BPS_twoway.Rdata')) #draws_BPS_twoway
load(file=paste0(project_path,'/data/Bayesian_model_BPS_supp.Rdata')) #draws_BPS_supp
load(file=paste0(project_path,'/data/Bayesian_model_SEPR_threewayinteraction.Rdata')) #draws_SEPR_threeway
load(file=paste0(project_path,'/data/Bayesian_model_MMN_threewayinteraction.Rdata')) #draws_MMN_threeway

### for supplements

#chain converge and stationary

# #draws_BPS_twoway - chain converge and stationary
var_names <- c("design_coef[1,1]" = "BPS_task_progression: autistic",
               "design_coef[2,1]" = "BPS_task_progression: non-autistic") #define parameter names

g1a<-mcmc_trace(draws_BPS_twoway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed))))
g1b<-mcmc_dens(draws_BPS_twoway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)))) 

# #draws_BPS_supp - chain converge and stationary
var_names <- c("design_coef[1,1]" = "BPS_sequence_position: autistic",
               "design_coef[2,1]" = "BPS_sequence_position: non-autistic") #define parameter names

g2a<-mcmc_trace(draws_BPS_supp, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed))))
g2b<-mcmc_dens(draws_BPS_supp, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)))) 

# #draws_SEPR_threeway - chain converge and stationary
var_names <- c("group_event_coefs[1,1]" = "SEPR_trend: autistic_standard",
               "group_event_coefs[2,1]" = "SEPR_trend: non-autistic_standard",
               "group_event_coefs[3,1]" = "SEPR_trend: autistic_pitch",
               "group_event_coefs[4,1]" = "SEPR_trend: non-autistic_pitch",
               "group_event_coefs[5,1]" = "SEPR_trend: autistic_length",
               "group_event_coefs[6,1]" = "SEPR_trend: non-autistic_length",
               "group_event_coefs[7,1]" = "SEPR_trend: autistic_pitch+length",
               "group_event_coefs[8,1]" = "SEPR_trend: non-autistic_pitch+length") #define parameter names

g3a<-mcmc_trace(draws_SEPR_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1))
g3b<-mcmc_dens(draws_SEPR_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1)) 

# #draws_MMN_threeway - chain converge and stationary
var_names <- c("group_event_coefs[1,1]" = "MMN_trend: autistic_standard",
               "group_event_coefs[2,1]" = "MMN_trend: non-autistic_standard",
               "group_event_coefs[3,1]" = "MMN_trend: autistic_pitch",
               "group_event_coefs[4,1]" = "MMN_trend: non-autistic_pitch",
               "group_event_coefs[5,1]" = "MMN_trend: autistic_length",
               "group_event_coefs[6,1]" = "MMN_trend: non-autistic_length",
               "group_event_coefs[7,1]" = "MMN_trend: autistic_pitch+length",
               "group_event_coefs[8,1]" = "MMN_trend: non-autistic_pitch+length") #define parameter names

g4a<-mcmc_trace(draws_MMN_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1))
g4b<-mcmc_dens(draws_MMN_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1)) 


#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend<-g_legend(g1a)

#arrange figures
g1<-grid.arrange(g1a+theme(legend.position = 'none'),g1b,top="BPS - task progression")
g2<-grid.arrange(g2a+theme(legend.position = 'none'),g2b,top="BPS - sequence position")
g3<-grid.arrange(g3a+theme(legend.position = 'none'),g3b,top="SEPR trend")
g4<-grid.arrange(g4a+theme(legend.position = 'none'),g4b,top="MMN trend")


g_figures<-grid.arrange(
  grid.arrange(g1,g2,ncol=2),
  g3,g4,
  nrow=3)


grid.arrange(g_figures,mylegend,ncol=2,widths=c(20,1))

#save to file
tiff(file=paste0(project_path,"/output/supplements/Bayesian_chains_convergence.tiff"), # create a file in tiff format in current working directory
  width=25, height=15, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(g_figures,mylegend,ncol=2,widths=c(20,1))
      
dev.off() #close operation and save file



```


## luminance within task

```{r}


g1<-ggplot(df_luminance,
       aes(x=trials,y=luminance_m),
       color='darkblue')+
  geom_smooth(aes(ymin=luminance_m-1.96*luminance_sd,
                  ymax=luminance_m+1.96*luminance_sd),fill='lightblue',stat = "identity")+
  geom_smooth(color='black',se=F)+
  geom_vline(xintercept=1400)+
  labs(x='trial number',y='relative luminance')

#plot in markdown
g1

#save to file
tiff(file=paste0(project_path,"/output/supplements/relative_luminance_stimulus_within_task.tiff"), # create a file in tiff format in current working directory
     width=5, height=2.5, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

g1

dev.off()

```


## covariate associations with BPS, SEPR, MMN

###BPS

```{r covariate associations BPS}

#full model  - save to supplements      
lmm_full<-lmer(scale(pd_res)~EventData_epoch*t1_diagnosis*stats::poly(EventCounter_epoch,3)+
      scale(luminance_m)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|id),data=df_epoch)

summary(lmm_full)

# luminance_eff<-cbind('video luminance',
#                round(fixef(lmm_full)['scale(luminance_m)'],2),
#                round(confint(lmm_full,parm = 'scale(luminance_m)'),2))
# 
# age_eff<-cbind('age (yrs)',
#                round(fixef(lmm_full)['scale(ageyrs)'],2),
#                round(confint(lmm_full,parm = 'scale(ageyrs)'),2))
# 
# samplingrate_eff<-confint(contrast(emmeans(lmm_full,~sampling_rate),
#                                    'revpairwise'))[c('contrast','estimate','asymp.LCL','asymp.UCL')]
# 
# centerdev_eff<-cbind('gaze center deviation (%)',
#                round(fixef(lmm_full)['scale(center_dev)'],2),
#                round(confint(lmm_full,parm = 'scale(center_dev)'),2))
# 
# 
# dquality_eff<-cbind('data quality / missing data (%)',
#                round(fixef(lmm_full)['scale(missing_data_trial)'],2),
#                round(confint(lmm_full,parm = 'scale(missing_data_trial)'),2))
# 
# 
# cov_eff_data<-rbind(luminance_eff,
#                     age_eff,
#       c('sampling rate (300Hz vs. 120Hz)',round(samplingrate_eff[1,2:4],2)),
#       centerdev_eff,
#       dquality_eff)
# 
# 
# cov_eff_table<-cov_eff_data %>%
#     kbl(caption = "Covariate effects on baseline pupil size (BPS)",
#         col.names = c('covariate','estimate','lower bound (2.5%)','upper bound (97.5%)'),
#         row.names = F) %>%
#     kable_classic(full_width = F, html_font = "Cambria")
#   
#   
#   cov_eff_table
#   
#   save_kable(cov_eff_table, file= paste0(project_path,'/output/supplements/table_BPS_covariateeffects.html'))

```

###SEPR

```{r covariate associations SEPR}

#full model  - save to supplements      
lmm_full<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|subjects)+(1|wave),data=df_trial)


dquality_eff<-cbind('missing data per trial (%)',
               round(fixef(lmm_full)['scale(missing_data_trial)'],2),
               round(confint(lmm_full,parm = 'scale(missing_data_trial)'),2))


cov_eff_data<-rbind(dquality_eff)

cov_eff_table<-cov_eff_data %>%
    kbl(caption = "Covariate effects on stimulus-evoked pupillary response (SEPR)",
        col.names = c('covariate','estimate','lower bound (2.5%)','upper bound (97.5%)'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  cov_eff_table
  
  save_kable(cov_eff_table, file= paste0(project_path,'/output/supplements/table_SEPR_covariateeffects.html'))

```

###MMN

```{r covariate associations MMN}

#full model  - save to supplements      
lmm_full<-lmer(scale(mmn)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|id),data=df_trial)


age_eff<-cbind('age (yrs)',
               round(fixef(lmm_full)['scale(ageyrs)'],2),
               round(confint(lmm_full,parm = 'scale(ageyrs)'),2))


dquality_eff<-cbind('missing data per trial (%)',
               round(fixef(lmm_full)['scale(missing_data_trial)'],2),
               round(confint(lmm_full,parm = 'scale(missing_data_trial)'),2))


cov_eff_data<-rbind(age_eff,
                    dquality_eff)

cov_eff_table<-cov_eff_data %>%
    kbl(caption = "Covariate effects on mismatch negativity (MMN)",
        col.names = c('covariate','estimate','lower bound (2.5%)','upper bound (97.5%)'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  cov_eff_table
  
  save_kable(cov_eff_table, file= paste0(project_path,'/output/supplements/table_MMN_covariateeffects.html'))

```




## moderation of BPS on SEPR and MMN interaction by group

```{r moderation by BPS}

#moderation of BPS on SEPR interaction with group
  lmm_REML_BPS<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+scale(pd_baseline)+
                   (1|subjects)+(1|wave),data=df_trial)

  anova(lmm_REML_BPS)
  
  
  ##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'BPS',
               'S x group','S x TP','group x TP','S x group x TP')
  
  
  table_data_model<-round(anova(lmm_REML_BPS),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: Moderation of group differences in dynamic stimulus-evoked pupillary response (SEPR) group differences by baseline pupil size (BPS)",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_moderation_SEPR_by_BPS.html'))


  #moderation of BPS on MMN interaction with group
  lmm_REML_BPS<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+scale(pd_baseline)+
                       (1|subjects)+(1|wave),data=df_trial)

  anova(lmm_REML_BPS)

  ##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'BPS',
               'S x group','S x TP','group x TP','S x group x TP')
  
  
  table_data_model<-round(anova(lmm_REML_BPS),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: Moderation of group differences in dynamic mismatch negativity (MMN) by baseline pupil size (BPS)",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_moderation_MMN_by_BPS.html'))

  
```


## intraclass correaltion of the dependent variables

```{r ICC_of_dependent_variables}

#select only individuals that have both waves
subjects_with_both_waves<-df_dem$subjects[df_dem$timepoints=='wave1+2']
df_trial_wave12<-df_trial[df_trial$subjects %in% subjects_with_both_waves,]

#BPS
baseline_pd_wave1<-with(df_trial_wave12[df_trial_wave12$wave=='wave1',],aggregate(pd_baseline,by=list(subjects),FUN=mean,na.rm=T))
baseline_pd_wave2<-with(df_trial_wave12[df_trial_wave12$wave=='wave2',],aggregate(pd_baseline,by=list(subjects),FUN=mean,na.rm=T))
names(baseline_pd_wave1)<-c('subjects','wave1')
names(baseline_pd_wave2)<-c('subjects','wave2')

baseline_pd_waves<-merge(baseline_pd_wave1,baseline_pd_wave2,by='subjects')

cor.test(baseline_pd_waves$wave1,baseline_pd_waves$wave2)

psych::ICC(baseline_pd_waves[,-1])
###--> ICC: 0.88 [0.81, 0.93]


#SEPR
rpd204_wave1<-with(df_trial_wave12[df_trial_wave12$EventData_epoch=='204' & df_trial_wave12$wave=='wave1',],aggregate(rpd_auc.204,by=list(subjects),FUN=mean,na.rm=T))
rpd204_wave2<-with(df_trial_wave12[df_trial_wave12$EventData_epoch=='204' & df_trial_wave12$wave=='wave2',],aggregate(rpd_auc.204,by=list(subjects),FUN=mean,na.rm=T))

names(rpd204_wave1)<-c('subjects','wave1')
names(rpd204_wave2)<-c('subjects','wave2')

rpd_waves<-merge(rpd204_wave1,rpd204_wave2,by='subjects')

cor.test(rpd_waves$wave1,rpd_waves$wave2)
##--> r =.23

psych::ICC(rpd_waves[,-1])
###--> ICC: 0.37 [-0.01, 0.60]


names(df_timepoint)

#NG
gain_wave1<-with(df_timepoint[df_timepoint$wave=='wave1',],aggregate(gain_pcdm,by=list(subjects),FUN=mean,na.rm=T))
gain_wave2<-with(df_timepoint[df_timepoint$wave=='wave2',],aggregate(gain_pcdm,by=list(subjects),FUN=mean,na.rm=T))

names(gain_wave1)<-c('subjects','wave1')
names(gain_wave2)<-c('subjects','wave2')

gain_waves<-merge(gain_wave1,gain_wave2,by='subjects')

cor.test(gain_waves$wave1,gain_waves$wave2)
##--> r =.23

psych::ICC(gain_waves[,-1])
###--> ICC: 0.37 [0.02, 0.59]

```

For participants with both measurement timepoints, we estimated a relative stability of BPS (intraclass correlation, ICC = 0.88 [0.81, 0.93]), while SEPR (ICC = 0.37 [-0.01, 0.60]) and NG (ICC = 0.37 [0.02, 0.59]) were variable across timepoints. EEG data was only available at one timepoint. 



