---
title: "data_analysis_oddball_leap_280423"
author: "Nico Bast"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

# setup

```{r setup, include=FALSE}

#chunk options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) #warning to false

## REQUIRED PACKAGES

#analysis
require(lme4) #linear mixed models
require(lmerTest) #p-value for fixed effects in LMM
require(emmeans) #estimated marginalized means
require(performance) #r²_nakagawa - marginalized and conditionalk R² 

#packages for MMN
require(pbapply) #progress bar apply
require(zoo) #rolling mean
require(data.table) #rbindlist

#visualization
require(kableExtra) #APA tables
library(greta) #Bayesian modelling
require(bayesplot) #plotting Bayesian
require(bayestestR) #credible interval ci, point estimates Bayesian
require(DiagrammeR) #plotting Bayesian
require(ggplot2) #plot posterior sampling estimates

require(R.matlab) # read matlab files
require(effectsize) # cohens D

require(RColorBrewer) #color palettes
require(gridExtra) #arrange plots
require(wesanderson) #custom color palettes

require(eegkit) #standard coordinates of EEG channels - data(eegcoord)
require(eegUtils)

require(car) #Anova function for type 3 SumSq 

#INFO: eegUtils is not on CRAN and thus loaded with remote packages from github
# install.packages("remotes")
# remotes::install_github("craddm/eegUtils")



#paths
project_path<-getwd()
#project_path<-"C:/Users/nico/PowerFolders/project_oddball_LEAP"
localation_of_EEG_allchannel_data<-"C:/Users/nico/Desktop/leap_oddball_eeg_list_allchannels.rdata"

##define parameters of posterior sampling in Bayesian modeling
number_of_warmups <- 4000
number_of_iterations <- 8000
credible_interval<-0.89 #credible interval Bayesian estimates


##define percentage of data sampling for visualization
fraction_to_sample<-100 ###relevant for scalp topography
sampling_factor<-1/fraction_to_sample ### releavnt for MMN, pupil size change figures

###graphic settings

  #graphic parameters
  theme_set(theme_bw())
  
  #define color palette
  custom_condition_colors <- wes_palette('FantasticFox1',5,type='discrete')[2:5]
  custom_contrast_colors <- c(brewer.pal(6, "Blues")[4:6],brewer.pal(6, "Oranges")[4:6])
  custom_group_color <- brewer.pal(3, "Dark2")[1:2]
 
  #RColorBrewer::display.brewer.all()

```

# load data

Eye tracking and EEG data in separate file that are already preprocessed, see

<li> [Eye-Tracking preprocessing]: https://github.com/nicobast/oddball_LEAP/blob/master/code/data_preprocessing_oddball_leap.R </li>
<li> [EEG preprocessing]: https://github.com/nicobast/oddball_LEAP/blob/master/code/read_eeg_data.R </li>


```{r load_data_ET_EEG}

## pupil data
load(file=paste0(project_path,'/data/mmn_leap_pd_final_dfs_21022023'))

## EEG data - reduced to FZ channel - see read_eeg_data.R
#load(paste0(project_path,"/data/leap_oddball_eeg_list.rdata"))
list_eeg_redux<-readRDS(paste0(project_path,"/data/leap_oddball_eeg_list_redux.rds"))


```

Also load and merge the data generated by the PCDM computational model, see

<li> [PCDM preprocessing]: https://github.com/nicobast/oddball_LEAP/blob/master/code/PCDM_data_conversion_matlab.R </li>
<li> [PCDM computation]: https://github.com/nicobast/PCDM </li>

```{r load_data_PCDM}

#read data estimated in MAtlab
mat_output<-readMat(paste0(project_path,"/data/pcdm_estimates_1run_290323.mat"))

#extract parameters
pcdm_parameters<-mat_output[[2]]
names(pcdm_parameters)<-c('interpolateBlinks','downsampleRate','fitTimeseries','change_cutoff_by_samplingRate','low_bandpass_filter_range','blinkint_vel_thres_on','blinkint_window_from_onset','blinkint_min_dur','rel_vel_thres','min_sac_dur')

#create empty list
list_pcdm_estimates<-list()

### extract required info from matlab object in for loop
for (i in seq(length(mat_output[[1]]))){

pcdm_estimate<-mat_output[[1]][i][[1]][[1]] #extract relevant data of single participant
names(pcdm_estimate)<-dimnames(pcdm_estimate)[[1]] #assign names of dimensions
pcdm_estimate<-lapply(pcdm_estimate,unlist) # remove internal list structure

list_pcdm_estimates[[i]]<-pcdm_estimate

}

#add names
names(list_pcdm_estimates)<-sapply(list_pcdm_estimates,function(x){x['name']})

#create vectors
gain_pcdm<-unlist(sapply(list_pcdm_estimates,function(x){x['gain']}))
Rsq_pcdm<-unlist(sapply(list_pcdm_estimates,function(x){x['Rsq']}))
offset_pcdm<-unlist(sapply(list_pcdm_estimates,function(x){x['offset']}))
id<-substr(names(list_pcdm_estimates),1,nchar(names(list_pcdm_estimates))-4)
id<-id[id!=""] #remove empty entries

#exclude outliers
gain_pcdm[gain_pcdm>0.2 | gain_pcdm<(-0.2)]<-NA

#merge data
df_pcdm<-data.frame(id,gain_pcdm,Rsq_pcdm,offset_pcdm)
df_timepoint<-merge(df_timepoint,df_pcdm,by='id',all.x=T)


```

# calculate mismatch negativity (MMN)

```{r calculate mismatch negativity}

###---> calculate MMN ####

fun_estimate_mmn<-function(one_set){

  #testing
  #one_set<-list_eeg[[100]]

  ###restrict to timeframe - 50-350ms
  one_set<-one_set[one_set$times>50 & one_set$times<350,]

  mmn_trial<-with(one_set,by(Fz,trial_counter,function(channel_data){

      channel_data_smooth<-rollmean(channel_data,k=20) #rolling mean of 20ms
      min(channel_data_smooth)

  }))

  trial_counter<-names(mmn_trial)
  mmn<-as.numeric(mmn_trial)
  mmn_per_trial<-data.frame(trial_counter,mmn)
  return(mmn_per_trial)

}

list_mmn<-pblapply(list_eeg_redux,fun_estimate_mmn)

##--> create data frame
ID<-substr(names(list_mmn),10,22)
condition<-substr(names(list_mmn),1,8)

list_mmn<-pbmapply(function(x,y,z){
  x$condition<-y
  x$ID<-z
  return(x)
  },x=list_mmn,y=condition,z=ID,SIMPLIFY = F)

df_mmn<-data.table::rbindlist(list_mmn)

df_mmn$trial_counter<-as.numeric(df_mmn$trial_counter)
df_mmn$condition<-as.factor(df_mmn$condition)

    
```

# merge eye-tracking and EEG data

```{r merge ET and EEG}

##prepare EEG data
df_mmn$ID<-sub('/','',df_mmn$ID)
df_mmn$ID<-sub('_','',df_mmn$ID)

    #with(df_timepoint,table(subjects,wave))

    #eye-tracking data with MMN data
    table(df_timepoint$subjects %in% df_mmn$ID)
    table(unique(df_mmn$ID) %in% df_timepoint$subjects)

#create merge variables    
df_mmn$merge_id<-interaction(df_mmn$ID,df_mmn$trial_counter)
df_trial$merge_id<-interaction(df_trial$subjects,df_trial$EventCounter)

#merge - per-trial level (df_trial)
df_trial<-merge(df_trial,df_mmn,by='merge_id',all.x=T)

    table(is.na(df_mmn$mmn))
    table(is.na(df_trial$rpd_auc))
    table(is.na(df_trial$mmn)) #missing in MMN --> to be expected
    ##--> not all trials analyzed in eye-tracking were selected for EEG
    
##some trials are not correctly matched - correct for that
with(df_trial,table(condition,EventData))
df_trial_mmn<-df_trial[(df_trial$EventData=='201' & df_trial$condition=='standard'|
                 df_trial$EventData=='203' & df_trial$condition=='duration'|
                 df_trial$EventData=='202' & df_trial$condition=='frequenc'|
                 df_trial$EventData=='204' & df_trial$condition=='combined'),]

###merge per-participant level (df_timepoint)
df_mmn_timepoint<-aggregate(mmn~id+EventData,data=df_trial_mmn,FUN=mean,na.rm=T)
df_mmn_timepoint<-reshape(df_mmn_timepoint, idvar = "id", timevar = "EventData", direction = "wide")

df_timepoint<-merge(df_timepoint,df_mmn_timepoint,by='id',all.x = T)
rm(df_mmn_timepoint)

###which participants have MMN data
table(is.na(df_timepoint$mmn.201),df_timepoint$t1_diagnosis) ##

```

# descriptives table of sample

generates a table on characteristics of the sample between groups

```{r samples description}

#### sample descriptives - PARTICIPANT specific ####

fun_return_descriptive<-function(variable,group,rounding=2,scaling=1){
  mean_values<-by(variable,group,function(x){round(mean(x,na.rm=T),rounding)})
  sd_values<-by(variable,group,function(x){round(sd(x,na.rm=T),rounding)})
  min_values<-by(variable,group,function(x){round(min(x,na.rm=T),rounding)})
  max_values<-by(variable,group,function(x){round(max(x,na.rm=T),rounding)})
  mean_values<-mean_values*scaling
  sd_values<-sd_values*scaling
  min_values<-min_values*scaling
  max_values<-max_values*scaling
  paste0(mean_values,'/',sd_values,' ','[',min_values,'-',max_values,']')
}

#standard descriptives
n_group<-with(df_dem,by(subjects,t1_diagnosis,function(x){length(unique(x))}))
gender<-with(df_dem,by(t1_sex,t1_diagnosis,function(x){paste0(table(x)[2],'/',table(x)[1])}))
timepoints<-with(df_dem,by(timepoints,t1_diagnosis,function(x){paste0(table(x)[1],'/',table(x)[2],'/',table(x)[3])}))
age<-with(df_dem,fun_return_descriptive(variable=t1_ageyrs,group=t1_diagnosis))
iq<-with(df_dem,fun_return_descriptive(variable=t1_fsiq,group=t1_diagnosis))
piq<-with(df_dem,fun_return_descriptive(variable=t1_piq,group=t1_diagnosis))
viq<-with(df_dem,fun_return_descriptive(variable=t1_viq,group=t1_diagnosis))

#clinical variables
srs<-with(df_dem,fun_return_descriptive(variable=t1_srs_rawscore,group=t1_diagnosis))
rbs_total<-with(df_dem,fun_return_descriptive(variable=t1_rbs_total,group=t1_diagnosis))
sdq_total<-with(df_dem,fun_return_descriptive(variable=t1_sdq_total_difficulties_p,group=t1_diagnosis))
adhd_inatt<-with(df_dem,fun_return_descriptive(variable=adhd_inatt,group=t1_diagnosis))
adhd_hyper<-with(df_dem,fun_return_descriptive(variable=adhd_hyper,group=t1_diagnosis))
anx_beck<-with(df_dem,fun_return_descriptive(variable=anx_beck,group=t1_diagnosis))
dep_beck<-with(df_dem,fun_return_descriptive(variable=dep_beck,group=t1_diagnosis))

##data quality
#precision<-with(df_dem,fun_return_descriptive(variable=Precision,group=t1_diagnosis))
#accuracy<-with(df_dem,fun_return_descriptive(variable=Accuracy,group=t1_diagnosis))
#sampling_rate<-with(df_dem,by(sampling_rate,t1_diagnosis,function(x){paste0(table(x)[2],'/',table(x)[1])}))
missing_data_trial<-with(df_dem,fun_return_descriptive(variable=missing_data_trial,group=t1_diagnosis,rounding=4,scaling = 100))
sampling_rate<-with(df_timepoint,by(sampling_rate,t1_diagnosis,function(x){paste0(table(x)[2],'/',table(x)[1])}))

#et vars
center_dev_id<-with(df_dem,fun_return_descriptive(variable=center_dev,group=t1_diagnosis,rounding=4,scaling=100))
screen_dist_id<-with(df_dem,fun_return_descriptive(variable=screen_dist,group=t1_diagnosis))
pd_baseline_id<-with(df_dem,fun_return_descriptive(variable=pd_baseline,group=t1_diagnosis))
# rpd_response201_id<-with(df_dem,fun_return_descriptive(variable=rpd_response.201,group=t1_diagnosis,rounding=4,scaling=100))
# rpd_response202_id<-with(df_dem,fun_return_descriptive(variable=rpd_response.202,group=t1_diagnosis,rounding=4,scaling=100))
# rpd_response203_id<-with(df_dem,fun_return_descriptive(variable=rpd_response.203,group=t1_diagnosis,rounding=4,scaling=100))
# rpd_response204_id<-with(df_dem,fun_return_descriptive(variable=rpd_response.204,group=t1_diagnosis,rounding=4,scaling=100))
# rpd_auc201_id<-with(df_dem,fun_return_descriptive(variable=rpd_auc.201,group=t1_diagnosis,rounding=4,scaling=100))
# rpd_auc202_id<-with(df_dem,fun_return_descriptive(variable=rpd_auc.202,group=t1_diagnosis,rounding=4,scaling=100))
# rpd_auc203_id<-with(df_dem,fun_return_descriptive(variable=rpd_auc.203,group=t1_diagnosis,rounding=4,scaling=100))
# rpd_auc204_id<-with(df_dem,fun_return_descriptive(variable=rpd_auc.204,group=t1_diagnosis,rounding=4,scaling=100))


rpd_auc_id<-with(df_dem,fun_return_descriptive(variable=rpd_auc,group=t1_diagnosis,rounding=4,scaling=100))


descriptives_table<-rbind(n_group,gender,timepoints,age,iq,piq,viq,
                          missing_data_trial,sampling_rate,
                          center_dev_id,screen_dist_id,pd_baseline_id,
                          srs,rbs_total,sdq_total,adhd_inatt,adhd_hyper,anx_beck,dep_beck
                          #rpd_response201_id,rpd_response202_id,rpd_response203_id,rpd_response204_id,
                          #rpd_auc201_id,rpd_auc202_id,rpd_auc203_id,rpd_auc204_id
                          )

##group differences
groupdiff_p<-c(NA,
               round(as.numeric(with(df_dem,chisq.test(t1_sex,t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,chisq.test(timepoints,t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_ageyrs~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_fsiq~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_piq~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_viq~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(missing_data_trial~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_timepoint,chisq.test(sampling_rate,t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(center_dev~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(screen_dist~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(pd_baseline~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_srs_rawscore~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_rbs_total~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(t1_sdq_total_difficulties_p~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(adhd_inatt~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(adhd_hyper~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(anx_beck~t1_diagnosis))['p.value']),3),
               round(as.numeric(with(df_dem,t.test(dep_beck~t1_diagnosis))['p.value']),3)
               # round(as.numeric(with(df_dem,t.test(rpd_response.201~t1_diagnosis))['p.value']),3),
               # round(as.numeric(with(df_dem,t.test(rpd_response.202~t1_diagnosis))['p.value']),3),
               # round(as.numeric(with(df_dem,t.test(rpd_response.203~t1_diagnosis))['p.value']),3),
               # round(as.numeric(with(df_dem,t.test(rpd_response.204~t1_diagnosis))['p.value']),3),
               # round(as.numeric(with(df_dem,t.test(rpd_auc.201~t1_diagnosis))['p.value']),3),
               # round(as.numeric(with(df_dem,t.test(rpd_auc.202~t1_diagnosis))['p.value']),3),
               # round(as.numeric(with(df_dem,t.test(rpd_auc.203~t1_diagnosis))['p.value']),3),
               # round(as.numeric(with(df_dem,t.test(rpd_auc.204~t1_diagnosis))['p.value']),3)
               )

groupdiff_p[which(groupdiff_p==0)]<-'<0.001'
groupdiff_p[which(is.na(groupdiff_p))]<-'-'

row_names<-c('n','gender (F/M)','timepoints (1/2/1+2)','age (in years)','IQ','perceptual IQ','verbal IQ',
             #'data quality: precision','data quality: accuracy',
             #'sampling rate (120Hz/300Hz)',
             'missing data per trial (%)','sampling rate (300Hz / 120Hz)',
             'gaze center deviation (%)','screen distance (mm)','mean pupil size (mm)',
             'SRS (total)','RBS (total)','SDQ (total)','ADHD inattention','ADHD hyperactivity','BAI (anxiety)','BDI (depression)'
             # 'pupillary response: amplitude - standard (%)','pupillary response: amplitude - oddball pitch (%)','pupillary response: amplitude - oddball length (%)','pupillary response: amplitude - oddball both (%)',
             # 'pupillary response: AUC - standard (%)','pupillary response: AUC - oddball pitch (%)','pupillary response: AUC - oddball length (%)','pupillary response: AUC - oddball both (%)'
             )

descriptives_table<-cbind(row_names,descriptives_table,groupdiff_p)

#add group differences
#descriptives_table<-ifelse(descriptives_table=='0','<0.001',descriptives_table)

table_sample<-descriptives_table %>%
  kbl(caption = "Sample description",
      col.names = c('','autistic individuals (ASD)','non-autistic individuals (NTC)','group difference (p)'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")


descriptives_table

table_sample

save_kable(table_sample, file=paste0(project_path,'/output/table_sampledescription.html'))




```

# group differences of dependent variables (per participant)

```{r group differences per participant}

#BPS
summary(lm<-lm(scale(pd)~t1_diagnosis,df_timepoint))
cohens_d(x=scale(pd)~t1_diagnosis,data=df_timepoint)

#NG
summary(lm(scale(gain_pcdm)~t1_diagnosis,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))
cohens_d(x=scale(pd)~t1_diagnosis,data=df_timepoint[df_timepoint$Rsq_pcdm>0.2,])

```

# covariate effects on dependent variables (per participant)

```{r covariate effects on dependent variables per participant}

#BPS
c(summary(lm(scale(pd)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(pd)~scale(ageyrs),df_timepoint))[2,])

#MMN - standards 
c(summary(lm(scale(mmn.201)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.201)~scale(ageyrs),df_timepoint))[2,])

#MMN - pitch oddballs 
c(summary(lm(scale(mmn.202)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.202)~scale(ageyrs),df_timepoint))[2,])

#MMN - length oddballs 
c(summary(lm(scale(mmn.203)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.203)~scale(ageyrs),df_timepoint))[2,])

#MMN - pitch & length oddballs
c(summary(lm(scale(mmn.204)~scale(ageyrs),df_timepoint))[['coefficients']][2,1],
      confint(lm(scale(mmn.204)~scale(ageyrs),df_timepoint))[2,])


```


# associations of dependent variables (per participant)

correlation coefficients on per-participant level of

<li> pupillary measure: baseline pupil size (BPS) </li>
<li> pupillary measure: stimulus-evoked pupilalry response (SEPR) </li>
<li> EEG measure: mismatch negativity (MMN) </li>
<li> computational model: neural gain (NG) </li>

Event Data coding

<li> 201 = standard </li>
<li> 202 = pitch </li>
<li> 203 = length </li>
<li> 204 = pitch & length </li>


```{r associations of dependent variables}

rm(gain_pcdm)
attach(df_timepoint)

cor_table<-data.frame(t(cbind(

#SEPR -PD
cbind(cor.test(rpd_auc.201,pd)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_auc.202,pd)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_auc.203,pd)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_auc.204,pd)[c('data.name','estimate','p.value')]),

#SEPR - MMN --> uncorrelated
cbind(cor.test(rpd_auc.201,mmn.201)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_auc.202,mmn.202)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_auc.203,mmn.203)[c('data.name','estimate','p.value')]),
cbind(cor.test(rpd_auc.204,mmn.204)[c('data.name','estimate','p.value')]),

#BPS - MMN --> substantial negative correlations
cbind(cor.test(pd,mmn.201)[c('data.name','estimate','p.value')]),
cbind(cor.test(pd,mmn.202)[c('data.name','estimate','p.value')]),
cbind(cor.test(pd,mmn.203)[c('data.name','estimate','p.value')]),
cbind(cor.test(pd,mmn.204)[c('data.name','estimate','p.value')]),

# NG - BPS
cbind(cor.test(gain_pcdm,pd)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,pd)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,pd)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,pd)[c('data.name','estimate','p.value')]),

#NG - SEPR
cbind(cor.test(gain_pcdm,rpd_auc.201)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,rpd_auc.202)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,rpd_auc.203)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,rpd_auc.204)[c('data.name','estimate','p.value')]),

#NG - MMN
cbind(cor.test(gain_pcdm,mmn.201)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,mmn.202)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,mmn.203)[c('data.name','estimate','p.value')]),
cbind(cor.test(gain_pcdm,mmn.204)[c('data.name','estimate','p.value')])


)))


#format table
cor_table$data.name<-unlist(cor_table$data.name)
cor_table$estimate<-unlist(cor_table$estimate)
cor_table$p.value<-unlist(cor_table$p.value)

cor_table$estimate<-round(cor_table$estimate,4)
cor_table$p.value<-round(cor_table$p.value,4)

cor_table$data.name<-gsub('rpd_auc','SEPR',cor_table$data.name)
cor_table$data.name<-gsub('pd','BPS',cor_table$data.name)
cor_table$data.name<-gsub('gain_pcdm','NG',cor_table$data.name)
cor_table$data.name<-gsub('mmn','MMN',cor_table$data.name)
cor_table$data.name<-gsub('.201',' (standard)',cor_table$data.name)
cor_table$data.name<-gsub('.202',' (pitch oddball)',cor_table$data.name)
cor_table$data.name<-gsub('.203',' (length oddball)',cor_table$data.name)
cor_table$data.name<-gsub('.204',' (pitch+length oddball)',cor_table$data.name)



#Bonferroni correction
number_of_comparisons<-nrow(cor_table)
cor_table$p_adjust<-cor_table$p.value*number_of_comparisons
cor_table$p_adjust<-ifelse(cor_table$p_adjust>1,1,cor_table$p_adjust)

#produce table
cor_table_format<-cor_table %>%
    kbl(caption = "correlations of baseline pupil size (BPS), 
        stimulus-evoked pupillary response (SEPR), 
        mismatch negativity (MMN), and neural gain (NG) for different stimuli.",
        col.names = c('correlation','coefficient (r)','p','p adjusted')) %>%
    kable_classic(full_width = F, html_font = "Cambria")

cor_table_format

save_kable(cor_table_format,file=paste0(project_path,'/output/supplements/associations_DV_perparticipant.html'))
  
```

# association of dependent variables (per trial)

```{r BPS_SEPR_association}

#interrelation of BPS and SEPR
lmm<-lmer(scale(rpd_auc)~scale(pd_baseline)+
            (1|subjects)+(1|wave),data=df_trial)
anova(lmm)

cbind(round(fixef(lmm)['scale(pd_baseline)'],2),
              round(confint(lmm,parm = 'scale(pd_baseline)'),2))

#model fit
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(rpd_auc)~scale(pd_baseline)+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(rpd_auc)~1+
                   (1|subjects)+(1|wave),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  

```  

```{r BPS_SEPR_association_Bayesian}

 # data - create greta arrays
    x <- as_data(scale(df_trial$pd_baseline))
    y <- as_data(scale(df_trial$rpd_auc))
    subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)

    
    # variables and priors
    int <- normal(0, 1)
    coef <- normal(0, 1)
    sd <- student(3, 0, 1, truncation = c(0, Inf))
    a_subject <- normal(0, 1, dim = max(subject_id))

    # operations
    mean <- int + coef * x + a_subject[subject_id]

    # likelihood - distribution over data
    distribution(y) <- normal(mean, sd)

    # defining the model
    m <- model(int, coef, sd)

    # plotting
    #require(DiagrammeR)
    plot(m)

    # sampling
    draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups)
    summary(draws)
    
    mcmc_trace(draws) #see whether chains converge and are stationary
    mcmc_dens(draws) # + stable estimates
    
    #posterior predictive checking
    coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
    coda::effectiveSize(draws) #ESS

    #save model
    draws_SEPR_BPS<-draws
    save(draws_SEPR_BPS,file=paste0(project_path,'/data/Bayesian_model_SEPR_BPS.Rdata'))


```

```{r BPS MMN association LMM}

###here use ID as random intercept (subject + wave) as MMN is only available for one timepoint

#interrelation of BPS and SEPR
lmm<-lmer(scale(mmn)~scale(pd_baseline)*EventData+
            (1|id),data=df_trial)

anova(lmm)

cbind(round(fixef(lmm)['scale(pd_baseline)'],2),
              round(confint(lmm,parm = 'scale(pd_baseline)'),2))

#model fit
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(mmn)~scale(pd_baseline)+EventData+
              (1|id),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(mmn)~EventData+
                   (1|id),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  



```

```{r SEPR MMN association LMM}

###here use ID as random intercept (subject + wave) as MMN is only available for one timepoint

#interrelation of BPS and SEPR
lmm<-lmer(scale(mmn)~scale(rpd_auc)*EventData+
            (1|id),data=df_trial)

anova(lmm)

cbind(round(fixef(lmm)['scale(rpd_auc)'],2),
              round(confint(lmm,parm = 'scale(rpd_auc)'),2))

#model fit
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(mmn)~scale(rpd_auc)+
              (1|id),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(mmn)~1+
                   (1|id),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  


```

# composite figure - pupil size and Fz amplitude change within trials

visualization of pupile size change (BPS,SEPR) and Fz amplitude change (MMN) within trials

## MMN - scalp topography (part 1)

```{r scalp topography, warnings=F}

#load all EEG channel data - LARGE FILE ON DIFFERENT DRIVE
load(file=localation_of_EEG_allchannel_data)

#for standard coordinates of electrodes
data(eegcoord) #load sample set with standard coordinates
#rownames(eegcoord) #names of channels

###select only participants that are in final sample
subjects_in_list<-substr(names(list_mmn),10,22)
subjects_in_list<-sub('/','',subjects_in_list)
subjects_in_list<-sub('_','',subjects_in_list)
list_elements_in_final_sample<-which(subjects_in_list %in% unique(df_trial$subjects))
list_eeg_selected<-list_eeg[list_elements_in_final_sample]

#participants_to_sample<-round(length(list_eeg)/4*fraction_to_sample)
participants_to_sample<-round(length(list_eeg_selected)/4*fraction_to_sample)

#select 30 random from each condition - same participants
combined_sampled<-sample(1:(length(list_eeg_selected)/4),participants_to_sample)
duration_sampled<-combined_sampled+length(list_eeg_selected)/4
frequency_sampled<-combined_sampled+2*length(list_eeg_selected)/4
standard_sampled<-combined_sampled+3*length(list_eeg_selected)/4

select_samples<-c(combined_sampled,
                  duration_sampled,
                  frequency_sampled,
                  standard_sampled)

#select thirty random participants - from participants that have ET data
list_eeg_selected<-list_eeg_selected[select_samples]

fun_prepare_channeldata_plotting<-function(one_set,sample_coordinates){

  ##testing
  # one_set<-list_eeg_selected[[1]]
  # sample_coordinates<-eegcoord
  
  #match names of sample coordnisates and data set
  names(one_set)<-toupper(names(one_set)) #capitalize names - naming convention of eegkit
  sample_coordinates<-sample_coordinates[rownames(sample_coordinates) %in% names(one_set),]
  sample_coordinates$variable<-rownames(sample_coordinates)


  #limit site to interested timeframe - 150-250 ms after trial onset
  one_set<-one_set[one_set$TIMES >= 150 & one_set$TIMES <= 250,]
      ###--> less flexibel but reduces file size
  
  #melt to long format
  one_set$TIMES<-as.factor(one_set$TIMES) #preserve in melting
  one_set<-reshape2::melt(one_set,value.name='amplitude') #convert to long format

  #merge with sampel coordinates
  one_set<-merge(one_set,sample_coordinates,by='variable') #merge with sample coordniates
  names(one_set)[1]<-'electrode' #set name for later plotting function
  one_set$TIMES<-as.numeric(levels(one_set$TIMES))[one_set$TIMES]
  
  return(one_set)

      }

#takes around 20 minutes
list_eeg_selected<-pblapply(list_eeg_selected,fun_prepare_channeldata_plotting,sample_coordinates=eegcoord)
df_plot<-data.table::rbindlist(list_eeg_selected)

#scalp topography separated for condition
limits_of_plotted_range<-c(-3,3)

gf_4a<-topoplot(df_plot[df_plot$CONDITION=='standard',], 
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='standard') #scale coordniates to head

gf_4b<-topoplot(df_plot[df_plot$CONDITION=='frequenc',], 
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='pitch oddball') #scale coordniates to head

gf_4c<-topoplot(df_plot[df_plot$CONDITION=='duration',], 
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='length oddball') #scale coordniates to head

gf_4d<-topoplot(df_plot[df_plot$CONDITION=='combined',],
         interp_limit='head',limits=limits_of_plotted_range)+labs(title='pitch & length oddball') #scale coordniates to head


#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend<-g_legend(gf_4a)

#plot in markdown
grid.arrange(
  arrangeGrob(
    gf_4a + theme(legend.position="none"),
    gf_4b + theme(legend.position="none"),
    gf_4c + theme(legend.position="none"),
    gf_4d + theme(legend.position="none"),nrow=2),
  mylegend, ncol=2, widths=c(5,1))


```

## MMN - Fz amplitude change within trials (part 2)

```{r Fz amplitude change}

df_eeg<-data.table::rbindlist(list_eeg_redux) ##very fast compared to do.call(rbind)

subsample<-sample(1:nrow(df_eeg),nrow(df_eeg)/sampling_factor) # select 10% of data for faster plotting

    ### use participants selected for scalp topography - much slower and same results
    # df_eeg$ID<-sub('/','',df_eeg$ID)
    # df_eeg$ID<-sub('_','',df_eeg$ID)
    # selected_participants<-substr(names(list_mmn)[combined_sampled],10,21)
    # subsample<-which(df_eeg$ID %in% selected_participants)

custom_condition_colors <- rev(wes_palette('FantasticFox1',5,type='discrete')[2:5]) #reverse custom colors to match color coding in other figures

gf_1<-ggplot(df_eeg[subsample,],aes(x=times/1000,y=Fz,group=condition,color=condition,fill=condition))+
  geom_smooth()+
  labs(y='electrode Fz - amplitute (uV)',x='trial duration (s)')+
  scale_fill_manual(values = custom_condition_colors, labels=c("standard" = "standard", "frequenc" = "pitch oddball", "duration" = "length oddball ", "combined" = "pitch & length oddball"))+
  scale_color_manual(values = custom_condition_colors, labels=c("standard" = "standard", "frequenc" = "pitch oddball", "duration" = "length oddball ", "combined" = "pitch & length oddball"))+
  xlim(-0.1,0.6) #allign with eye-tracking epoch for plotting

#gf_1

```


## pupil size change within trial - progression between groups (part 3+4)

```{r pupil size change signal}

#sample data
sampled_df<-df[sample(1:nrow(df),nrow(df)/sampling_factor),]
#ON FULL DATA - takes 1hour+ to complete

gf_2<-ggplot(sampled_df[sampled_df$time_event<0.6,],aes(x=time_event,y=rpd,group=EventData,color=EventData,fill=EventData))+
  geom_smooth(alpha=0.4)+
  labs(x='trial duration (s)',y='pupil size (mm)')+
  scale_fill_manual(values = custom_condition_colors[c(1,3,2,4)], labels=c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball ", "204" = "pitch & length oddball"))+ #[c(1,3,2,4)] = colors like EEG fig
  scale_color_manual(values = custom_condition_colors[c(1,3,2,4)], labels=c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball ", "204" = "pitch & length oddball"))+
  xlim(-0.1,0.6) #allign with eeg epoch


#add demographics data
df_dem_merge<-df_dem[,c('subjects','t1_group','t1_diagnosis','t1_sex','t1_ageyrs','t2_ageyrs','t1_fsiq','t1_piq','t1_viq')]
df_dem_merge$subjects<-as.character(df_dem_merge$subjects)
sampled_df$subjects<-substr(sampled_df$id,1,12)
sampled_df<-merge(sampled_df,df_dem_merge,by='subjects') ###--> takes a while

    
#define quartiles
sampled_df$trial_position<-ifelse(sampled_df$EventCounter<350,'first quarter',
  ifelse(sampled_df$EventCounter>1050,'last quarter',
         ifelse(sampled_df$EventCounter>=350 & sampled_df$EventCounter<=700,'second quarter',
                ifelse(sampled_df$EventCounter>700 & sampled_df$EventCounter<=1050,'third quarter',NA
                                       ))))

sampled_df$trial_position<-factor(sampled_df$trial_position,
                                  levels=c('first quarter','second quarter','third quarter','last quarter'))

#plot split by group and quartiles
gf_3<-ggplot(sampled_df[sampled_df$time_event<0.60 & !is.na(sampled_df$trial_position),],
       aes(x=time_event,y=rpd,group=interaction(t1_diagnosis,EventData),
           color=EventData,fill=EventData,linetype=t1_diagnosis))+
  facet_grid(EventData ~ trial_position)+
  geom_smooth()+
  #geom_smooth(formula = y ~ x + poly(x,4))
  scale_linetype_discrete(labels=c("ASD" = "autistic", "Control" = "non-autistic"))+
  scale_fill_manual(values = custom_condition_colors,
                     labels=c("201" = "standard",
                              "202" = "pitch oddball",
                              "203" = "length oddball ",
                              "204" = "pitch & length oddball"))+
  scale_color_manual(values = custom_condition_colors,
                     labels=c("201" = "standard",
                              "202" = "pitch oddball",
                              "203" = "length oddball ",
                              "204" = "pitch & length oddball"))+
  labs(x='trial duration (s)',y='pupil size (mm)',color='stimulus',fill='stimulus',linetype='group')+
  theme(strip.text.y = element_blank()) #remove labels on facet axis


#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend_pupil<-g_legend(gf_3)

#print in markdown
#grid.arrange(gf_2+theme(legend.position = 'none'),gf_3,widths=c(1,1.3))    
     
```

## save composite figure

```{r save composite figure to file}

#save to file
tiff(file=paste0(project_path,"/output/figures/figure_pupil_mmn_withintrial2.tiff"), # create a file in tiff format in current working directory
     width=12, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(
  arrangeGrob(gf_2+theme(legend.position = 'none'),
              gf_3+theme(legend.position = 'none'),
              mylegend_pupil,ncol=3,widths=c(5,5,2)), #pupil size change Grob 
  arrangeGrob(gf_1+theme(legend.position = 'none'), #MMN Grob
              arrangeGrob(
                gf_4a + theme(legend.position="none"),
                gf_4b + theme(legend.position="none"),
                gf_4c + theme(legend.position="none"),
                gf_4d + theme(legend.position="none"),nrow=2),
              mylegend, ncol=3, widths=c(5,5,2)),
  nrow=2)

dev.off() #close operation and save file


```


# task effects

investigates influence of stimulus and task progression on pupillary measures (BPS, SEPR) and EEG measure (MMN)
in separate linear mixed models and supported by Bayesian posterior sampling

## task condition / stimulus effect

```{r stimulus_on_BPS}

##BPS
## include trial characteristics - TRIAL CONDITION
lmm<-lmer(scale(pd_baseline)~EventData+
            (1|subjects)+(1|wave),data=df_trial)

anova(lmm)
emmeans(lmm,~EventData)
confint(contrast(emmeans(lmm,~EventData),'pairwise'))
###--> 201 > 202 + 204; 203 > 202

      ###restricted to trial that are preceded by standards
      lmm_r<-lmer(scale(pd_baseline)~EventData+
            (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position!=1,])

      anova(lmm_r)
      emmeans(lmm_r,~EventData)
      confint(contrast(emmeans(lmm_r,~EventData),'pairwise'))



r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(pd_baseline)~EventData+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

# null_lmm_ML <- lmer(scale(pd_baseline)~1+
#                    (1|subjects)+(1|wave),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  
 
```

```{r stimulus_on_BPS_Bayesian}

    # data - create greta arrays
    y <- as_data(scale(df_trial$pd_baseline))
    events <- as_data(model.matrix(~ EventData - 1, df_trial)) #alternative to dummy coding?
    subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)

    # variables and priors
    int <- normal(0, 1)
    event_coef<- normal(0, 1, dim=ncol(events))
    sd <- student(3, 0, 1, truncation = c(0, Inf))
    a_subject <- normal(0, 1, dim = max(subject_id))

    # operations
    mean <- events %*% event_coef + a_subject[subject_id]

    # likelihood - distribution over data
    distribution(y) <- normal(mean, sd)

    # defining the model
    m <- model(event_coef)

    # sampling
    draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups)
    summary(draws)
    
    mcmc_intervals(draws, prob= .66, prob_outer = .95)
    
    mcmc_trace(draws) #see whether chains converge and are stationary
    mcmc_dens(draws) # + stable estimates

    #posterior predictive checking
    coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
    coda::effectiveSize(draws) #ESS

    #save model
    draws_BPS_trialcondition<-draws
    save(draws_BPS_trialcondition,file=paste0(project_path,'/data/Bayesian_model_BPS_trialcondition.Rdata'))

```

```{r stimulus_on_SEPR}

##SEPR
  
## include trial characteristics - TRIAL CONDITION
lmm<-lmer(scale(rpd_auc)~EventData+
            (1|subjects)+(1|wave),data=df_trial)

anova(lmm)
emmeans(lmm,~EventData)
confint(contrast(emmeans(lmm,~EventData),'pairwise'))
###--> 201 > 202 + 204; 203 > 202

r2_nakagawa(lmm)  
  
#BIC Bayes Factor  
lmm_ML<-lmer(scale(rpd_auc)~EventData+
          (1|subjects)+(1|wave),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(rpd_auc)~1+
                 (1|subjects)+(1|wave),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
BF_BIC
  
```

```{r stimulus_on_SEPR_Bayesian}

    # data - create greta arrays
    y <- as_data(scale(df_trial$rpd_auc))
    events <- as_data(model.matrix(~ EventData - 1, df_trial)) #alternative to dummy coding?
    subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)

    # variables and priors
    int <- normal(0, 1)
    event_coef<- normal(0, 1, dim=ncol(events))
    sd <- student(3, 0, 1, truncation = c(0, Inf))
    a_subject <- normal(0, 1, dim = max(subject_id))

    # operations
    mean <- int + events %*% event_coef + a_subject[subject_id]

    # likelihood - distribution over data
    distribution(y) <- normal(mean, sd)

    # defining the model
    m <- model(event_coef)

    # sampling
    draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups)
    summary(draws)
    
    mcmc_intervals(draws, prob= .66, prob_outer = .95)+scale_y_discrete(labels=c('standard','pitch oddball','length oddball','pitch & length oddball'))+coord_flip()+theme_bw()
    
    mcmc_trace(draws) #see whether chains converge and are stationary
    mcmc_dens(draws) # + stable estimates

    
    #posterior predictive checking
    coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
    coda::effectiveSize(draws) #ESS

    #save model
    draws_SEPR_taskcondition<-draws
    save(draws_SEPR_taskcondition,file=paste0(project_path,'/data/Bayesian_model_SEPR_taskcondition.Rdata'))
        
```

```{r stimulus_on_MMN}

## include trial characteristics - TRIAL CONDITION
lmm<-lmer(scale(mmn)~EventData+
            (1|id),data=df_trial)

anova(lmm)
emmeans(lmm,~EventData)
confint(contrast(emmeans(lmm,~EventData),'pairwise'))

r2_nakagawa(lmm)  
  
#BIC Bayes Factor  
lmm_ML<-lmer(scale(mmn)~EventData+
          (1|id),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(mmn)~1+
                 (1|id),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
BF_BIC
  
```

## task progression effect

```{r task_progression_on_BPS}
  
## include trial characteristics - TASK PROGRESSION
  lmm<-lmer(scale(pd_baseline)~stats::poly(EventCounter,3)+
              (1|subjects)+(1|wave),data=df_trial)
  
  anova(lmm)
  
  cbind(round(fixef(lmm)['EventCounter'],9),
                round(confint(lmm,parm = 'EventCounter'),9))

  confint(contrast(emmeans(lmm,~EventCounter,at=list(EventCounter =     seq(1,1400,699))),'revpairwise'))

r2_nakagawa(lmm)  
 

#BIC Bayes Factor
lmm_ML<-lmer(scale(pd_baseline)~stats::poly(EventCounter,3)+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(pd_baseline)~1+
                   (1|subjects)+(1|wave),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  


```

```{r task progression on BPS - Bayesian}

 # data - create greta arrays
    y <- as_data(scale(df_trial$pd_baseline))
    trials <- as_data(scale(df_trial$EventCounter))
    subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)

    
    # variables and priors
    int <- normal(0, 1)
    coef <- normal(0, 1)
    sd <- student(3, 0, 1, truncation = c(0, Inf))
    a_subject <- normal(0, 1, dim = max(subject_id))

    # operations
    mean <- int + coef * trials + a_subject[subject_id]

    # likelihood - distribution over data
    distribution(y) <- normal(mean, sd)

    # defining the model
    m <- model(int, coef, sd)

    # sampling
    draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups)
    summary(draws)
    
    #diangostics
    mcmc_trace(draws) #see whether chains converge and are stationary
    mcmc_dens(draws) # + stable estimates

    #posterior predictive checking
    coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
    coda::effectiveSize(draws) #ESS

    #save model
    draws_BPS_taskprogression<-draws
    save(draws_BPS_taskprogression,file=paste0(project_path,'/data/Bayesian_model_BPS_taskprogression.Rdata'))


```

```{r task progression on SEPR}

## include trial characteristics - TASK PROGRESSION
  lmm<-lmer(scale(rpd_auc)~EventCounter+
              (1|subjects)+(1|wave),data=df_trial)
  anova(lmm)
  
  cbind(round(fixef(lmm)['EventCounter'],9),
                round(confint(lmm,parm = 'EventCounter'),9))

  confint(contrast(emmeans(lmm,~EventCounter,at=list(EventCounter =   seq(1,1400,1399))),'pairwise'))
  
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(rpd_auc)~EventCounter+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(rpd_auc)~1+
                   (1|subjects)+(1|wave),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  


```

```{r task progression on SEPR - Bayesian}

 # data - create greta arrays
    y <- as_data(scale(df_trial$rpd_auc))
    trials <- as_data(scale(df_trial$EventCounter))
    subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)

    
    # variables and priors
    int <- normal(0, 1)
    coef <- normal(0, 1)
    sd <- student(3, 0, 1, truncation = c(0, Inf))
    a_subject <- normal(0, 1, dim = max(subject_id))

    # operations
    mean <- coef * trials + a_subject[subject_id]

    # likelihood - distribution over data
    distribution(y) <- normal(mean, sd)

    # defining the model
    m <- model(coef)

    # sampling
    draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups)
    summary(draws)
    
    mcmc_trace(draws) #see whether chains converge and are stationary
    mcmc_dens(draws) # + stable estimates

    #posterior predictive checking
    coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
    coda::effectiveSize(draws) #ESS

    #save model
    draws_SEPR_taskprogression<-draws
    save(draws_SEPR_taskprogression,file=paste0(project_path,'/data/Bayesian_model_SEPR_taskprogression.Rdata'))
    
```

```{r task progression on MMN}

## include trial characteristics - TASK PROGRESSION
  lmm<-lmer(scale(mmn)~EventCounter+
              (1|id),data=df_trial)
  anova(lmm)
  
  cbind(round(fixef(lmm)['EventCounter'],9),
                round(confint(lmm,parm = 'EventCounter'),9))

  confint(contrast(emmeans(lmm,~EventCounter,at=list(EventCounter =   seq(1,1400,1399))),'pairwise'))
  
r2_nakagawa(lmm)  
  
#BIC Bayes Factor
lmm_ML<-lmer(scale(mmn)~EventCounter+
              (1|id),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(mmn)~1+
                   (1|id),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  
BF_BIC  

```

create a combined figure of stimulus and task progression effects on BPS, SEPR, MMN

```{r figure_task_effects}

#y-axis limitation for combined plot
y_axis_limitation<-c(-0.1, 0.15)

#BPS
lmm<-lmer(scale(pd_baseline)~EventData+
              (1|subjects)+(1|wave),data=df_trial)

#create plot --> of contrasts
plot_task_effect<-plot(contrast(emmeans(lmm,~EventData),'pairwise'))[['data']]

#modify plot
g1<-ggplot(plot_task_effect,aes(x=contrast,y=the.emmean))+
  #conventional error bar
  geom_errorbar(aes(min=asymp.LCL,max=asymp.UCL),width=0.2)+
  
  # #overplot predicted values
  # geom_jitter(data=df_plot_predicted_bps[complete.cases(df_plot_predicted_bps),],
  #             aes(x=predicted_bps_eventdata,
  #                 y=predicted_bps_mean,color=predicted_bps_eventdata),alpha=0.3,width=0.4,show.legend=F)+
  #boxplot
  geom_boxplot(aes(fill=contrast,
                   middle=the.emmean,
                   lower=the.emmean-1.5*SE,
                   upper=the.emmean+1.5*SE,
                   ymin=asymp.LCL,
                   ymax=asymp.UCL),stat = "identity",alpha=0.7)+
  scale_fill_manual(values = custom_contrast_colors,labels = c('standard - pitch','standard - length','standard - pitch & length','pitch - length','pitch - pitch & length','length - pitch & length'))+
  scale_x_discrete(labels = NULL, breaks = NULL)+ #remove x-axis tick labels
  coord_cartesian(ylim = y_axis_limitation)+
  labs(x='contrast of stimuli',y='BPS difference (z)')
  

#SEPR -pupillary response - technical model
lmm<-lmer(scale(rpd_auc)~EventData+
              (1|subjects)+(1|wave),data=df_trial)

#create plot --> of contrasts
plot_task_effect<-plot(contrast(emmeans(lmm,~EventData),'pairwise'))[['data']]

#modify plot
g2<-ggplot(plot_task_effect,aes(x=contrast,y=the.emmean))+
  #conventional error bar
  geom_errorbar(aes(min=asymp.LCL,max=asymp.UCL),width=0.2)+
  
    #boxplot
  geom_boxplot(aes(fill=contrast,
                   middle=the.emmean,
                   lower=the.emmean-1.5*SE,
                   upper=the.emmean+1.5*SE,
                   ymin=asymp.LCL,
                   ymax=asymp.UCL),stat = "identity",alpha=0.7)+
  scale_fill_manual(values = custom_contrast_colors,labels = c('standard - pitch','standard - length','standard - pitch & length','pitch - length','pitch - pitch & length','length - pitch & length'))+
  scale_x_discrete(labels = NULL, breaks = NULL)+ #remove x-axis tick labels
  coord_cartesian(ylim = y_axis_limitation)+
  labs(x='contrast of stimuli',y='SEPR difference (z)')

##MMN
lmm<-lmer(scale(mmn)~EventData+(1|id),df_trial)

      #create plot --> of contrasts
      plot_task_effect<-plot(contrast(emmeans(lmm,~EventData),'pairwise'))[['data']]

      #modify plot
      g3<-ggplot(plot_task_effect,aes(x=contrast,y=the.emmean))+
        #conventional error bar
        geom_errorbar(aes(min=asymp.LCL,max=asymp.UCL),width=0.2)+
        geom_boxplot(aes(fill=contrast,
                         middle=the.emmean,
                         lower=the.emmean-1.5*SE,
                         upper=the.emmean+1.5*SE,
                         ymin=asymp.LCL,
                         ymax=asymp.UCL),stat = "identity",alpha=0.7)+
        scale_fill_manual(values = custom_contrast_colors,labels = c('standard - pitch','standard - length','standard - pitch & length','pitch - length','pitch - pitch & length','length - pitch & length'))+
        scale_x_discrete(labels = NULL, breaks = NULL)+ #remove x-axis tick labels
        coord_cartesian(ylim = y_axis_limitation)+
        labs(x='contrast of stimuli',y='MMN difference (z)')


###task progression BPS
BPS_progression<-lmer(scale(pd_baseline)~stats::poly(EventCounter,3)+
              (1|subjects)+(1|wave),data=df_trial)
      
model_plot_data_BPS<-as.data.frame(emmeans(BPS_progression,~EventCounter,at=list(EventCounter = seq(1,1400,140))))
    
g4<-ggplot(model_plot_data_BPS,aes(x=EventCounter,y=emmean))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE),alpha=0.8,position='dodge',fill='grey35')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  coord_cartesian(ylim = c(-0.2,0.25))+
  labs(x='task progression (trial)',y='baseline pupil size (BPS, z)')

###task progression SEPR
SEPR_progression<-lmer(scale(rpd_auc)~EventCounter+
              (1|subjects)+(1|wave),data=df_trial)
      
      
model_plot_data_SEPR<-as.data.frame(emmeans(SEPR_progression,~EventCounter,at=list(EventCounter = seq(1,1400,140))))
    
g5<-ggplot(model_plot_data_SEPR,aes(x=EventCounter,y=emmean))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE),alpha=0.8,position='dodge',fill='grey35')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  coord_cartesian(ylim = c(-0.2,0.25))+
  labs(x='task progression (trial)',y='pupillary response (SEPR, z)')

###task progression MMN
MMN_progression<-lmer(scale(mmn)~EventCounter+
              (1|id),data=df_trial)
      
model_plot_data_MMN<-as.data.frame(emmeans(MMN_progression,~EventCounter,at=list(EventCounter = seq(1,1400,140))))
    
g6<-ggplot(model_plot_data_MMN,aes(x=EventCounter,y=emmean))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE),alpha=0.8,position='dodge',fill='grey35')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  coord_cartesian(ylim = c(-0.2,0.25))+
  labs(x='task progression (trial)',y='mismatch negativity (MMN, z)')

#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend_contrast<-g_legend(g3)

      
###composite figure layout
lay <- rbind(c(1,2,3,7),
             c(4,5,6,7))


#plot in markdown
grid.arrange(g1+theme(legend.position="none")+xlab(''),
             g2+theme(legend.position="none"),
             g3+theme(legend.position="none")+xlab(''),
             g4+xlab(''),
             g5,
             g6+xlab(''),
             mylegend_contrast,layout_matrix=lay)

#save to file
tiff(file=paste0(project_path,"/output/figures/task_effects_contrasts_progression.tiff"), # create a file in tiff format in 
  width=8, height=5, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(g1+theme(legend.position="none")+xlab(''),
             g2+theme(legend.position="none"),
             g3+theme(legend.position="none")+xlab(''),
             g4+xlab(''),
             g5,
             g6+xlab(''),
             mylegend_contrast,layout_matrix=lay)
      

dev.off() #close operation and save file


```

# group differences

## baseline pupil size (BPS)

```{r BPS task progression between groups - LMM}

#main model
lmm_REML<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
              (1|subjects)+(1|wave),data=df_trial)

#fixed effect significances
anova(lmm_REML)

#model fit
r2_nakagawa(lmm_REML)

#POST-HOC
contrast(emmeans(lmm_REML,~EventData|t1_diagnosis),'pairwise')
##--> lower BPS in ASD for length versus standards

#Post-hoc - between group
emtrends(lmm_REML,~t1_diagnosis,var='EventCounter')

#emmeans(lmm_REML,~t1_diagnosis|EventCounter,at=list(EventCounter = seq(1,1400,1399)))
emmeans(lmm_REML,~t1_diagnosis|EventCounter,at=list(EventCounter = seq(1,1400,699)))


  confint(contrast(emmeans(lmm_REML,~EventCounter|t1_diagnosis,at=list(EventCounter =   seq(1,1400,699))),'revpairwise'))


#Bayes factor - group x task progression
lmm_ML<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(pd_baseline)~EventData+t1_diagnosis+stats::poly(EventCounter,3)+
                   EventData:EventCounter+EventData:t1_diagnosis+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
BF_BIC

### compare full model for influence of covariates

#full model
lmm_full<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*EventCounter+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|subjects)+(1|wave),data=df_trial)
              #--> group effect (interaction) is robust in reduced and full model

anova(lmm_full)

  
```

```{r BPS task progression between groups - Bayesian}

####--> reun with more iterations

# A. translate data - make greta arrays
y <- as_data(scale(df_trial$pd_baseline))
design <- as.matrix(cbind(ifelse(df_trial$t1_diagnosis=='ASD',T,F),df_trial$EventCounter))
subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)

#define priors
int <- normal(0, 1)
design_coef<- normal(0, 1, dim=ncol(design))
sd <- cauchy(0, 3, truncation = c(0, Inf))
#sd <- student(3, 0, 1, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- int + design %*% design_coef + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
m <- model(design_coef)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups,) ###--> test
Sys.time() - start

#save model
draws_BPS_twoway<-draws
save(draws_BPS_twoway,file=paste0(project_path,'/data/Bayesian_model_BPS_twoway.Rdata'))

#posterior estimates
point_estimate(draws_BPS_twoway)
ci(draws_BPS_twoway,ci=credible_interval)

```

```{r BPS habituation between groups - LMM}

  with(df_trial,table(sequence_position,EventData))
  ###--> weird distribution?

  #recoding
  df_trial$sequence_position_z<-scale(df_trial$sequence_position)

          #figure
        df_trial$standardtrial<-with(df_trial,ifelse(EventData==201,T,F))
        ggplot(df_trial[df_trial$sequence_position<=10,],
               aes(x=sequence_position,y=scale(pd_baseline),group=interaction(standardtrial,t1_diagnosis),
                   color=standardtrial,linetype=t1_diagnosis))+
          geom_smooth()+labs(x='sequence position',y='BPS (z)')+scale_x_continuous(breaks=1:10)+
          theme_bw()
        
        ggplot(df_trial[df_trial$sequence_position<=10,],
               aes(x=sequence_position,y=scale(rpd_response),group=interaction(standardtrial,t1_diagnosis),
                   color=t1_diagnosis,linetype=standardtrial))+
          geom_smooth()+labs(x='sequence position',y='SEPR (z)')+scale_x_continuous(breaks=1:10)+
          theme_bw()


  #BPS - continuous
  lmm<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*stats::poly(sequence_position_z,2)+
              (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<10,])

  #model fit
  anova(lmm)
  
        ### supplementary analysis
        lmm<-lmer(scale(pd_baseline)~t1_diagnosis*sequence_position_z+
                    (1|subjects)+(1|wave),data=df_trial[df_trial$EventData==201 & df_trial$sequence_position<10,])

        #model fit
        anova(lmm)
        
        r2_nakagawa(lmm)



        #BIC Bayes Factor
        lmm_ML<-lmer(scale(pd_baseline)~t1_diagnosis*sequence_position_z+
                       (1|subjects)+(1|wave),data=df_trial[df_trial$EventData==201 & df_trial$sequence_position<10,],REML=F)

        null_lmm_ML <- lmer(scale(pd_baseline)~sequence_position_z+
                              (1|subjects)+(1|wave),data=df_trial[df_trial$EventData==201 & df_trial$sequence_position<10,],REML=F)

        BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)
        BF_BIC

  #BPS categorical
  lmm<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*as.factor(sequence_position)+
              (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<4,])

  anova(lmm)
  contrast(emmeans(lmm,~as.factor(sequence_position)),'pairwise')
  ###-> 1-3 effect
  ###inclusion of covariates leads to boundary singular

  lmm<-lmer(scale(pd_baseline)~t1_diagnosis*as.factor(sequence_position)+
              (1|subjects)+(1|wave),data=df_trial[df_trial$EventData==201 & df_trial$sequence_position<4,])

```

```{r BPS habituation between groups - Bayesian}


df_bps_bayesian_habituation<-df_trial[df_trial$EventData==201 & df_trial$sequence_position<10,]

# A. translate data - make greta arrays
y <- as_data(scale(df_bps_bayesian_habituation$pd_baseline))
design <- as.matrix(cbind(ifelse(df_bps_bayesian_habituation$t1_diagnosis=='ASD',T,F),df_bps_bayesian_habituation$sequence_position))
subject_id <- as.integer(factor(df_bps_bayesian_habituation$subjects)) #random intercept (in LMM)

#define priors
int <- normal(0, 1)
design_coef<- normal(0, 1, dim=ncol(design))
sd <- cauchy(0, 3, truncation = c(0, Inf))
#sd <- student(3, 0, 1, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- int + design %*% design_coef + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
m <- model(design_coef)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups,) ###--> test
Sys.time() - start

#save model
draws_BPS_supp<-draws
save(draws_BPS_supp,file=paste0(project_path,'/data/Bayesian_model_BPS_supp.Rdata'))

#posterior estimates
point_estimate(draws_BPS_supp)
ci(draws_BPS_supp,ci=credible_interval)


```

## stimulus-evoked pupillary response

```{r SEPR task progression between groups LMM}

#MAIN MODEL - pupillary response by group - random intercept
lmm_REML<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+
            (1|subjects)+(1|wave),data=df_trial)
            #--> group effect (interaction) is robust in reduced and full model

#fixed effect significances
anova(lmm_REML)

#Bayes factor - group x task progression
lmm_ML<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+
                (1|subjects)+(1|wave),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(rpd_auc)~EventData+EventCounter+
                 EventData:EventCounter+EventData:t1_diagnosis:EventCounter+
                (1|subjects)+(1|wave),data=df_trial,REML=F)

  BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
  BF_BIC


#full model
lmm_full<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|subjects)+(1|wave),data=df_trial)
              #--> group effect (interaction) is robust in reduced and full model

anova(lmm_full)

confint(contrast(emmeans(lmm_full,~EventData|t1_diagnosis),'revpairwise'))

#coefficient of determination
r2_nakagawa(lmm_full)


#post hoc effects
confint(contrast(emtrends(lmm_REML,~t1_diagnosis|EventData,var='EventCounter'),'pairwise'))
      plot(emtrends(lmm_REML,~t1_diagnosis|EventData,var='EventCounter'))

confint(contrast(emmeans(lmm_REML,~EventCounter|t1_diagnosis+EventData,at=list(EventCounter =   seq(1,1400,1399))),'revpairwise'))

      

```

```{r SEPR_three_way_interaction Bayesian}

# A. translate data - make greta arrays
y <- as_data(scale(df_trial$rpd_auc))
trials <- as_data(scale(df_trial$EventCounter))
subject_id <- as.integer(factor(df_trial$subjects)) #random intercept (in LMM)
events <- as_data(model.matrix(~ EventData - 1, df_trial)) #alternative to dummy coding?
interaction_variable<-with(df_trial,interaction(t1_diagnosis,EventData))
group_by_events <- as_data(model.matrix(~interaction_variable-1, df_trial)) #alternative to dummy coding - remove intercept (-1)
#table(group_by_events[,1],group_by_events[,9])

#define priors
event_coefs<- normal(0, 1, dim=ncol(events))
group_event_coefs <- normal(0,1, dim=ncol(group_by_events))

sd <- cauchy(0, 3, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- trials * group_by_events %*% group_event_coefs + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
#m <- model(int, coef_201, coef_202, coef_203, coef_204, sd)
m <- model(group_event_coefs)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups) ###--> test
Sys.time() - start

#save model
draws_SEPR_threeway<-draws
save(draws_SEPR_threeway,file=paste0(project_path,'/data/Bayesian_model_SEPR_threewayinteraction.Rdata'))

#posterior estimates
table(interaction_variable) #names of coefficients
point_estimate(draws_SEPR_threeway)
ci(draws_SEPR_threeway,ci=credible_interval)


```

```{r SEPR_habituation_LMM}

  #SEPR - continuous habituation
  lmm<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*sequence_position_z+
              (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<10,])

  #fixed effect significance
  anova(lmm)
  
  #model fit
  r2_nakagawa(lmm)
  
        ##with covariates
        lmm_full<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*sequence_position_z+
                      scale(ageyrs)+scale(t1_piq)+sex+
                       scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
                    (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<10,])
          
        anova(lmm_full)
        ###--> if an oddballs follows an oddball - higehr response in ASD
        ###--> if oddball follows a standard - higher response in TD
        

  #SEPR - categorical habituation
  lmm<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*as.factor(sequence_position)+
              (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<4,])

   anova(lmm)
       

      #BIC Bayes Factor
      lmm_ML<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*as.factor(sequence_position)+
                     (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<4,],REML=F)

      null_lmm_ML <- lmer(scale(rpd_auc)~1+
                            (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<4,],REML=F)

      BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)
      BF_BIC

      BIC(null_lmm_ML)
      BIC(lmm_ML)

  
  #full model with covariates
  lmm_full<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*as.factor(sequence_position)+
                 scale(ageyrs)+scale(t1_piq)+sex+
                 scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
                (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<4,])
                #--> group effect (interaction) is robust in reduced and full model

  anova(lmm_full)
  
  contrast(emmeans(lmm,~t1_diagnosis|EventData+as.factor(sequence_position)),'pairwise')
  confint(contrast(emmeans(lmm,~t1_diagnosis|EventData+as.factor(sequence_position)),'revpairwise'))

```

```{r SEPR_habituation_Bayesian}

df_sepr_bayesian_habituation<-df_trial[df_trial$sequence_position<4,]

# A. translate data - make greta arrays
y <- as_data(scale(df_sepr_bayesian_habituation$rpd_auc))
interaction_variable<-with(df_sepr_bayesian_habituation,interaction(t1_diagnosis,sequence_position))
group_by_events <- as_data(model.matrix(~interaction_variable-1, df_sepr_bayesian_habituation)) #alternative to dummy coding - remove intercept (-1)
subject_id <- as.integer(factor(df_sepr_bayesian_habituation$subjects)) #random intercept (in LMM)

#define priors
int <- normal(0, 1)
design_coef<- normal(0, 1, dim=ncol(group_by_events))
sd <- cauchy(0, 3, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- int + group_by_events %*% design_coef + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
m <- model(design_coef)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups,) ###--> test
Sys.time() - start


#save model
draws_SEPR_habituation<-draws
save(draws_SEPR_habituation,file=paste0(project_path,'/data/Bayesian_model_SEPR_habituation.Rdata'))

#posterior estimates
table(interaction_variable) #names of coefficients
point_estimate(draws_SEPR_habituation)
ci(draws_SEPR_habituation,ci=credible_interval)



```

## mismatch negativity between groups

```{r MMN task progression - LLM}

#MAIN MODEL - MN by group - random intercept
lmm_REML<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+
            (1|id),data=df_trial)

#fixed effect significances
anova(lmm_REML)

#Bayes factor - group x task progression
lmm_ML<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+
                (1|id),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(mmn)~EventData+EventCounter+
                 EventData:EventCounter+EventData:t1_diagnosis:EventCounter+
                (1|id),data=df_trial,REML=F)

  BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
  BF_BIC


#full model
lmm_full<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|id),data=df_trial)
              #--> group effect (interaction) is robust in reduced and full model

anova(lmm_full)

#coefficient of determination
r2_nakagawa(lmm_full)


confint(contrast(emtrends(lmm_REML,~t1_diagnosis|EventData,var='EventCounter'),'pairwise'))

      plot(emtrends(lmm_REML,~t1_diagnosis|EventData,var='EventCounter'))

      
##translate to changes across the task
confint(contrast(emmeans(lmm_REML,~EventCounter|t1_diagnosis+EventData,at=list(EventCounter =   seq(1,1400,1399))),'revpairwise'))



```

```{r MMN task progression - Bayesian}

#remove NA
df_bayesian_mmn<-df_trial[!is.na(df_trial$mmn),]

# A. translate data - make greta arrays
y <- as_data(scale(df_bayesian_mmn$mmn))
trials <- as_data(scale(df_bayesian_mmn$EventCounter))
subject_id <- as.integer(factor(df_bayesian_mmn$subjects)) #random intercept (in LMM)
events <- as_data(model.matrix(~ EventData - 1, df_bayesian_mmn)) #alternative to dummy coding?
interaction_variable<-with(df_bayesian_mmn,interaction(t1_diagnosis,EventData))
group_by_events <- as_data(model.matrix(~interaction_variable-1, df_bayesian_mmn)) #alternative to dummy coding - remove intercept (-1)
#table(group_by_events[,1],group_by_events[,9])

#define priors
event_coefs<- normal(0, 1, dim=ncol(events))
group_event_coefs <- normal(0,1, dim=ncol(group_by_events))

sd <- cauchy(0, 3, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- trials * group_by_events %*% group_event_coefs + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
#m <- model(int, coef_201, coef_202, coef_203, coef_204, sd)
m <- model(group_event_coefs)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups) ###--> test
Sys.time() - start


#estimator convergence
mcmc_trace(draws) #see whether chains converge and are stationary
mcmc_dens(draws) # + stable estimates

##results
mcmc_areas(draws, prob= .66, prob_outer = .95)+scale_y_discrete(labels=c('ASD-standard','TD-standard ','ASD-pitch','TD-pitch','ASD-length ','TD-length','ASD-pitch+length','TD-pitch+length'))+coord_flip()+theme_bw()

#analyse
summary(draws)


#posterior predictive checking
coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
coda::effectiveSize(draws) #ESS - the higher the better

#save model
draws_MMN_threeway<-draws
save(draws_MMN_threeway,file=paste0(project_path,'/data/Bayesian_model_MMN_threewayinteraction.Rdata'))

#posterior estimate
table(interaction_variable) #names of coefficients
point_estimate(draws_MMN_threeway)
ci(draws_MMN_threeway,ci=credible_interval)


```

```{r MMN habituation LMM}

#MMN - continuous habituation
lmm_REML<-lmer(scale(mmn)~EventData*t1_diagnosis*sequence_position_z+
            (1|id),data=df_trial[df_trial$sequence_position<10,])

anova(lmm_REML)

#coefficient of determination
r2_nakagawa(lmm_REML)


#Bayes factor - group x sequence position
lmm_ML<-lmer(scale(mmn)~EventData*t1_diagnosis*sequence_position_z+
                (1|id),data=df_trial,REML=F)

null_lmm_ML <- lmer(scale(mmn)~EventData+EventCounter+
                 EventData:EventCounter+EventData:t1_diagnosis:EventCounter+
                (1|id),data=df_trial,REML=F)

  BF_BIC <- exp((BIC(null_lmm_ML) - BIC(lmm_ML))/2)  # BICs to Bayes factor
  BF_BIC


  
#full model
lmm_full<-lmer(scale(mmn)~EventData*t1_diagnosis*sequence_position_z+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|id),data=df_trial[df_trial$sequence_position<10,])
              


anova(lmm_full)

confint(contrast(emmeans(lmm_REML,~EventData|t1_diagnosis),'pairwise'))




#MMN - categorical habituation
lmm_REML<-lmer(scale(mmn)~EventData*t1_diagnosis*as.factor(sequence_position)+
            (1|id),data=df_trial[df_trial$sequence_position<4,])

anova(lmm_REML)
  


```

```{r MMN habituation - Bayesian}

#remove NA
df_mmn_bayesian_habituation<-df_trial[!is.na(df_trial$mmn) & df_trial$sequence_position<10,]

# A. translate data - make greta arrays
y <- as_data(scale(df_mmn_bayesian_habituation$mmn))
subject_id <- as.integer(factor(df_mmn_bayesian_habituation$subjects)) #random intercept (in LMM)
events <- as_data(model.matrix(~ EventData - 1, df_trial)) #alternative to dummy coding?
interaction_variable<-with(df_mmn_bayesian_habituation,interaction(t1_diagnosis,EventData))
group_by_events <- as_data(model.matrix(~interaction_variable-1, df_mmn_bayesian_habituation)) #alternative to dummy coding - remove intercept (-1)
#table(group_by_events[,1],group_by_events[,9])

#define priors
event_coefs<- normal(0, 1, dim=ncol(events))
group_event_coefs <- normal(0,1, dim=ncol(group_by_events))

sd <- cauchy(0, 3, truncation = c(0, Inf))
a_subject <- normal(0, 1, dim = max(subject_id))

#OPERATIONS - MODEL
mean <- group_by_events %*% group_event_coefs + a_subject[subject_id]

# likelihood - distribution over data
distribution(y) <- normal(mean, sd)

# defining the model
#m <- model(int, coef_201, coef_202, coef_203, coef_204, sd)
m <- model(group_event_coefs)

# sampling
start <-Sys.time()
draws <- mcmc(m, n_samples = number_of_iterations, warmup = number_of_warmups) ###--> test
Sys.time() - start

# #estimator convergence
# mcmc_trace(draws) #see whether chains converge and are stationary
# mcmc_dens(draws) # + stable estimates
# 
# ##results
# mcmc_areas(draws, prob= .66, prob_outer = .89)+scale_y_discrete(labels=c('ASD-standard','TD-standard ','ASD-pitch','TD-pitch','ASD-length ','TD-length','ASD-pitch+length','TD-pitch+length'))+coord_flip()+theme_bw()
# 
# #analyse
# summary(draws)
# 
# #posterior predictive checking
# coda::gelman.diag(draws) #Rhat <= 1.01 is cutoff for reliable estimator
# coda::effectiveSize(draws) #ESS - the higher the better

#save model
draws_MMN_habituation<-draws
save(draws_MMN_habituation,file=paste0(project_path,'/data/Bayesian_model_MMN_habituation_continuous.Rdata'))

#posterior estimate
table(interaction_variable) #names of coefficients
point_estimate(draws_MMN_habituation)
ci(draws_MMN_habituation,ci=credible_interval)


```



## computational modelling - PCDM - neural gain

```{r PCDM_neural_gain_analysis}


#data quality for neural gain
with(df_timepoint,table(Rsq_pcdm>=0.2,t1_diagnosis))
chisq.test(with(df_timepoint,table(Rsq_pcdm>=0.2,t1_diagnosis)))

#higher gain for standards in ASD
lm_gain<-lm(scale(gain_pcdm)~t1_diagnosis,df_timepoint[df_timepoint$Rsq_pcdm>0.2,])
summary(lm_gain) #higher gain in ASD for standards
anova(lm_gain) #higher gain in ASD for standards

t.test(scale(gain_pcdm)~t1_diagnosis,df_timepoint[df_timepoint$Rsq_pcdm>0.2,])
cohens_d(x=scale(gain_pcdm)~t1_diagnosis,data=df_timepoint[df_timepoint$Rsq_pcdm>0.2,])
#--> d=0.4 in ASD

#visualize
ggplot(df_timepoint[df_timepoint$Rsq_pcdm>0.2, ],aes(x=scale(gain_pcdm),fill=t1_diagnosis))+geom_histogram(aes(y = after_stat(density)),position='dodge',bins=20)+theme_bw()

###scaling with demographic data
summary(lm(scale(gain_pcdm)~ageyrs,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))
summary(lm(scale(gain_pcdm)~t1_piq,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))
summary(lm(scale(gain_pcdm)~sex,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))

# summary(lm(scale(gain_pcdm)~rpd_auc.201,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))
# summary(lm(scale(gain_pcdm)~rpd_auc.202,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))
# summary(lm(scale(gain_pcdm)~rpd_auc.203,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))
# summary(lm(scale(gain_pcdm)~rpd_auc.204,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))
# summary(lm(scale(gain_pcdm)~pd_baseline,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))

#not altered by covariates
lmm_full<-lm(scale(gain_pcdm)~t1_diagnosis+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)
               ,df_timepoint[df_timepoint$Rsq_pcdm>0.2,])

anova(lmm_full) #higher gain in ASD for standards



```

## figure group differences LMM

added to a compostie figure - not evaluated

```{r figure_BPS_SEP_MMN_between_groups, , eval=FALSE}
#main models
BPS_progression<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
              (1|subjects)+(1|wave),data=df_trial)

BPS_habituation<-lmer(scale(pd_baseline)~t1_diagnosis*sequence_position+
                    (1|subjects)+(1|wave),data=df_trial[df_trial$EventData==201 & df_trial$sequence_position<=10,])

SEPR_progression<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+
              (1|subjects)+(1|wave),data=df_trial)

MMN_progression<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+
              (1|id),data=df_trial)
  

#plot interaction
model_plot_data_BPS1<-as.data.frame(emmeans(BPS_progression,~t1_diagnosis+EventCounter,
                                       at=list(EventCounter = seq(1,1400,140))))

g1<-ggplot(model_plot_data_BPS1,aes(x=EventCounter,y=emmean,group=t1_diagnosis,color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  labs(x='task progression (trial)')+
  theme(axis.title.y=element_blank(),legend.position = 'none')

model_plot_data_BPS2<-as.data.frame(emmeans(BPS_habituation,~t1_diagnosis+sequence_position,
                                       at=list(sequence_position = seq(1,10))))

g2<-ggplot(model_plot_data_BPS2,aes(x=sequence_position,y=emmean,group=t1_diagnosis,color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_x_continuous(breaks=seq(0,10,1))+
  labs(x='sequence position')+
  theme(axis.title.y=element_blank(),legend.position = 'none')


model_plot_data_SEPR<-as.data.frame(emmeans(SEPR_progression,~t1_diagnosis+EventData+EventCounter,at=list(EventCounter = seq(1,1400,140))))
    
    labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch & length oddball")
g3<-ggplot(model_plot_data_SEPR,aes(x=EventCounter,y=emmean,
                           group=interaction(EventData,t1_diagnosis),color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  facet_wrap(~EventData,labeller=labeller(EventData = labels))+
  labs(x='task progression (trial)',y='stimulus-evoked pupillary response - SEPR (z)')


model_plot_data_MMN<-as.data.frame(emmeans(MMN_progression,~t1_diagnosis+EventData+EventCounter,at=list(EventCounter = seq(1,1400,140))))
    
    labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch & length oddball")
g4<-ggplot(model_plot_data_MMN,aes(x=EventCounter,y=emmean,
                           group=interaction(EventData,t1_diagnosis),color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  facet_wrap(~EventData,labeller=labeller(EventData = labels))+
  labs(x='task progression (trial)',y='mismatch negativity - MMN (z)')

  
lay <- rbind(c(1,1,3,3,3,5),
             c(1,1,3,3,3,5),
             c(1,1,3,3,3,5),
             c(2,2,4,4,4,5),
             c(2,2,4,4,4,5),
             c(2,2,4,4,4,5))


g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend_group<-g_legend(g4)


#plot in markdown
grid.arrange(g1,g2,
             g3+theme(legend.position = 'none'),
             g4+theme(legend.position = 'none'),
             mylegend_group, left='baseline pupil size (z)', layout_matrix=lay)


#save to file
tiff(file=paste0(project_path,"/output/figures/group_differences.tiff"), # create a file in tiff format in current working directory
  width=9, height=7, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure
      
grid.arrange(g1,g2,
             g3+theme(legend.position = 'none'),
             g4+theme(legend.position = 'none'),
             mylegend_group, left='baseline pupil size (z)', layout_matrix=lay)

dev.off() #close operation and save file




```

## figure group differences Bayesian

added to a compostie figure - not evaluated

```{r figure group differences Bayesian, eval=FALSE}

#load Bayesian models
load(file=paste0(project_path,'/data/Bayesian_model_BPS_twoway.Rdata')) #draws_BPS_twoway
load(file=paste0(project_path,'/data/Bayesian_model_BPS_supp.Rdata')) #draws_BPS_supp
load(file=paste0(project_path,'/data/Bayesian_model_SEPR_threewayinteraction.Rdata')) #draws_SEPR_threeway
load(file=paste0(project_path,'/data/Bayesian_model_MMN_threewayinteraction.Rdata')) #draws_MMN_threeway

# #posterior predictive checking
# coda::gelman.diag(draws_BPS_twoway) #Rhat <= 1.01 is cutoff for reliable estimator
# coda::effectiveSize(draws_BPS_twoway) #ESS
# 
# coda::gelman.diag(draws_BPS_supp) #Rhat <= 1.01 is cutoff for reliable estimator
# coda::effectiveSize(draws_BPS_supp) #ESS
# 
# coda::gelman.diag(draws_SEPR_threeway) #Rhat <= 1.01 is cutoff for reliable estimator
# coda::effectiveSize(draws_SEPR_threeway) #ESS
# 
# coda::gelman.diag(draws_MMN_threeway) #Rhat <= 1.01 is cutoff for reliable estimator
# coda::effectiveSize(draws_MMN_threeway) #ESS

#posterior estimates
# point_estimate(draws_BPS_twoway)
# ci(draws_BPS_twoway,ci=credible_interval)
# 
# point_estimate(draws_BPS_supp)
# ci(draws_BPS_supp,ci=credible_interval)
# 
# point_estimate(draws_SEPR_threeway)
# ci(draws_SEPR_threeway,ci=credible_interval)
# 
# point_estimate(draws_MMN_threeway)
# ci(draws_MMN_threeway,ci=credible_interval)


#draws_BPS_twoway
plot_data<-mcmc_areas_data(draws_BPS_twoway,prob_outer = credible_interval)
g1a<-ggplot(plot_data[plot_data$parameter=="design_coef[1,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('AUT'))+
  labs(y='posterior sampling: BPS - task progression',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[1],
                    labels=c('autistic'))
g1b<-ggplot(plot_data[plot_data$parameter=="design_coef[2,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('NON-AUT'))+
  labs(y='',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[2],
                    labels=c('non-autistic'))
g1<-grid.arrange(g1a,g1b,ncol=2)

#draws_BPS_supp
plot_data<-mcmc_areas_data(draws_BPS_supp,prob_outer = credible_interval)
g2a<-ggplot(plot_data[plot_data$parameter=="design_coef[1,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('AUT'))+
  labs(y='posterior sampling: BPS - sequence position',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[1],
                    labels=c('autistic'))
g2b<-ggplot(plot_data[plot_data$parameter=="design_coef[2,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('NON-AUT'))+
  labs(y='',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[2],
                    labels=c('non-autistic'))
g2<-grid.arrange(g2a,g2b,ncol=2)


#draws_SEPR_threeway
plot_data<-mcmc_areas_data(draws_SEPR_threeway,prob_outer = credible_interval)
g3<-ggplot(plot_data,aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75),scale='width')+
  scale_x_discrete(labels=rep(c('AUT','NON-AUT'),4))+
  labs(y='posterior sampling: SEPR trend',x='')+
  scale_fill_manual(name = 'parameter: group x stimulus',
                    values = rep(custom_condition_colors,each=2),
                    labels=c('autistic - standard trial','non-autistic - standard trial',
                            'autistic - pitch oddball','non-autistic - pitch oddball',
                            'autistic - length oddball','non-autistic - length oddball',
                            'autistic - pitch+length oddball','non-autistic - pitch+length oddball'))
  

#draws_MMN_threeway 
plot_data<-mcmc_areas_data(draws_MMN_threeway,prob_outer = credible_interval)
g4<-ggplot(plot_data,aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75),scale='width')+
  scale_x_discrete(labels=rep(c('AUT','NON-AUT'),4))+
  labs(y='posterior sampling: MMN trend',x='')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = rep(custom_condition_colors,each=2),
                    labels=c('standard trial','standard trial',
                            'pitch oddball','pitch oddball',
                            'length oddball','length oddball',
                            'pitch+length oddball','pitch+length oddball'))

###extract legends
#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}


plot_data<-mcmc_areas_data(draws_BPS_supp,prob_outer = credible_interval)
g2_legend<-ggplot(plot_data,aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('AUT','NON-AUT'))+
  labs(y='posterior sampling: BPS - sequence position',x='')+
  scale_fill_manual(name = 'parameter: group',
                    values = custom_group_color,
                    labels=c('autistic','non-autistic'))

mylegend_group<-g_legend(g2_legend)
mylegend_stimulus<-g_legend(g3)


grid.arrange(g1,
             g2,
             mylegend_group,
             g3+theme(legend.position = 'none'),
             g4+theme(legend.position = 'none'),
             mylegend_stimulus,
             nrow=2,
             ncol=3,
             widths=c(1,1,0.5))


#save to file
tiff(file=paste0(project_path,"/output/figures/group_differences_Bayesian.tiff"), # create a file in tiff format in current working directory
  width=12, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure
      
grid.arrange(g1,
             g2,
             mylegend_group,
             g3+theme(legend.position = 'none'),
             g4+theme(legend.position = 'none'),
             mylegend_stimulus,
             nrow=2,
             ncol=3,
             widths=c(1,1,0.5))

dev.off() #close operation and save file



```

## figure group differences LMM + Bayesian

```{r figure group differences LMM and Bayesian}

#main models - LMM --> calculate
BPS_progression<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
              (1|subjects)+(1|wave),data=df_trial)

BPS_habituation<-lmer(scale(pd_baseline)~t1_diagnosis*sequence_position+
                    (1|subjects)+(1|wave),data=df_trial[df_trial$EventData==201 & df_trial$sequence_position<=10,])

SEPR_progression<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+
              (1|subjects)+(1|wave),data=df_trial)

MMN_progression<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+
              (1|id),data=df_trial)

#main models - Bayesian --> load
load(file=paste0(project_path,'/data/Bayesian_model_BPS_twoway.Rdata')) #draws_BPS_twoway
load(file=paste0(project_path,'/data/Bayesian_model_BPS_supp.Rdata')) #draws_BPS_supp
load(file=paste0(project_path,'/data/Bayesian_model_SEPR_threewayinteraction.Rdata')) #draws_SEPR_threeway
load(file=paste0(project_path,'/data/Bayesian_model_MMN_threewayinteraction.Rdata')) #draws_MMN_threeway

#figure BPS task progression ####

## figure BPS - LMM
model_plot_data_BPS1<-as.data.frame(emmeans(BPS_progression,~t1_diagnosis+EventCounter,
                                       at=list(EventCounter = seq(1,1400,140))))
g1a<-ggplot(model_plot_data_BPS1,aes(x=EventCounter,y=emmean,group=t1_diagnosis,color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  labs(x='task progression (trial)',y='marginalized mean')
  
#figure BPS task progression - draws_BPS_twoway
plot_data<-mcmc_areas_data(draws_BPS_twoway,prob_outer = credible_interval)
g1b1<-ggplot(plot_data[plot_data$parameter=="design_coef[1,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('autistic'))+
  labs(y='posterior estimate of trend',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[1],
                    labels=c('autistic'))

g1b2<-ggplot(plot_data[plot_data$parameter=="design_coef[2,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('non-autistic'))+
  labs(y='',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[2],
                    labels=c('non-autistic'))


#figure BPS habituation ####
 
# BPS habituation --> LMM
model_plot_data_BPS2<-as.data.frame(emmeans(BPS_habituation,~t1_diagnosis+sequence_position,
                                       at=list(sequence_position = seq(1,10))))

g2a<-ggplot(model_plot_data_BPS2,aes(x=sequence_position,y=emmean,group=t1_diagnosis,color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_x_continuous(breaks=seq(0,10,1))+
  labs(x='sequence position',y='marginalized mean')

#draws_BPS_supp
plot_data<-mcmc_areas_data(draws_BPS_supp,prob_outer = credible_interval)
g2b1<-ggplot(plot_data[plot_data$parameter=="design_coef[1,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('autistic'))+
  labs(y='posterior estimate',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[1],
                    labels=c('autistic'))
g2b2<-ggplot(plot_data[plot_data$parameter=="design_coef[2,1]",],aes(x=parameter,y=x,group=parameter,fill=parameter))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75))+
  scale_x_discrete(labels=c('non-autistic'))+
  labs(y='',x='')+
  theme(legend.position = 'none')+
  scale_fill_manual(name = 'parameter: stimulus',
                    values = custom_group_color[2],
                    labels=c('non-autistic'))


#figure SEPR ####

##SEPR LMM
model_plot_data_SEPR<-as.data.frame(emmeans(SEPR_progression,~t1_diagnosis+EventData+EventCounter,
                                            at=list(EventCounter = seq(1,1400,140))))
labels <- c("201" = "standard", 
            "202" = "pitch oddball", 
            "203" = "length oddball", 
            "204" = "pitch & length oddball")
    
g3a<-ggplot(model_plot_data_SEPR,aes(x=EventCounter,y=emmean,
                           group=interaction(EventData,t1_diagnosis),color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  facet_wrap(~EventData,labeller=labeller(EventData = labels),ncol=4)+
  labs(x='task progression (trial)',y='marginalized mean')

#SEPR Bayesian
#draws_SEPR_threeway
plot_data<-mcmc_areas_data(draws_SEPR_threeway,prob_outer = credible_interval)

plot_data$group<-with(plot_data,
                         dplyr::case_when(
                           parameter %in% c("group_event_coefs[1,1]",
                                            "group_event_coefs[3,1]",
                                            "group_event_coefs[5,1]",
                                            "group_event_coefs[7,1]") ~ "autistic",
                           parameter %in% c("group_event_coefs[2,1]",
                                            "group_event_coefs[4,1]",
                                            "group_event_coefs[6,1]",
                                            "group_event_coefs[8,1]") ~ "non-autistic"))
                           
plot_data$stimulus<-with(plot_data,
                         dplyr::case_when(
                           parameter %in% c("group_event_coefs[1,1]",
                                            "group_event_coefs[2,1]") ~ "standard",
                           parameter %in% c("group_event_coefs[3,1]",
                                            "group_event_coefs[4,1]") ~ "pitch oddball",
                           parameter %in% c("group_event_coefs[5,1]",
                                            "group_event_coefs[6,1]") ~ "length oddball",
                           parameter %in% c("group_event_coefs[7,1]",
                                            "group_event_coefs[8,1]") ~ "pitch+length oddball"))

plot_data$stimulus<-factor(plot_data$stimulus,levels=c('standard','pitch oddball','length oddball','pitch+length oddball'))

g3b<-ggplot(plot_data,aes(x=group,y=x,fill=group))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75),scale='width')+
  labs(y='posterior estimate of trend',x='')+
  scale_fill_manual(name = 'group',
                    values = custom_group_color)+
  facet_wrap(~stimulus,ncol=4)
  

#figure MMN ####

#MMN LMM
model_plot_data_MMN<-as.data.frame(emmeans(MMN_progression,~t1_diagnosis+EventData+EventCounter,at=list(EventCounter = seq(1,1400,140))))
    
labels <- c("201" = "standard", 
            "202" = "pitch oddball", 
            "203" = "length oddball", 
            "204" = "pitch & length oddball")
    
g4a<-ggplot(model_plot_data_MMN,aes(x=EventCounter,y=emmean,
                           group=interaction(EventData,t1_diagnosis),color=t1_diagnosis))+
  geom_crossbar(aes(ymin=emmean-1*SE,ymax=emmean+1*SE,fill=t1_diagnosis),alpha=0.8,position='dodge')+
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),position='dodge')+
  scale_color_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  scale_fill_manual(name = "groups", labels = c("autistic", "non-autistic"),values = custom_group_color)+
  facet_wrap(~EventData,labeller=labeller(EventData = labels),ncol=4)+
  labs(x='task progression (trial)',y='marginalized mean')

#MMN Bayesian
#draws_MMN_threeway
plot_data<-mcmc_areas_data(draws_MMN_threeway,prob_outer = credible_interval)

plot_data$group<-with(plot_data,
                         dplyr::case_when(
                           parameter %in% c("group_event_coefs[1,1]",
                                            "group_event_coefs[3,1]",
                                            "group_event_coefs[5,1]",
                                            "group_event_coefs[7,1]") ~ "autistic",
                           parameter %in% c("group_event_coefs[2,1]",
                                            "group_event_coefs[4,1]",
                                            "group_event_coefs[6,1]",
                                            "group_event_coefs[8,1]") ~ "non-autistic"))
                           
plot_data$stimulus<-with(plot_data,
                         dplyr::case_when(
                           parameter %in% c("group_event_coefs[1,1]",
                                            "group_event_coefs[2,1]") ~ "standard",
                           parameter %in% c("group_event_coefs[3,1]",
                                            "group_event_coefs[4,1]") ~ "pitch oddball",
                           parameter %in% c("group_event_coefs[5,1]",
                                            "group_event_coefs[6,1]") ~ "length oddball",
                           parameter %in% c("group_event_coefs[7,1]",
                                            "group_event_coefs[8,1]") ~ "pitch+length oddball"))

plot_data$stimulus<-factor(plot_data$stimulus,levels=c('standard','pitch oddball','length oddball','pitch+length oddball'))

g4b<-ggplot(plot_data,aes(x=group,y=x,fill=group))+
  geom_violin(draw_quantiles = c(0.25,0.5,0.75),scale='width')+
  labs(y='posterior estimate of trend',x='')+
  scale_fill_manual(name = 'group',
                    values = custom_group_color)+
  facet_wrap(~stimulus,ncol=4)


#create composite figure - formatting####
g1b<-grid.arrange(g1b1,g1b2,ncol=2)
g1<-grid.arrange(g1a+theme(legend.position = 'none'),
             g1b,
                 left='baseline pupil size (BPS, z)')

g2b<-grid.arrange(g2b1,g2b2,ncol=2)
g2<-grid.arrange(g2a+theme(legend.position = 'none'),
                 g2b,
                 left='baseline pupil size (BPS, z)')

g3<-grid.arrange(g3a+theme(legend.position = 'none'),
                 g3b+theme(legend.position = 'none'),
                 nrow=2,left='stimulus-evoked pupillary response (SEPR, z)')

g4<-grid.arrange(g4a+theme(legend.position = 'none'),
                 g4b+theme(legend.position = 'none'),
                 nrow=2,left='mismatch-negativity-associated amplitude (MMN-amp, z)')

lay <- rbind(c(1,1,2,2),
             c(3,3,3,3),
             c(4,4,4,4))

grid.arrange(g1,g2,g3,g4,layout_matrix=lay)


#save to file ####
tiff(file=paste0(project_path,"/output/figures/group_differences_LMM_Bayesian_combined.tiff"), # create a file in tiff format in current working directory
  width=8, height=16, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(g1,g2,g3,g4,layout_matrix=lay)

dev.off() #close operation and save file



```

# supplementary analyses
## descriptives per participant 
### descriptive statistics of dependent variables - per participant

```{r descriptive statistics of dependent variables - per-participant, warnings=F}

descriptive_table_data<-
  
  rbind(
    psych::describe(df_timepoint$pd)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_auc.201)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_auc.202)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_auc.203)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$rpd_auc.204)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.201)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.202)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.203)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$mmn.204)[c('n','mean','sd','min','max','skew','kurtosis')],
    psych::describe(df_timepoint$gain_pcdm)[c('n','mean','sd','min','max','skew','kurtosis')]
    )


descriptive_table_data<-round(descriptive_table_data,3)

  variable<-c('BPS',
    'SEPR - standard','SEPR - pitch oddball',
    'SEPR - length oddball','SEPR - pitch+length oddball',
    'MMN - standard','MMN - pitch oddball',
    'MMN - length oddball','MMN - pitch+length oddball',
    'NG')

descriptive_table_data<-cbind(variable,descriptive_table_data)

#produce table
descriptive_table_format<-descriptive_table_data %>%
    kbl(caption = "Per-participant level: Descriptive statistics of baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), mismatch negativity (MMN), and neural gain (NG)",row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

descriptive_table_format

save_kable(descriptive_table_format,file=paste0(project_path,'/output/supplements/descriptivestatistics_DV_perparticipant.html'))


```

### histograms of dependent variables - per participant

```{r historgram of dependent variables - per participant}

grid.arrange(
  ggplot(df_timepoint,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot(df_timepoint,aes(x=gain_pcdm))+geom_histogram(bins=30)+labs(x='',title='neural gain (z)'),
  ggplot(df_timepoint,aes(x=rpd_auc.201))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.202))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.203))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.204))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_timepoint,aes(x=mmn.201))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_timepoint,aes(x=mmn.202))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.203))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.204))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

#save to file
tiff(file=paste0(project_path,"/output/supplements/histograms_dependentvariables_perparticipant.tiff"), # create a file in tiff format in current working directory
  width=8, height=12, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(
  ggplot(df_timepoint,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot(df_timepoint,aes(x=gain_pcdm))+geom_histogram(bins=30)+labs(x='',title='neural gain (z)'),
  ggplot(df_timepoint,aes(x=rpd_auc.201))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.202))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.203))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_timepoint,aes(x=rpd_auc.204))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_timepoint,aes(x=mmn.201))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_timepoint,aes(x=mmn.202))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.203))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_timepoint,aes(x=mmn.204))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

dev.off() #close operation and save file


```






## group comparisons per-participant

```{r group comparisons per participant - table}
### table

BPS_group<-c(summary(lm<-lm(scale(pd)~t1_diagnosis,df_timepoint))[['fstatistic']],
  summary(lm<-lm(scale(pd)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])

NG_group<-c(summary(lm<-lm(scale(gain_pcdm)~t1_diagnosis,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))[['fstatistic']],
             summary(lm<-lm(scale(gain_pcdm)~t1_diagnosis,df_timepoint[df_timepoint$Rsq_pcdm>0.2,]))[['coefficients']][2,c(1,2,4)])

SEPR_201_group<-c(summary(lm<-lm(scale(rpd_auc.201)~t1_diagnosis,df_timepoint))[['fstatistic']],
             summary(lm<-lm(scale(rpd_auc.201)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
SEPR_202_group<-c(summary(lm<-lm(scale(rpd_auc.202)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(rpd_auc.202)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
SEPR_203_group<-c(summary(lm<-lm(scale(rpd_auc.203)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(rpd_auc.203)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
SEPR_204_group<-c(summary(lm<-lm(scale(rpd_auc.204)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(rpd_auc.204)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])

MMN_201_group<-c(summary(lm<-lm(scale(mmn.201)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.201)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
MMN_202_group<-c(summary(lm<-lm(scale(mmn.202)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.202)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
MMN_203_group<-c(summary(lm<-lm(scale(mmn.203)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.203)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])
MMN_204_group<-c(summary(lm<-lm(scale(mmn.204)~t1_diagnosis,df_timepoint))[['fstatistic']],
                  summary(lm<-lm(scale(mmn.204)~t1_diagnosis,df_timepoint))[['coefficients']][2,c(1,2,4)])

group_comparison_table<-round(rbind(BPS_group,NG_group,
      SEPR_201_group,SEPR_202_group,SEPR_203_group,SEPR_204_group,
      MMN_201_group,MMN_202_group,MMN_203_group,MMN_204_group),3)


row_names<-c('BPS - across stimuli','NG - standards',
            'SEPR - standards','SEPR - pitch oddball','SEPR - length oddball','SEPR - pitch+length oddball',
            'MMN - standards','MMN - pitch oddball','MMN - length oddball','MMN - pitch+length oddball')

group_comparison_table<-cbind(row_names,group_comparison_table)

#produce table
group_comparison_table_format<-group_comparison_table %>%
  kbl(caption = "Per-participant level: Group comparisons of baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), mismatch negativity (MMN), and neural gain (NG)",
      row.names = F, col.names = c('','F','df1','df2','b','std. error','p')) %>%
  kable_classic(full_width = F, html_font = "Cambria")

group_comparison_table_format

save_kable(group_comparison_table_format,file=paste0(project_path,'/output/supplements/table_groupcomparison_perparticipant.html'))

```

## covariate effects on group comparisons per-partiicpant

```{r covariate effects on group comparisons per participant}

# significant covariate models ####
#full model - PD
lm<-lm(scale(pd)~t1_diagnosis+
         scale(ageyrs)+scale(t1_piq)+sex+
         as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)
       ,df_timepoint)

  ##supplements - full table
  row_names<-c('intercept','group',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'residuals')


  table_data_model<-round(Anova(lm,type=3),3)
  table_data_model<-cbind(row_names,table_data_model)

  table_data_model

  table_model<-table_data_model %>%
    kbl(caption = "Per participant: Full Linear model of baseline pupil size (BPS) with all potential covariates",
        col.names = c('','Sum Sq','df','F','p'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

  table_model

  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_linearmodel_BPS_allcovariates.html'))

#significant covariate model - PD
lm<-lm(scale(pd)~t1_diagnosis+
         scale(ageyrs)+scale(t1_piq)
       ,df_timepoint)

  ##supplements - full table
  row_names<-c('intercept','group',
               'age','perceptual IQ','residuals')


  table_data_model<-round(Anova(lm,type=3),3)
  table_data_model<-cbind(row_names,table_data_model)

  table_data_model

  table_model<-table_data_model %>%
    kbl(caption = "Per participant: Reduced Linear model of baseline pupil size (BPS) with significant covariates",
        col.names = c('','Sum Sq','df','F','p'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

  table_model

  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_linearmodel_BPS_signcovariates.html'))



# full modelNG
lm<-lm(scale(gain_pcdm)~t1_diagnosis+
         scale(ageyrs)+scale(t1_piq)+sex+
         as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)
       ,df_timepoint[df_timepoint$Rsq_pcdm>0.2,])

  ##supplements - full table
  row_names<-c('intercept','group',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'residuals')


  table_data_model<-round(Anova(lm,type=3),3)
  table_data_model<-cbind(row_names,table_data_model)

  table_data_model

  table_model<-table_data_model %>%
    kbl(caption = "Per participant: Full Linear model of neural gain (NG) with all potential covariates",
        col.names = c('','Sum Sq','df','F','p'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

  table_model

  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_linearmodel_NG_allcovariates.html'))



#significant covariate model - NG
lm<-lm(scale(gain_pcdm)~t1_diagnosis
       ,df_timepoint[df_timepoint$Rsq_pcdm>0.2,])

    ##supplements - full table
    row_names<-c('intercept','group','residuals')


    table_data_model<-round(Anova(lm,type=3),3)
    table_data_model<-cbind(row_names,table_data_model)

    table_data_model

    table_model<-table_data_model %>%
      kbl(caption = "Per participant: Reduced Linear model of neural gain (NG) with significant covariates",
          col.names = c('','Sum Sq','df','F','p'),
          row.names = F) %>%
      kable_classic(full_width = F, html_font = "Cambria")

    table_model

    save_kable(table_model, file= paste0(project_path,'/output/supplements/table_linearmodel_NG_signcovariates.html'))


```


## descriptives per trial
### descriptive statistics of dependent variables - per trial

```{r descriptive statistics of dependent variables - per-trial, warnings=F}

descriptive_table_data<-
  
rbind(
psych::describe(df_trial$pd_baseline)[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$rpd_auc[df_trial$EventData=='201'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$rpd_auc[df_trial$EventData=='202'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$rpd_auc[df_trial$EventData=='203'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$rpd_auc[df_trial$EventData=='204'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$mmn[df_trial$EventData=='201'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$mmn[df_trial$EventData=='202'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$mmn[df_trial$EventData=='203'])[c('n','mean','sd','min','max','skew','kurtosis')],
psych::describe(df_trial$mmn[df_trial$EventData=='204'])[c('n','mean','sd','min','max','skew','kurtosis')]

)


descriptive_table_data<-round(descriptive_table_data,3)

  variable<-c('BPS',
    'SEPR - standard','SEPR - pitch oddball',
    'SEPR - length oddball','SEPR - pitch+length oddball',
    'MMN - standard','MMN - pitch oddball',
    'MMN - length oddball','MMN - pitch+length oddball')

descriptive_table_data<-cbind(variable,descriptive_table_data)

#produce table
descriptive_table_format<-descriptive_table_data %>%
    kbl(caption = "Per-trial level: Descriptive statistics of baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), and mismatch negativity (MMN)",row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")

descriptive_table_format

save_kable(descriptive_table_format,file=paste0(project_path,'/output/supplements/descriptivestatistics_DV_pertrial.html'))

```



### histograms of dependent variables - per trial

```{r historgram of dependent variables - per trial}

grid.arrange(
  ggplot(df_trial,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot()+theme_void(),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

#save to file
tiff(file=paste0(project_path,"/output/supplements/histograms_dependentvariables_pertrial.tiff"), # create a file in tiff format in current working directory
  width=8, height=12, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(
  ggplot(df_trial,aes(x=pd))+geom_histogram(bins=30)+labs(x='',title='BPS (mm)'),
  ggplot()+theme_void(),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - standard (mm)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=rpd_auc))+geom_histogram(bins=30)+labs(x='',title='SEPR - pitch+length oddball (mm)'),
  ggplot(df_trial[df_trial$EventData=='201',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - standard (uV)'),
  ggplot(df_trial[df_trial$EventData=='202',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='203',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - length oddball (uV)'),
  ggplot(df_trial[df_trial$EventData=='204',],
         aes(x=mmn))+geom_histogram(bins=30)+labs(x='',title='MMN - pitch+length oddball (uV)'),
  ncol=2)

dev.off() #close operation and save file


```








## linearity of effect of trial number

```{r linearity SEPR}

###polynomial fit of trial counter
  linear_fit<-lmer(scale(rpd_auc)~scale(EventCounter)+
          (1|subjects)+(1|wave),data=df_trial,REML=F)

  quadratic_fit<-lmer(scale(rpd_auc)~stats::poly(scale(EventCounter),2)+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

  cubic_fit<-lmer(scale(rpd_auc)~stats::poly(scale(EventCounter),3)+
               (1|subjects)+(1|wave),data=df_trial,REML=F)
  
  anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-cbind(c('linear fit','quadratic fit','cubic fit'),table_model_compare)
  
  table_formatted<-table_model_compare %>%
  kbl(caption = "Model comparison on the linearity of trial number on stimulus-evoked pupillary response (SEPR)",
      col.names = c('','number of parameters','AIC','BIC','log likelihood','deviance','Chi-squared','df','p-value'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_formatted

save_kable(table_formatted, file=paste0(project_path,'/output/supplements/table_modelcomparison_linearity_SEPR.html'))
  
  
```

```{r linearity BPS}

###polynomial fit of trial counter
  linear_fit<-lmer(scale(pd_baseline)~scale(EventCounter)+
          (1|subjects)+(1|wave),data=df_trial,REML=F)

  quadratic_fit<-lmer(scale(pd_baseline)~stats::poly(scale(EventCounter),2)+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

  cubic_fit<-lmer(scale(pd_baseline)~stats::poly(scale(EventCounter),3)+
               (1|subjects)+(1|wave),data=df_trial,REML=F)
  
  anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-cbind(c('linear fit','quadratic fit','cubic fit'),table_model_compare)
  
  table_formatted<-table_model_compare %>%
  kbl(caption = "Model comparison on the linearity of trial number on baseline pupil size (BPS)",
      col.names = c('','number of parameters','AIC','BIC','log likelihood','deviance','Chi-squared','df','p-value'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_formatted

save_kable(table_formatted, file=paste0(project_path,'/output/supplements/table_modelcomparison_linearity_BPS.html'))

```

```{r linearity MMN}

###polynomial fit of trial counter
  linear_fit<-lmer(scale(mmn)~scale(EventCounter)+
          (1|subjects)+(1|wave),data=df_trial,REML=F)

  quadratic_fit<-lmer(scale(mmn)~stats::poly(scale(EventCounter),2)+
              (1|subjects)+(1|wave),data=df_trial,REML=F)

  cubic_fit<-lmer(scale(mmn)~stats::poly(scale(EventCounter),3)+
               (1|subjects)+(1|wave),data=df_trial,REML=F)
  
  anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-cbind(c('linear fit','quadratic fit','cubic fit'),table_model_compare)
  
  table_formatted<-table_model_compare %>%
  kbl(caption = "Model comparison on the linearity of trial number on mismatch negativity (MMN)",
      col.names = c('','number of parameters','AIC','BIC','log likelihood','deviance','Chi-squared','df','p-value'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_formatted

save_kable(table_formatted, file=paste0(project_path,'/output/supplements/table_modelcomparison_linearity_MMN.html'))

```

```{r linearity of sequence position on BPS}

###polynomial fit of trial counter
  linear_fit<-lmer(scale(pd_baseline)~scale(sequence_position)+
          (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<10,],REML=F)

  quadratic_fit<-lmer(scale(pd_baseline)~stats::poly(scale(sequence_position),2)+
              (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<10,],REML=F)

  cubic_fit<-lmer(scale(pd_baseline)~stats::poly(scale(sequence_position),3)+
               (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<10,],REML=F)
  
  anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-anova(linear_fit,quadratic_fit,cubic_fit)
  
  table_model_compare<-cbind(c('linear fit','quadratic fit','cubic fit'),table_model_compare)
  
  table_formatted<-table_model_compare %>%
  kbl(caption = "Model comparison on the linearity of trial number on mismatch negativity (MMN)",
      col.names = c('','number of parameters','AIC','BIC','log likelihood','deviance','Chi-squared','df','p-value'),
      row.names = F) %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_formatted

save_kable(table_formatted, file=paste0(project_path,'/output/supplements/table_modelcomparison_linearity_sequenceposition.html'))

```

## intraclass correaltion of the dependent variables

```{r ICC_of_dependent_variables}

#select only individuals that have both waves
subjects_with_both_waves<-df_dem$subjects[df_dem$timepoints=='wave1+2']
df_trial_wave12<-df_trial[df_trial$subjects %in% subjects_with_both_waves,]

#BPS
baseline_pd_wave1<-with(df_trial_wave12[df_trial_wave12$wave=='wave1',],aggregate(pd_baseline,by=list(subjects),FUN=mean,na.rm=T))
baseline_pd_wave2<-with(df_trial_wave12[df_trial_wave12$wave=='wave2',],aggregate(pd_baseline,by=list(subjects),FUN=mean,na.rm=T))
names(baseline_pd_wave1)<-c('subjects','wave1')
names(baseline_pd_wave2)<-c('subjects','wave2')

baseline_pd_waves<-merge(baseline_pd_wave1,baseline_pd_wave2,by='subjects')

cor.test(baseline_pd_waves$wave1,baseline_pd_waves$wave2)

psych::ICC(baseline_pd_waves[,-1])
###--> ICC: 0.88 [0.81, 0.93]


#SEPR
rpd204_wave1<-with(df_trial_wave12[df_trial_wave12$EventData=='204' & df_trial_wave12$wave=='wave1',],aggregate(rpd_auc.204,by=list(subjects),FUN=mean,na.rm=T))
rpd204_wave2<-with(df_trial_wave12[df_trial_wave12$EventData=='204' & df_trial_wave12$wave=='wave2',],aggregate(rpd_auc.204,by=list(subjects),FUN=mean,na.rm=T))

names(rpd204_wave1)<-c('subjects','wave1')
names(rpd204_wave2)<-c('subjects','wave2')

rpd_waves<-merge(rpd204_wave1,rpd204_wave2,by='subjects')

cor.test(rpd_waves$wave1,rpd_waves$wave2)
##--> r =.23

psych::ICC(rpd_waves[,-1])
###--> ICC: 0.37 [-0.01, 0.60]


names(df_timepoint)

#NG
gain_wave1<-with(df_timepoint[df_timepoint$wave=='wave1',],aggregate(gain_pcdm,by=list(subjects),FUN=mean,na.rm=T))
gain_wave2<-with(df_timepoint[df_timepoint$wave=='wave2',],aggregate(gain_pcdm,by=list(subjects),FUN=mean,na.rm=T))

names(gain_wave1)<-c('subjects','wave1')
names(gain_wave2)<-c('subjects','wave2')

gain_waves<-merge(gain_wave1,gain_wave2,by='subjects')

cor.test(gain_waves$wave1,gain_waves$wave2)
##--> r =.23

psych::ICC(gain_waves[,-1])
###--> ICC: 0.37 [0.02, 0.59]

```

For participants with both measurement timepoints, we estimated a relative stability of BPS (intraclass correlation, ICC = 0.88 [0.81, 0.93]), while SEPR (ICC = 0.37 [-0.01, 0.60]) and NG (ICC = 0.37 [0.02, 0.59]) were variable across timepoints. EEG data was only available at one timepoint. 



## full model tables

```{r BPS full model LMM}

#full model  - save to supplements      
lmm_full<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|subjects)+(1|wave),data=df_trial)


##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'S x group','S x TP','group x TP','S x group x TP')
  
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: baseline pupil size (BPS) - task progression",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_BPS.html'))

```

```{r BPS full model LMM - habituation}

#recoding
df_trial$sequence_position_z<-scale(df_trial$sequence_position)


  ### supplementary analysis
lmm_full<-lmer(scale(pd_baseline)~t1_diagnosis*sequence_position_z+
            scale(ageyrs)+scale(t1_piq)+sex+
            as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
            (1|subjects)+(1|wave),data=df_trial[df_trial$EventData==201 & df_trial$sequence_position<10,])

anova(lmm_full)

##supplements - full table
  row_names<-c('group','sequence position - continuous (SP)',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'group x SP')
  
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: baseline pupil size (BPS) - habituation",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_BPS_habituation.html'))


```

```{r SEPR full model LMM}

#full model
lmm_full<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|subjects)+(1|wave),data=df_trial)
              #--> group effect (interaction) is robust in reduced and full model


##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'S x group','S x TP','group x TP','S x group x TP')
   
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: stimulus-evoked pupillary response (SEPR) - task progression",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_SEPR.html'))


```

```{r SEPR full model LMM - habituation}

#SEPR - categorical habituation
  lmm_full<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*as.factor(sequence_position)+
            scale(ageyrs)+scale(t1_piq)+sex+
             scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|subjects)+(1|wave),data=df_trial[df_trial$sequence_position<4,])

   anova(lmm_full)
  
##supplements - full table
  row_names<-c('stimulus (S)','group','sequence position - categorical (SP)',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'S x group','S x SP','group x SP','S x group x SP')
   
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: stimulus-evoked pupillary response (SEPR) - habituation",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_SEPR_habituation.html'))



```

```{r MMN full model LMM}

#full model
lmm_full<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|id),data=df_trial)
              #--> group effect (interaction) is robust in reduced and full model


##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'S x group','S x TP','group x TP','S x group x TP')
   
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: mismatch negativity - task progression",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_MMN.html'))


```

```{r MMN full model LMM - habituation}

#full model
lmm_full<-lmer(scale(mmn)~EventData*t1_diagnosis*sequence_position_z+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)+
              (1|id),data=df_trial)
              #--> group effect (interaction) is robust in reduced and full model


##supplements - full table
  row_names<-c('stimulus (S)','group','sequence position - continuous (SP)',
               'age','perceptual IQ','sex',
               'sampling rate','gaze center deviation','data quality',
               'S x group','S x SP','group x SP','S x group x SP')
   
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: mismatch negativity - habituation",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_MMN_habituation.html'))


```

```{r NG full model LM}

lmm_full<-lm(scale(gain_pcdm)~t1_diagnosis+
               scale(ageyrs)+scale(t1_piq)+sex+
               scale(center_dev)+as.factor(sampling_rate)+scale(missing_data_trial)
               ,df_timepoint[df_timepoint$Rsq_pcdm>0.2,])

anova(lmm_full) #higher gain in ASD for standards

##supplements - full table
  row_names<-c('group',
               'age','perceptual IQ','sex',
               'gaze center deviation','sampling rate','data quality',
               'residuals')
   
  
  table_data_model<-round(anova(lmm_full),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear model: neural gain",
        col.names = c('','Df','Sum Sq','Mean Sq','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_model_NG.html'))




```

## covariate associations with BPS, SEPR, MMN

###BPS

```{r covariate associations BPS}

#full model  - save to supplements      
lmm_full<-lmer(scale(pd_baseline)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|subjects)+(1|wave),data=df_trial)


age_eff<-cbind('age (yrs)',
               round(fixef(lmm_full)['scale(ageyrs)'],2),
               round(confint(lmm_full,parm = 'scale(ageyrs)'),2))

samplingrate_eff<-confint(contrast(emmeans(lmm_full,~sampling_rate),
                                   'revpairwise'))[c('contrast','estimate','asymp.LCL','asymp.UCL')]

centerdev_eff<-cbind('gaze center deviation (%)',
               round(fixef(lmm_full)['scale(center_dev)'],2),
               round(confint(lmm_full,parm = 'scale(center_dev)'),2))


dquality_eff<-cbind('missing data per trial (%)',
               round(fixef(lmm_full)['scale(missing_data_trial)'],2),
               round(confint(lmm_full,parm = 'scale(missing_data_trial)'),2))


cov_eff_data<-rbind(age_eff,
      c('sampling rate (300Hz vs. 120Hz)',round(samplingrate_eff[1,2:4],2)),
      centerdev_eff,
      dquality_eff)


cov_eff_table<-cov_eff_data %>%
    kbl(caption = "Covariate effects on baseline pupil size (BPS)",
        col.names = c('covariate','estimate','lower bound (2.5%)','upper bound (97.5%)'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  cov_eff_table
  
  save_kable(cov_eff_table, file= paste0(project_path,'/output/supplements/table_BPS_covariateeffects.html'))

```

###SEPR

```{r covariate associations SEPR}

#full model  - save to supplements      
lmm_full<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|subjects)+(1|wave),data=df_trial)


dquality_eff<-cbind('missing data per trial (%)',
               round(fixef(lmm_full)['scale(missing_data_trial)'],2),
               round(confint(lmm_full,parm = 'scale(missing_data_trial)'),2))


cov_eff_data<-rbind(dquality_eff)

cov_eff_table<-cov_eff_data %>%
    kbl(caption = "Covariate effects on stimulus-evoked pupillary response (SEPR)",
        col.names = c('covariate','estimate','lower bound (2.5%)','upper bound (97.5%)'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  cov_eff_table
  
  save_kable(cov_eff_table, file= paste0(project_path,'/output/supplements/table_SEPR_covariateeffects.html'))

```

###MMN

```{r covariate associations MMN}

#full model  - save to supplements      
lmm_full<-lmer(scale(mmn)~EventData*t1_diagnosis*stats::poly(EventCounter,3)+
      scale(ageyrs)+scale(t1_piq)+sex+
      as.factor(sampling_rate)+scale(center_dev)+scale(missing_data_trial)+
      (1|id),data=df_trial)


age_eff<-cbind('age (yrs)',
               round(fixef(lmm_full)['scale(ageyrs)'],2),
               round(confint(lmm_full,parm = 'scale(ageyrs)'),2))


dquality_eff<-cbind('missing data per trial (%)',
               round(fixef(lmm_full)['scale(missing_data_trial)'],2),
               round(confint(lmm_full,parm = 'scale(missing_data_trial)'),2))


cov_eff_data<-rbind(age_eff,
                    dquality_eff)

cov_eff_table<-cov_eff_data %>%
    kbl(caption = "Covariate effects on mismatch negativity (MMN)",
        col.names = c('covariate','estimate','lower bound (2.5%)','upper bound (97.5%)'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  cov_eff_table
  
  save_kable(cov_eff_table, file= paste0(project_path,'/output/supplements/table_MMN_covariateeffects.html'))

```



## scaling of BPS SEPR and MMN for demographic variables between groups

### SEPR scaling with demographic variables

```{r SEPR scaling with demographics between groups}

  ###SCALING of pupillary response AUC WITH MAIN DEMOGRAPHIC VARIABLES ####

  hist(df_trial$ageyrs) ## pick 10 - 25 years
  hist(df_trial$t1_piq) ### IQ 80-135

    labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch+length oddball")

  #age
  g1<-ggplot(df_trial[df_trial$ageyrs>=10 & df_trial$ageyrs<=25,],aes(x=ageyrs,y=scale(rpd_auc),group=interaction(EventData,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='age (years)',y='')+
    coord_cartesian(ylim = c(-0.05, 0.15))+
    theme_bw()

  #IQ
  g2<-ggplot(df_trial[df_trial$t1_piq > 80 & df_trial$t1_piq < 135,],aes(x=t1_piq,y=scale(rpd_auc),group=interaction(EventData,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='perceptual IQ',y='')+
    coord_cartesian(ylim = c(-0.05, 0.15))+
    theme_bw()


    #summary functions
    stat_sum_df <- function(fun, geom="crossbar", ...) {
      stat_summary(fun.data=fun, geom=geom, width=0.2, ...)
    }

    stat_sum_single <- function(fun, geom="point", ...) {
      stat_summary(fun=fun, geom=geom, size = 3, ...)
    }

  group_labels<-c("female","female","male","male")
  g3<-ggplot(df_trial,aes(x=interaction(t1_diagnosis,sex),y=scale(rpd_auc),group=interaction(EventData,t1_diagnosis,sex),color=t1_diagnosis))+
    stat_sum_df("mean_cl_normal", geom = "errorbar")+
    stat_sum_single("mean")+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_x_discrete(labels=group_labels)+
    labs(x='biological sex',y='')+
    coord_cartesian(ylim = c(-0.05, 0.15))+
    theme_bw()

  
  grid.arrange(g1,g2,g3,nrow=3)
  
  
  #save to file
tiff(file=paste0(project_path,"/output/supplements/SEPR_scaling_withdemographics_between_groups.tiff"), # create a file in tiff format in current working directory
  width=10, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

  grid.arrange(g1,g2,g3,nrow=3,left='stimulus-evoked pupillary response (SEPR, z)')

        
dev.off() #close operation and save file



```

### BPS scaling with demographic variables

```{r BPS scaling with demographics between groups}
  ###SCALING of baseline pupil size WITH MAIN DEMOGRAPHIC VARIABLES ####

  labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch+length oddball")
  #age
  g1<-ggplot(df_trial[df_trial$ageyrs>=10 & df_trial$ageyrs<=25,],aes(x=ageyrs,y=pd_baseline,group=interaction(EventData,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='age (years)',y='')+
    theme_bw()

  #IQ
  g2<-ggplot(df_trial[df_trial$t1_piq > 80 & df_trial$t1_piq < 135,],aes(x=t1_piq,y=pd_baseline,group=interaction(EventData,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='perceptual IQ',y='')+
    theme_bw()


  #summary functions
  stat_sum_df <- function(fun, geom="crossbar", ...) {
    stat_summary(fun.data=fun, geom=geom, width=0.2, ...)
  }

  stat_sum_single <- function(fun, geom="point", ...) {
    stat_summary(fun=fun, geom=geom, size = 3, ...)
  }

  group_labels<-c("female","female","male","male")
  g3<-ggplot(df_trial,aes(x=interaction(t1_diagnosis,sex),y=pd_baseline,group=interaction(EventData,t1_diagnosis,sex),color=t1_diagnosis))+
    stat_sum_df("mean_cl_normal", geom = "errorbar")+
    stat_sum_single("mean")+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_x_discrete(labels=group_labels)+
    labs(x='biological sex',y='')+
    theme_bw()

  grid.arrange(g1,g2,g3,nrow=3)
  ####--> higher IQ and lower age in autistic females?

      with(df_dem,by(t1_piq,interaction(t1_sex,t1_diagnosis),summary))
      with(df_dem,by(t1_ageyrs,interaction(t1_sex,t1_diagnosis),summary))

  
    #save to file
tiff(file=paste0(project_path,"/output/supplements/BPS_scaling_withdemographics_between_groups.tiff"), # create a file in tiff format in current working directory
  width=10, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

  grid.arrange(g1,g2,g3,nrow=3,left='baseline pupil size (mm)')

        
dev.off() #close operation and save file
    
      

```

MMN scaling with demographic variables

```{r MMN scaling with demographics between groups}
  ###SCALING of MMN WITH MAIN DEMOGRAPHIC VARIABLES ####

    labels <- c("201" = "standard", "202" = "pitch oddball", "203" = "length oddball", "204" = "pitch+length oddball")

  #age
  g1<-ggplot(df_trial[df_trial$ageyrs>=10 & df_trial$ageyrs<=25,],aes(x=ageyrs,y=mmn,group=interaction(EventData,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='age (years)',y='')+
    theme_bw()

  #IQ
  g2<-ggplot(df_trial[df_trial$t1_piq > 80 & df_trial$t1_piq < 135,],aes(x=t1_piq,y=mmn,group=interaction(EventData,t1_diagnosis),color=t1_diagnosis,fill=t1_diagnosis))+
    geom_smooth(method='lm')+facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    scale_fill_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    labs(x='perceptual IQ',y='')+
    theme_bw()


  #summary functions
  stat_sum_df <- function(fun, geom="crossbar", ...) {
    stat_summary(fun.data=fun, geom=geom, width=0.2, ...)
  }

  stat_sum_single <- function(fun, geom="point", ...) {
    stat_summary(fun=fun, geom=geom, size = 3, ...)
  }

  group_labels<-c("female","female","male","male")
  g3<-ggplot(df_trial,aes(x=interaction(t1_diagnosis,sex),y=mmn,group=interaction(EventData,t1_diagnosis,sex),color=t1_diagnosis))+
    stat_sum_df("mean_cl_normal", geom = "errorbar")+
    stat_sum_single("mean")+
    scale_color_manual(name = "group", labels = c("autistic", "non-autistic"),values = custom_group_color)+
    facet_grid(~EventData,labeller=labeller(EventData = labels))+
    scale_x_discrete(labels=group_labels)+
    labs(x='biological sex',y='')+
    theme_bw()

  grid.arrange(g1,g2,g3,nrow=3)
  ####--> higher IQ and lower age in autistic females?

  with(df_dem,by(t1_piq,interaction(t1_sex,t1_diagnosis),summary))
  with(df_dem,by(t1_ageyrs,interaction(t1_sex,t1_diagnosis),summary))


  #save to file
tiff(file=paste0(project_path,"/output/supplements/MMN_scaling_withdemographics_between_groups.tiff"), # create a file in tiff format in current working directory
  width=10, height=8, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

  grid.arrange(g1,g2,g3,nrow=3,left='mismatch negativity (uV)')

        
dev.off() #close operation and save file





```



## Bayesian chains - convergence and stationarity of estimates

```{r figure Bayesian models chain convergence stationarity}

###load saved Bayeisan posterior estimates
load(file=paste0(project_path,'/data/Bayesian_model_BPS_twoway.Rdata')) #draws_BPS_twoway
load(file=paste0(project_path,'/data/Bayesian_model_BPS_supp.Rdata')) #draws_BPS_supp
load(file=paste0(project_path,'/data/Bayesian_model_SEPR_threewayinteraction.Rdata')) #draws_SEPR_threeway
load(file=paste0(project_path,'/data/Bayesian_model_MMN_threewayinteraction.Rdata')) #draws_MMN_threeway

### for supplements

#chain converge and stationary

# #draws_BPS_twoway - chain converge and stationary
var_names <- c("design_coef[1,1]" = "BPS_task_progression: autistic",
               "design_coef[2,1]" = "BPS_task_progression: non-autistic") #define parameter names

g1a<-mcmc_trace(draws_BPS_twoway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed))))
g1b<-mcmc_dens(draws_BPS_twoway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)))) 

# #draws_BPS_supp - chain converge and stationary
var_names <- c("design_coef[1,1]" = "BPS_sequence_position: autistic",
               "design_coef[2,1]" = "BPS_sequence_position: non-autistic") #define parameter names

g2a<-mcmc_trace(draws_BPS_supp, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed))))
g2b<-mcmc_dens(draws_BPS_supp, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)))) 

# #draws_SEPR_threeway - chain converge and stationary
var_names <- c("group_event_coefs[1,1]" = "SEPR_trend: autistic_standard",
               "group_event_coefs[2,1]" = "SEPR_trend: non-autistic_standard",
               "group_event_coefs[3,1]" = "SEPR_trend: autistic_pitch",
               "group_event_coefs[4,1]" = "SEPR_trend: non-autistic_pitch",
               "group_event_coefs[5,1]" = "SEPR_trend: autistic_length",
               "group_event_coefs[6,1]" = "SEPR_trend: non-autistic_length",
               "group_event_coefs[7,1]" = "SEPR_trend: autistic_pitch+length",
               "group_event_coefs[8,1]" = "SEPR_trend: non-autistic_pitch+length") #define parameter names

g3a<-mcmc_trace(draws_SEPR_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1))
g3b<-mcmc_dens(draws_SEPR_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1)) 

# #draws_MMN_threeway - chain converge and stationary
var_names <- c("group_event_coefs[1,1]" = "MMN_trend: autistic_standard",
               "group_event_coefs[2,1]" = "MMN_trend: non-autistic_standard",
               "group_event_coefs[3,1]" = "MMN_trend: autistic_pitch",
               "group_event_coefs[4,1]" = "MMN_trend: non-autistic_pitch",
               "group_event_coefs[5,1]" = "MMN_trend: autistic_length",
               "group_event_coefs[6,1]" = "MMN_trend: non-autistic_length",
               "group_event_coefs[7,1]" = "MMN_trend: autistic_pitch+length",
               "group_event_coefs[8,1]" = "MMN_trend: non-autistic_pitch+length") #define parameter names

g4a<-mcmc_trace(draws_MMN_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1))
g4b<-mcmc_dens(draws_MMN_threeway, 
           facet_args = list(labeller = labeller(parameter = as_labeller(var_names, label_parsed)),nrow=1)) 


#extract legend
g_legend<-function(x){
  tmp <- ggplot_gtable(ggplot_build(x))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend<-g_legend(g1a)

#arrange figures
g1<-grid.arrange(g1a+theme(legend.position = 'none'),g1b,top="BPS - task progression")
g2<-grid.arrange(g2a+theme(legend.position = 'none'),g2b,top="BPS - sequence position")
g3<-grid.arrange(g3a+theme(legend.position = 'none'),g3b,top="SEPR trend")
g4<-grid.arrange(g4a+theme(legend.position = 'none'),g4b,top="MMN trend")


g_figures<-grid.arrange(
  grid.arrange(g1,g2,ncol=2),
  g3,g4,
  nrow=3)


grid.arrange(g_figures,mylegend,ncol=2,widths=c(20,1))

#save to file
tiff(file=paste0(project_path,"/output/supplements/Bayesian_chains_convergence.tiff"), # create a file in tiff format in current working directory
  width=25, height=15, units="in", res=300, compression='lzw') #define size and resolution of the resulting figure

grid.arrange(g_figures,mylegend,ncol=2,widths=c(20,1))
      
dev.off() #close operation and save file



```


## moderation of BPS on SEPR and MMN interaction by group

```{r moderation by BPS}

#moderation of BPS on SEPR interaction with group
  lmm_REML_BPS<-lmer(scale(rpd_auc)~EventData*t1_diagnosis*EventCounter+scale(pd_baseline)+
                   (1|subjects)+(1|wave),data=df_trial)

  anova(lmm_REML_BPS)
  
  
  ##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'BPS',
               'S x group','S x TP','group x TP','S x group x TP')
  
  
  table_data_model<-round(anova(lmm_REML_BPS),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: Moderation of group differences in dynamic stimulus-evoked pupillary response (SEPR) group differences by baseline pupil size (BPS)",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_moderation_SEPR_by_BPS.html'))


  #moderation of BPS on MMN interaction with group
  lmm_REML_BPS<-lmer(scale(mmn)~EventData*t1_diagnosis*EventCounter+scale(pd_baseline)+
                       (1|subjects)+(1|wave),data=df_trial)

  anova(lmm_REML_BPS)

  ##supplements - full table
  row_names<-c('stimulus (S)','group','task progression (TP)',
               'BPS',
               'S x group','S x TP','group x TP','S x group x TP')
  
  
  table_data_model<-round(anova(lmm_REML_BPS),3)
  table_data_model<-cbind(row_names,table_data_model)
  
  table_data_model
  
  table_model<-table_data_model %>%
    kbl(caption = "Linear mixed model: Moderation of group differences in dynamic mismatch negativity (MMN) by baseline pupil size (BPS)",
        col.names = c('','Sum Sq','Mean Sq','df1','df2','F','p adjusted'),
        row.names = F) %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  
  table_model
  
  save_kable(table_model, file= paste0(project_path,'/output/supplements/table_moderation_MMN_by_BPS.html'))


  
  
```

