---
title: "supplementary information"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(htmltools)
require(knitr)
require(tiff)

project_path<-getwd()

```

# table of content

abbreviations: baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), mismatch-negativity-associated amplitude (MMN-amp), neural gain (NG)

-   [supplement 1 - descriptive statistics per participant][supplement 1]
-   [supplement 2 - descriptive statistics per trial][supplement 2]
-   [supplement 3 - group comparisons on per-participant estimates][supplement 3]
-   [supplement 4 - covariate effects: BPS group difference][supplement 4]
-   [supplement 5 - covariate effects: NG group difference][supplement 5]
-   [supplement 6 - BPS association with age, IQ, biological sex][supplement 6]
-   [supplement 7 - SEPR association with age, IQ, biological sex][supplement 7]
-   [supplement 8 - MMN-amp association with age, IQ, biological sex][supplement 8]
-   [supplement 9 - correlation table of BPS, SEPR, MMN-amp, NG][supplement 9]
-   [supplement 10 - polynomial fit comparison for BPS, SEPR; MMN-amp][supplement 10]
-   [supplement 11 - BPS linear mixed model - task progression][supplement 11]
-   [supplement 12 - stimulus habituation of BPS, SEPR, and MMN-amp][supplement 12]
-   [supplement 13 - BPS linear mixed model - habituation][supplement 13]
-   [supplement 14 - SEPR linear mixed model - task progression][supplement 14]
-   [supplement 15 - SEPR linear mixed model - habituation][supplement 15]
-   [supplement 16 - MMN-amp linear mixed model - task progression][supplement 16]
-   [supplement 17 - MMN-amp linear mixed model - task progression][supplement 17]
-   [supplement 18 - Bayesian posterior estimate MCMC chain convergence][supplement 18]
-   [supplement 19 - histograms of per-participant estimates][supplement 19]
-   [supplement 20 - histograms of per-trial estimates][supplement 20]


# supplement 1

```{r descriptive statistics per participant}

htmltools::includeHTML("./output/supplements/descriptivestatistics_DV_perparticipant.html")

```

# supplement 2

```{r descriptive statistics per trial}

htmltools::includeHTML("./output/supplements/descriptivestatistics_DV_pertrial.html")

```

# supplement 3 

```{r group comparisons per participant}

htmltools::includeHTML("./output/supplements/table_groupcomparison_perparticipant.html")

```

# supplement 4 

```{r covariate effects BPS}

htmltools::includeHTML("./output/supplements/table_linearmodel_BPS_allcovariates.html")

htmltools::includeHTML("./output/supplements/table_linearmodel_BPS_signcovariates.html")


```

# supplement 5 

```{r covariate effects NG}

htmltools::includeHTML("./output/supplements/table_linearmodel_NG_allcovariates.html")

htmltools::includeHTML("./output/supplements/table_linearmodel_NG_signcovariates.html")


```

# supplement 6

```{r figure BPS scaling}

grid::grid.raster(tiff::readTIFF("./output/supplements/BPS_scaling_withdemographics_between_groups.tiff") )

```

# supplement 7

```{r figure SEPR scaling}

grid::grid.raster(tiff::readTIFF("./output/supplements/SEPR_scaling_withdemographics_between_groups.tiff") )

```

# supplement 8

```{r figure MMN scaling}

grid::grid.raster(tiff::readTIFF("./output/supplements/MMN_scaling_withdemographics_between_groups.tiff") )

```

# supplement 9 

```{r correlation table}

htmltools::includeHTML("./output/supplements/associations_DV_perparticipant.html")


```

# supplement 10 

```{r polynomial fit comparisons}

htmltools::includeHTML("./output/supplements/table_modelcomparison_linearity_BPS.html")

htmltools::includeHTML("./output/supplements/table_modelcomparison_linearity_SEPR.html")

htmltools::includeHTML("./output/supplements/table_modelcomparison_linearity_MMN.html")

```

# supplement 11 

```{r BPS linear mixed models - task progression}

htmltools::includeHTML("./output/supplements/table_model_BPS.html")

htmltools::includeHTML("./output/supplements/table_BPS_covariateeffects.html")

```

# supplement 12

In this supplementary analyses, we investigated stimulus habituation of baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), and mismatch-negativity-associated amplitude. For this analyses, we exchanged the fixed effect of task progression with sequence position, i.e., trials since last oddball. This was modeled for sequence position as continuous variable (sequence position: 1-10) or as categorical variable (sequence position: 1 vs. 2 vs 3).

BPS did not differ between groups for sequence position as continuous (F(1,246853) = 1.45, p = .229) or categorical variable (F(1,133874) < 1). When limiting the analysis to the frequent standard stimuli (mR² = 0.01, cR² = 0.76, figure 4 – top right), BPS was characterized by an interaction of group x sequence position (F(1,204703) = 3.87, p = .049, BF < 1), which was not moderated by including covariates (F(1,204685) = 3.99, p < .045, supplement 13). Post-hoc analysis showed that the autistic group was associated with a BPS increase for standard stimuli in a sequence (Δβ = 0.01 [0.00, 0.02]) compared to the non-autistic group (Δβ = -0.01 [-0.02, 0.00]). This is supported by Bayesian posterior estimates that showed a BPS increase with standard stimuli sequences (b = 0.20 [0.07, 0.30]) compared to non-autistic individuals (b = 0.00 [0.00, 0.00]).

SEPR did not differ between groups for sequence position as continuous (F(1,246857) = 2.11, p = .146). For sequence position as categorical variable (1 vs. 2 vs. 3, mR² = 0.00, cR² = 0.00), SEPR was characterized by an interaction of group x sequence position (F(2,133893) = 3.43, p = .032, BF < 1), which was not moderated by the inclusion of covariates (F(2,133891) = 3.45, p < .031, supplement 15). Post-hoc analysis showed that this was driven by a lower SEPR to pitch oddballs at sequence position 2 in autistic versus non-autistic individuals (Δβ = -0.10 [-0.18, -0.02]). This was not supported by Bayesian posterior sampling that showed no group differences in posterior estimates.

MMN-amp differed between groups for sequence position as continuous variable (mR² = 0.01, cR² = 0.13), MMN-amp was characterized by an interaction of group x stimulus (F(3, 37901) = 2.29, p = 0.045, BF < 1), which was not moderated by the inclusion of covariates (F(3, 37906) = 2.67, p = 0.045, supplement 17). Post-hoc analysis showed a higher MMN-amp for oddball versus standard trials in autistic individuals (pitch: Δβ = 0.10 [0.06, 0.15]; length: Δβ = 0.15 [0.10, 0.19]; pitch & length: Δβ = 0.12 [0.07, 0.17]), compared to non-autistic individuals (pitch: Δβ = 0.02 [-0.03, 0.08]; length: Δβ = 0.07 [0.01, 0.13]; pitch & length: Δβ = 0.05 [0.00, 0.11]). This was not supported by Bayesian posterior sampling that showed no group differences in posterior estimates. For sequence position as categorical variable, MMN-amp did not differ between groups (F(3, 19273 = 1.57, p = 0.19).

# supplement 13 

```{r BPS linear mixed models - habituation}

htmltools::includeHTML("./output/supplements/table_model_BPS_habituation.html")

```

# supplement 14 

```{r SEPR linear mixed models - task progression}

htmltools::includeHTML("./output/supplements/table_model_SEPR.html")

htmltools::includeHTML("./output/supplements/table_SEPR_covariateeffects.html")

```

# supplement 15 

```{r SEPR linear mixed models - habituation}

htmltools::includeHTML("./output/supplements/table_model_SEPR_habituation.html")

```

# supplement 16 

```{r MMN linear mixed models - task progression}

htmltools::includeHTML("./output/supplements/table_model_MMN.html")

htmltools::includeHTML("./output/supplements/table_MMN_covariateeffects.html")

```

# supplement 17 


```{r MMN linear mixed models - habituation}

htmltools::includeHTML("./output/supplements/table_model_MMN_habituation.html")

```

# supplement 18

```{r Bayesian posterior estimate chain convergence,as.is=TRUE}

grid::grid.raster(tiff::readTIFF("./output/supplements/Bayesian_chains_convergence.tiff"))

```

# supplement 19

```{r histograms per-partiticpant}

grid::grid.raster(tiff::readTIFF("./output/supplements/histograms_dependentvariables_perparticipant.tiff") )

```

# supplement 20

```{r histograms per-trial}

grid::grid.raster(tiff::readTIFF("./output/supplements/histograms_dependentvariables_pertrial.tiff") )


```

