---
title: "supplementary information"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(htmltools)
require(knitr)
require(tiff)

project_path<-getwd()

```

# table of content

abbreviations: baseline pupil size (BPS), stimulus-evoked pupillary response (SEPR), mismatch negativity (MMN), neural gain (NG)

-   [supplement 1 - descriptive statistics per participant][supplement 1]
-   [supplement 2 - descriptive statistics per trial][supplement 2]
-   [supplement 3 - group comparisons on per-participant estimates][supplement 3]
-   [supplement 4 - covariate effects: BPS group difference][supplement 4]
-   [supplement 5 - covariate effects: NG group difference][supplement 5]
-   [supplement 6 - BPS association with age, IQ, biological sex][supplement 6]
-   [supplement 7 - SEPR association with age, IQ, biological sex][supplement 7]
-   [supplement 8 - MMN association with age, IQ, biological sex][supplement 8]
-   [supplement 9 - correlation table of BPS, SEPR, MMN, NG][supplement 9]
-   [supplement 10 - polynomial fit comparison for BPS, SEPR; MMN][supplement 10]
-   [supplement 11 - BPS linear mixed model - task progression][supplement 11]
-   [supplement 12 - BPS linear mixed model - habituation][supplement 12]
-   [supplement 13 - SEPR linear mixed model - task progression][supplement 13]
-   [supplement 14 - SEPR linear mixed model - habituation][supplement 14]
-   [supplement 15 - MMN linear mixed model - task progression][supplement 15]
-   [supplement 16 - MMN linear mixed model - task progression][supplement 16]
-   [supplement 17 - Bayesian posterior estimate MCMC chain convergence][supplement 17]
-   [supplement 18 - histograms of per-participant estimates][supplement 18]
-   [supplement 19 - histograms of per-trial estimates][supplement 19]


# supplement 1

```{r descriptive statistics per participant}

htmltools::includeHTML("./output/supplements/descriptivestatistics_DV_perparticipant.html")

```

# supplement 2

```{r descriptive statistics per trial}

htmltools::includeHTML("./output/supplements/descriptivestatistics_DV_pertrial.html")

```

# supplement 3 

```{r group comparisons per participant}

htmltools::includeHTML("./output/supplements/table_groupcomparison_perparticipant.html")

```

# supplement 4 

```{r covariate effects BPS}

htmltools::includeHTML("./output/supplements/table_linearmodel_BPS_allcovariates.html")

htmltools::includeHTML("./output/supplements/table_linearmodel_BPS_signcovariates.html")


```

# supplement 5 

```{r covariate effects NG}

htmltools::includeHTML("./output/supplements/table_linearmodel_NG_allcovariates.html")

htmltools::includeHTML("./output/supplements/table_linearmodel_NG_signcovariates.html")


```

# supplement 6

```{r figure BPS scaling}

grid::grid.raster(tiff::readTIFF("./output/supplements/BPS_scaling_withdemographics_between_groups.tiff") )

```

# supplement 7

```{r figure SEPR scaling}

grid::grid.raster(tiff::readTIFF("./output/supplements/SEPR_scaling_withdemographics_between_groups.tiff") )

```

# supplement 8

```{r figure MMN scaling}

grid::grid.raster(tiff::readTIFF("./output/supplements/MMN_scaling_withdemographics_between_groups.tiff") )

```

# supplement 9 

```{r correlation table}

htmltools::includeHTML("./output/supplements/associations_DV_perparticipant.html")


```

# supplement 10 

```{r polynomial fit comparisons}

htmltools::includeHTML("./output/supplements/table_modelcomparison_linearity_BPS.html")

htmltools::includeHTML("./output/supplements/table_modelcomparison_linearity_SEPR.html")

htmltools::includeHTML("./output/supplements/table_modelcomparison_linearity_MMN.html")

```

# supplement 11 

```{r BPS linear mixed models - task progression}

htmltools::includeHTML("./output/supplements/table_model_BPS.html")

htmltools::includeHTML("./output/supplements/table_BPS_covariateeffects.html")

```

# supplement 12 

```{r BPS linear mixed models - habituation}

htmltools::includeHTML("./output/supplements/table_model_BPS_habituation.html")

```

# supplement 13 

```{r SEPR linear mixed models - task progression}

htmltools::includeHTML("./output/supplements/table_model_SEPR.html")

htmltools::includeHTML("./output/supplements/table_SEPR_covariateeffects.html")

```

# supplement 14 

```{r SEPR linear mixed models - habituation}

htmltools::includeHTML("./output/supplements/table_model_SEPR_habituation.html")

```

# supplement 15 

```{r MMN linear mixed models - task progression}

htmltools::includeHTML("./output/supplements/table_model_MMN.html")

htmltools::includeHTML("./output/supplements/table_MMN_covariateeffects.html")

```

# supplement 16 


```{r MMN linear mixed models - habituation}

htmltools::includeHTML("./output/supplements/table_model_MMN_habituation.html")

```

# supplement 17

```{r Bayesian posterior estimate chain convergence,as.is=TRUE}

grid::grid.raster(tiff::readTIFF("./output/supplements/Bayesian_chains_convergence.tiff"))

```

# supplement 18

```{r histograms per-partiticpant}

grid::grid.raster(tiff::readTIFF("./output/supplements/histograms_dependentvariables_perparticipant.tiff") )

```

# supplement 19

```{r histograms per-trial}

grid::grid.raster(tiff::readTIFF("./output/supplements/histograms_dependentvariables_pertrial.tiff") )


```

